---
title: ""
bibliography: references.bib
---

```{r, echo = FALSE}
library(lme4)
library(lattice)
library(sjstats)
library(jtools)
library(ROCR)
library(sjPlot)
library(performance)
library(XML)
library(kableExtra)
library(tidyverse)
library(caTools)
library(fastDummies)
library(semptools)
library(semoutput)
library(lavaan)
library(semPlot)
library(stargazer)
library(semTable)
library(rstatix)
library(gtsummary)
library(forcats)
library(haven)
load("waves123.long.rda")
load("waves123.long.complete.rda")
load("waves123.wide.rda")
load("waves12.wide.complete.rda")
load("waves23.wide.complete.rda")
load("waves13.wide.complete.rda")
# Tutorials
# https://lmudge13.github.io/sample_code/mixed_effects.html
# Converting html to pdf https://github.com/gorkang/html2latex/
```

```{r path-data-recodes, echo = FALSE, results='hide'}
# Order Education levels
path.waves123.wide <- waves123.wide %>%
  mutate(waves123.wide = factor(Education_W1, levels = c("Less than High School", "HS Diploma/GED", 
                                                "Some College/AA Degree", "College Grad",
                                                "Graduate or Professional Degree"), labels = 1:5, ordered = TRUE))
# Convert to numeric to be used as ordinal variable
path.waves123.wide$Education_W1 <- as.numeric(path.waves123.wide$Education_W1)
# Order education levels
path.waves123.wide <- path.waves123.wide %>%
  mutate(path.waves123.wide = factor(Education_W2, levels = c("Less than High School", "HS Diploma/GED", 
                                                "Some College/AA Degree", "College Grad",
                                                "Graduate or Professional Degree"), labels = 1:5, ordered = TRUE))
# Convert to numeric to be used as ordinal variable
path.waves123.wide$Education_W2 <- as.numeric(path.waves123.wide$Education_W2)
# Order Education levels
path.waves123.wide <- path.waves123.wide %>%
  mutate(path.waves123.wide = factor(Education_W3, levels = c("Less than High School", "HS Diploma/GED", 
                                                "Some College/AA Degree", "College Grad",
                                                "Graduate or Professional Degree"), labels = 1:5, ordered = TRUE))
# Convert to numeric to be used as ordinal variable
path.waves123.wide$Education_W3 <- as.numeric(path.waves123.wide$Education_W3)

# Create dummy variables for race
path.waves123.wide <- dummy_cols(path.waves123.wide, select_columns = 'Race_W2', ignore_na = TRUE)
path.waves123.wide <- rename(path.waves123.wide, 
                             MexicanHispanic_W2 = "Race_W2_Mexican American and Other Hispanic",
                             Black_W2 = "Race_W2_Non-Hispanic Black", 
                             OtherRace_W2 = "Race_W2_Other Race",
                             White_W2 = "Race_W2_Non-Hispanic White")

# Order SRH levels
path.waves123.wide <- path.waves123.wide %>%
  mutate(waves123.wide = factor(SRH_W2, levels = c("Poor", "Fair", "Good",
                                                   "Very Good", "Excellent"), 
                                labels = 1:5, 
                                ordered = TRUE))
# Convert to numeric to be used as ordinal variable
path.waves123.wide$SRH_W2 <- as.numeric(path.waves123.wide$SRH_W2)

# Order SRH levels
path.waves123.wide <- path.waves123.wide %>%
  mutate(waves123.wide = factor(SRH_W3, levels = c("Poor", "Fair", "Good",
                                                   "Very Good", "Excellent"), 
                                labels = 1:5, 
                                ordered = TRUE))
# Convert to numeric to be used as ordinal variable
path.waves123.wide$SRH_W3 <- as.numeric(path.waves123.wide$SRH_W3)
```

```{r listwise-delete-path-data, echo =FALSE}
myvars <- c("mindbody_W2", "mindbody_W3", "altmed_W2", "altmed_W3", "pna_W2", "pna_W3", "manipulative_W2", "manipulative_W3", "totalCams_W2", "totalCams_W3", "SumChron_W2", "SumChron_W3", "SRH_W2", "SRH_W3", "LogIncome_W1", "Education_W1", "Sex_W1", "Age_W2", "Age_W3", "Marital_W2", "Marital_W3", "HlthInsurance_W2", "HlthInsurance_W3", "HLSelf_W2", "HLSelf_W3", "Spiritual_W2", "Spiritual_W3", "Religion_W2", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Diabetes_W1", "Diabetes_W2", "Diabetes_W3", "Respiratory_W1", "Respiratory_W2", "Respiratory_W3", "DepAnx_W1", "DepAnx_W2", "DepAnx_W3", "HighBP_W1", "HighBP_W2", "HighBP_W3", "ChronicPain_W2", "ChronicPain_W3", "Depression_W1", "Depression_W2", "Depression_W3")

path.complete <- path.waves123.wide[myvars]

path.complete <- path.complete[complete.cases(path.complete),]

```

```{r, echo = FALSE}
require(foreign)
write.dta(path.complete, "path.dta")
library(Statamarkdown)

stataexe <- "C:/Program Files/Stata17/StataBE-64.exe"
knitr::opts_chunk$set(engine.path=list(stata=stataexe))

```

# Chapter 3: Alternative Health Lifestyle as Mediator of the SES-Health Relationship

## Questions and Hypotheses

Health lifestyle represents an important pathway by which socioeconomic status, contributing to the well-known positive association between social position and health (i.e. the social gradient in health) [@adler1994; @brunello2016; @burdette2008; @lantz1998; @lantz2001]. This chapter considers the role of AHL as a mediator of the socioeconomic status-health relationship by examining the relationships between SES, CAM use, and health. Do individuals with higher SES (education and income) use more CAMs? And does this CAM use contribute to better health for those with higher SES? In short, can CAM use explain some of the association between SES and health. The alternative is that the relationship between SES and health is unrelated to individuals' use of CAMs.

From a public health perspective, if health lifestyles can explain some of the relationship between SES and health, then campaigns to alter individual behaviors have the potential to address health disparities. However, an emphasis on individual behavioral attributes as the source of health lifestyles ignores the social context within which behaviors are enacted [@frohlich2001]. Health lifestyle theory emphasizes that behaviors are not simply the product of individual motivations or will but reflect the enabling and constraining aspects of structural position [@abel2012; @cockerham2005; @frohlich2001]. Thus, I expect to find that individuals sharing a similar socioeconomic position will also share similar health lifestyles, reflecting the important role of life circumstances and social experience in shaping health lifestyles.

To investigate CAM use as a potential mediator of the SES-health relationship, I conducted a two-part analysis. First, I address the question of whether CAM use is related to SES as measured by income and education. The CAM use measures include total CAMs used in the past year as well as binary indicators of whether respondents use each of the types of CAMs identified in the previous chapter (mind-body medicines, alternative medicines, physical and nutritional approaches, and manipulative treatments). I hypothesize that CAM use is positively related to income (*Hypothesis 1a*) and education (*Hypothesis 1b*).

-   *Hypothesis 1* SES is positively related to CAM use.
    -   *Hypothesis 1a* Income will be positively related to CAM use.
    -   *Hypothesis 1b* Education will be positively related to CAM use.

In the second step, I examine two possible mediation processes. Path models will provide a longitudinal test of *Hypothesis 1*, as well as test the relationship between SES and health. As previous research has observed, I expect to find higher SES to be related to better health outcomes (*Hypothesis 2*).

-   *Hypothesis 2* SES is positively related to health.
    -   *Hypothesis 2a* Income will be positively related to health.
    -   *Hypothesis 2b* Education will be positively related to health.

The first mediation process I will examine is whether CAM use mediates the relationship between SES and health. More specifically, CAM use may contribute to better health so that more CAM use among higher SES groups contributes to better health outcomes in these groups. In short, I expect that individuals with high SES are more likely to use CAMs, and that this in turn leads to better health outcomes (*Hypothesis 3*).

-   *Hypothesis 3* CAM use will mediate the relationship between SES and health, with SES being positively related to CAM use, and CAM use being positively related to health.
    -   *Hypothesis 3a* CAM use will mediate the relationship between income and health, with income being positively related to CAM use, and CAM use being positively related to health.
    -   *Hypothesis 3b* CAM use will mediate the relationship between education and health, with education being positively related to CAM use, and CAM use being positively related to health.

The second mediation process I examine is whether health mediates the relationship between SES and CAM use. While I expect that CAM use influences health, it is also possible that health influences CAM use. Individuals who experience a health crisis or develop chronic health conditions may be more likely to use CAM to address their health problems. Thus, it is important to examine both potential directions of this relationship to tease out whether individuals tend to use CAMs for prevention or treatment of disease. I expect to find that SES influences health, and that health in turn influences CAM use (*Hypothesis 4*).

-   *Hypothesis 4* Health will mediate the relationship between SES and CAM use, with SES being positively related to health, and individuals who report poor health being more likely to utilize CAMs.
    -   *Hypothesis 4a* Health will mediate the relationship between income and CAM use, with income being positively associated with health, and individuals with poor health being more likely to utilize CAMs.
    -   *Hypothesis 4b* Health will mediate the relationship between education and CAM use, with education being positively associated with health, and individuals with poor health being more likely to utilize CAMs.

## Data and Measures

To test these hypotheses, I use two samples from the MIDUS survey (See Chapter 1). In the first stage of analysis I use generalized linear mixed modeling to analyze all respondents that responded to at least two waves (*n* = `r length(unique(waves123.long.complete$M2ID))` ). The second stage of analysis uses structural equation modeling (SEM) and includes only respondents with complete data on all variables that were included in analyses (*n* = `r nrow(path.complete)`).

CAM variables include total CAMs, mind-body medicine, alternative medicine, physical and nutritional approaches, and manipulative treatments. Total CAMs was calculated by summing the number of CAMs used in the past year at each wave. Mind-body medicine,alternative medicine, physical and nutritional approaches, and manipulative treatments are dummy variables indicating whether the respondent used any of the associated CAMs. In accordance with the factor analysis from the previous chapter, mind-body medicines include prayer or spiritual healing, energy healing, imagery techniques, relaxation and meditation techniques, and hypnosis. Alternative medicines or medical systems include homeopathy, herbal therapy, and acupuncture. Physical and nutritional approaches include vitamins, exercise and movement, and special diets. Manipulative treatments include massage and chiropractic care.

The analyses utilize multiple measures of health. There are two measures of general health:e th sum of chronic conditions and self-reported health (SRH). The sum of chronic conditions is simply the total chronic conditions reported. Chronic conditions in the MIDUS survey include: asthma, bronchitis, or emphysema; tuberculosis; other lung problems; bone or joint diseases; sciatica, lumbago, recurring backache; persistent skin trouble; thyroid disease; hay fever; recurring stomach trouble; urinary and bladder problems; being constipated; gall bladder trouble; persistent foot trouble; varicose veins requiring treatment; AIDs or HIV infection; autoimmune disorders; trouble with gums or mouth; persistent trouble with teeth; high blood pressure; emotional disorder; alcohol or drug problems; migraine headaches; chronic sleeping problems; diabetes or high blood sugar; neurological disorders; stroke; ulcer; hernia or rupture; and piles or hemorrhoids. Self-reported health is measured as the respondent's response to the question "In general, would you say your physical health is excellent, very good, good, fair, or poor?" Responses are coded as 1 through 5, with higher values indicating better health.

There are four measures of specific chronic conditions: diabetes, chronic pain, high blood pressure, and depression/anxiety. Respondents were asked whether they had experienced each chronic condition in the past 12 months. Each condition is coded as a dummy variable with 1 representing having experienced or been treated for the condition in the past 12 months.

Socioeconomic status is measured with income and education. Income represents household income. The original variable is coded in US dollars and top coded at \$300,000. Income is logged in analyses to address skewness. The variable for education is an ordinal variable consisting of the following categories: less than high school, high school diploma or GED, some college or an associate's, college graduate, and graduate or professional degree. College graduate is the reference category in analyses.

Race, sex, marital status, health insurance, age, spirituality and religiosity are included as controls. Race at wave 1 includes categories for non-Hispanic White, non-Hispanic Black, and other race. Race additionally includes a category for Mexican American and other Hispanic at waves 2 and 3. White is the reference category in analyses. Sex is a dummy variable coded as 1 = female, 0 = male. Health insurance is a dummy variable coded as 1 = insured, 0 = uninsured. Age is measured in years. Spirituality is a summed index ranging from 0 to 8. Respondents were asked to indicate on a scale of 1 to 4 (1 = Very, 2 = Somewhat, 3 = Not very, and 4 = Not at all) their response to the following questions: "How spiritual are you?" and "How important is spirituality in your life?". Religiosity is a summed index ranging from 0 to 24 based on responses on a scale of 1 to 4 (1 = Very, 2 = Somewhat, 3 = Not very, and 4 = Not at all) for the following questions: "How religious are you?"; "How important is religion in your life?"; "How important is it for you -- or would it be if you had children now -- to send your children for religious or spiritual services or instruction?"; "How closely do you identify with being a member of your religious group?"; "How much do you prefer to be with other people who are the same religion as you?"; "How important do you think it is for people of your religion to marry other people who are the same religion?"; "How important is it for you to celebrate or practice on religious holidays with your family, friends, or members of your religious community?"

## Methods

This chapter examines whether socioeconomic status (SES) predicts the use of mind body approaches, alternative medical systems, physical and nutritional approaches, manipulative treatments, and total CAMs. To address this question, I used generalized linear mixed models (GLMMs). I used logistic GLMMs to model the effect of SES on each of the CAM factor indices and a Poisson GLMM to model the effect of SES on total CAM use.

GLMMs are an extension of generalized linear models that account for the nested nature of data. In the MIDUS data, each respondent participated in the survey at one to three time points. Therefore measures at different time points are nested within the individual. GLMMs include fixed effects for level 1 predictors and random effects for level 2. The fixed effects in the models below include all predictors and covariates. Like in a fixed effects model, the random effects represent the unmeasured characteristics of the individual respondents. In this type of GLMM for change, the intercept for each individual can vary randomly. Analyses are completed in R using the lme4 package.

Then I conducted mediation analyses using structural equation modeling (SEM). In cases like this, where all variables are observed as opposed to unobserved (i.e. latent), it is also known as path analysis. SEM is an extension of multiple regression. It allows for the inclusion of more than one dependent variable in a model, and allows for the testing of models that specify chains of causality. The SEM models in this chapter simultaneously estimate two mediation processes. First, I examine CAM use as a mediator of the relationship between SES and health. Second, I analyze health as a mediator of the relationship between SES and CAM use. I measured SES by income and education at wave 1. I measured CAM use and health at waves 2 and 3, with the wave 2 variables representing mediators and the wave 3 variables being the outcome of interest. Wave 3 variables are lagged, meaning the wave 2 variable is included as a control. This allows the models to isolate the effect of the mediator, independent of the effect of the outcome variable at the previous wave.

Analyses includes tests of the relationships between all CAMs and measures of general health (sum of chronic conditions and self-reported health). The review of CAM efficacy in Chapter 1 guided the selection of specific health outcomes to be included in models with each CAM. Analyses test the relationship between total CAMs used and measures of general health, diabetes, chronic pain, high blood pressure, and depression/anxiety. Analyses test the relationship between use of mind-body medicine and measures of general health, chronic pain, and depression/anxiety. Analyses test the relationship between use of alternative medical systems and measures of general health, high blood pressure, diabetes, chronic pain, and depression/anxiety. Analyses test the relationships between use of physical and nutritional approaches and measures of general health, high blood pressure, chronic pain, and depression/anxiety. Finally, analyses test the relationships between use of manipulative treatments and measures of general health, chronic pain, and depression/anxiety.

An important aspect of SEM is model fit. Fit refers to how well the relationships in the model reproduce the relationships observed in the data. Goodness of fit indices used in the following analyses to assess model fit include $\chi$^2^ tests, comparative fit index (CFI), Tucker Lewis Index (TLI), root mean square error of approximation (RMSEA), and standardized root mean square residual (SRMR). A non-significant $\chi$^2^ indicates good model fit, though the test is affected by sample size, with larger sample sizes almost always producing a significant result. Cutoffs indicating good fit include CFI \> 0.90, TLI \> 0.90, RMSEA \< 0.08, and SRMR \< 0.08.

I first completed the analyses in R version 4.2.0 using the lavaan program. However, the outcome variables in the models are a mixture of binary, count, and ordinal outcomes. Lavaan does not allow the user to specify non-normal distributions for endogenous (i.e. outcome) variables, assuming all variables are normally distributed. To address this violation of the model assumptions, I also completed analyses in Stata Version 17 using the gsem command for generalized SEM (GSEM). The gsem command allows the user to specify non-normal distributions such as logit or poisson. In GSEM analyses below, I estimated total CAMs with a Poisson distribution, binary outcomes with a binomial distribution and logit link, and SRH, an ordinal outcome, with an ordinal distribution and ordered probit link function. Goodness of fit indices presented below are from SEM analyses in lavaan. Goodness of fit indices are currently not available in gsem. Both SEM analyses using lavaan and GSEM analyses using gsem are presented below.

```{r wave-comp-educ-recodes1, echo=FALSE}


#  Education.2 so that college graduate is the reference category. 
waves123.long.complete$Education.2 <- factor(waves123.long.complete$Education, 
                                              levels = c("College Grad", 
                                                         "Less than High School", 
                                                         "HS Diploma/GED", 
                                                         "Some College/AA Degree", 
                                                         "Graduate or Professional Degree"))

waves12.wide.complete$Education_W1.2 <- factor(waves12.wide.complete$Education_W1, 
                                              levels = c("College Grad", 
                                                         "Less than High School", 
                                                         "HS Diploma/GED", 
                                                         "Some College/AA Degree", 
                                                         "Graduate or Professional Degree"))


waves23.wide.complete$Education_W2.2 <- factor(waves23.wide.complete$Education_W2, 
                                              levels = c("College Grad", 
                                                         "Less than High School", 
                                                         "HS Diploma/GED", 
                                                         "Some College/AA Degree", 
                                                         "Graduate or Professional Degree"))

waves13.wide.complete$Education_W1.2 <- factor(waves13.wide.complete$Education_W1, 
                                              levels = c("College Grad", 
                                                         "Less than High School", 
                                                         "HS Diploma/GED", 
                                                         "Some College/AA Degree", 
                                                         "Graduate or Professional Degree"))

```

## Results

Table \ref{tab:descriptives-table} shows summary statistics for variables across all three waves included in the following analyses along with information about missing data.

```{r descriptives-table, echo = FALSE}
waves123.long %>% 
  ungroup %>% 
  dplyr::select(wave, totalCams, mindbody, altmed, pna, manipulative, SumChron, SRH, 
         Income, Education, Race, Sex, Marital, HlthInsurance, Age, Spiritual, Religion) %>%
  gtsummary::tbl_summary(
    by = wave,
    type = list(Spiritual ~ "continuous2",
                totalCams ~ "continuous2", 
                SumChron ~ "continuous2", 
                Income ~ "continuous2", 
                Age ~ "continuous2", 
                Religion ~ "continuous2"),
    statistic = list(all_continuous() ~ c("{N_nonmiss}/{N_obs}", 
                                          "{mean} ({sd})",
                                          "{min}, {max}"),
                     all_categorical() ~ "{n}/{N_nonmiss} ({p})%"),
    digits = all_continuous() ~ 2, 
    label = list(totalCams ~ "Total CAMs", 
                 mindbody ~ "Mind-Body Medicine",
                 altmed ~ "Alternative Medicines", 
                 pna ~ "Physical and Nutritional Approaches",
                 manipulative ~ "Manipulative Treatments", 
                 SumChron ~ "Sum of Chronic Conditions", 
                 SRH ~ "Self-Reported Health",
                 Marital ~ "Marital Status",
                 HlthInsurance ~ "Health Insurance",
                 Spiritual ~ "Spirituality",
                 Religion ~ "Religiosity"),
    missing_text = "(Missing)") %>% 
    bold_labels() %>% 
    as_kable_extra(
      longtable = TRUE,
      booktabs=TRUE,
      caption="Descriptive Statistics All Waves",
      escape = F,
      col.names = c("Variable", "Wave 1", "Wave 2", "Wave 3")) %>%
  kable_classic(html_font = "CMU Serif") %>%
  kable_styling(latex_options = c("striped", "repeat_header"),
                font_size = 7) 
```

### Relationship Between SES and CAM Use

#### Mind Body

```{r model-glmm-mb, echo = FALSE}
# baseline model 
baseline <- glmer(mindbody ~ 1 + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Model with income
mb1 <- glmer(mindbody ~ LogIncome + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Model with education
mb2 <- glmer(mindbody ~ Education.2 + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Model with income and education
mb3 <- glmer(mindbody ~ LogIncome + Education.2 + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

#sjPlot::plot_model(mb2, 
#                   show.values=TRUE, show.p=TRUE,
#                   title="Effect of Income and Education on Mind Body Medicine Use")


# Model with all covariates
mb4 <- glmer(mindbody ~ LogIncome + Education.2 + Age + Sex + Race + Marital + HlthInsurance + HLSelf + Spiritual + Religion + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Store results for reporting 
summ.mb1 <- summary(mb1)
summ.mb2 <- summary(mb2)
summ.mb3 <- summary(mb3)
summ.mb4 <- summary(mb4)

# Create table with results
mb.tab.model <- tab_model(mb1, mb2, mb3, mb4,
          show.intercept = FALSE,
          pred.labels = c("Logged Income",  "Less than high school", "High School Diploma/GED",
                          "Some College/AA", "Graduate/Professional Degree", "Age", 
                          "Female", "Non-Hispanic Black", "Other Race", 
                          "Mexican American and Other Hispanic", "Married", "Insured",
                          "Health Locus of Control", "Spirituality", "Religiosity"),
          dv.labels= c("Model 1", "Model 2", "Model 3", "Model 4"),
          title = "Mixed Model: Mind Body Approaches")

 # first you parse the html, then you put it as table, taking the first
df <- data.frame(readHTMLTable(htmlParse(mb.tab.model))[1])
# first row as colnames
colnames(df) <- df[1,]
# remove the fake first row
df <- df[-1,]
```

Table \ref{tab:mindbody-fixed-table} shows results for the level 1 predictors and Table \ref{tab:mindbody-random-table} shows results for the level 2 predictors for all mind body approaches. Model 1 assesses the association between the use of mind body approaches and logged income. It reveals that the odds of using mind body approaches at different levels of logged income do not differ significantly. The marginal $R^2$ is 0.00 and the conditional $R^2$ is 0.47, indicating that income does not explain any of the variance in the model.

```{r mindbody-fixed-table, echo=FALSE}
# Fixed effects table
# Drop random effects rows
df.1 <- df[c(1:15),]
# Drop row names
rownames(df.1)<-NULL
kable(df.1,
      booktabs = T,
      caption = "Mind Body Models: Fixed Effects") %>%
  add_header_above(c(" " = 1, "Model 1" = 3, "Model 2" = 3, 
                     "Model 3" = 3, "Model 4" = 3)) %>%
  kable_classic(html_font = "CMU Serif") %>% 
  kableExtra::landscape() %>% 
  kable_styling(latex_options = c("striped", "scale_down"))

# Compare models
# anova(mb1, mb2, mb3, mb4)
```

```{r mindbody-random-table, echo=FALSE}
# Random Effects Table 
# Drop fixed effects rows
df.1 <- df[c(17:22),]
# change column names
colnames(df.1) <- c(" ", "Model 1", "Model 2", "Model 3", "Model 4", "6", "7", "8", "9")
# Drop NA columns 
df.2 <- df.1 %>% 
  dplyr::select(-(6:13)) %>% 
  dplyr::select(-(1))

# Rename rows 
rownames(df.2) <- c("$\\sigma^2$", "$\\tau$00", "ICC", "N", "Observations", "Marginal R$^2$/Conditional R$^2$")

kable(df.2,
      booktabs = T,
      escape = F,
      caption = "Mind Body Models: Random Effects") %>%
   kable_classic(html_font = "CMU Serif") %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>% 
  footnote(general = c("$\\\\sigma^2$ = within subject variance. 
                       $\\\\tau$00 = between subject variance.
                       ICC = intraclass correlation coefficient"), 
           escape = F)

```

Model 2 assesses the association between the use of mind body approaches and levels of education, using a reference category of college degree. The odds of using mind body approaches increase with each level of education, so that the odds of using mind-body approaches are 63% lower for those with less than high school (OR = `r round(exp(summ.mb2$coefficients[2]),3)`, 95% CI: `r round(exp(summ.mb2$coefficients[2] - 1.96 * summ.mb2$coefficients[7]),3)`, `r round(exp(summ.mb2$coefficients[2] + 1.96 * summ.mb2$coefficients[7]),3)`, *p* \< 0.001), 50% lower for those with a high school diploma or GED (OR = `r round(exp(summ.mb2$coefficients[3]),3)`, 95% CI: `r round(exp(summ.mb2$coefficients[3] - 1.96 * summ.mb2$coefficients[8]),3)`, `r round(exp(summ.mb2$coefficients[3] + 1.96 * summ.mb2$coefficients[8]),3)`, *p* \< 0.001), and 25% lower for those with some college or an associate's degree (OR = `r round(exp(summ.mb2$coefficients[4]),3)`, 95% CI: `r round(exp(summ.mb2$coefficients[4] - 1.96 * summ.mb2$coefficients[9]),3)`, `r round(exp(summ.mb2$coefficients[4] + 1.96 * summ.mb2$coefficients[9]),3)`, *p* \< 0.01) compared to those with a college degree. There is no significant difference in use of mind-body medicines between those with a college degree and those with a graduate or professional degree. The marginal $R^2$ is 0.018 and the conditional $R^2$ is 0.467, indicating that education explains approximately 1.8% of the variance in the model. Models 1 and 2 suggest that education has a strong, significant, direct relationship with use of mind-body approaches, while income does not.

Model 3 includes both logged income and education as predictors of the use of mind body approaches. When accounting for education, logged income has a significant negative effect. For each one unit increase in logged income the odds of using mind body approaches decreases by 11% (OR = `r round(exp(summ.mb3$coefficients[2]),3)`, 95% CI: `r round(exp(summ.mb3$coefficients[2] - 1.96 * summ.mb3$coefficients[8]),3)`, `r round(exp(summ.mb3$coefficients[2] + 1.96 * summ.mb3$coefficients[8]),3)`, *p* \< 0.001). In this model, the odds of using mind-body approaches are 67% lower for those with less than a high school education (OR = `r round(exp(summ.mb3$coefficients[3]),3)`, 95% CI: `r round(exp(summ.mb3$coefficients[3] - 1.96 * summ.mb3$coefficients[9]),3)`, `r round(exp(summ.mb3$coefficients[3] + 1.96 * summ.mb3$coefficients[9]),3)`, *p* \< 0.001), 53% lower for those with a high school diploma or GED (OR = `r round(exp(summ.mb3$coefficients[4]),3)`, 95% CI: `r round(exp(summ.mb3$coefficients[4] - 1.96 * summ.mb3$coefficients[10]),3)`, `r round(exp(summ.mb3$coefficients[4] + 1.96 * summ.mb3$coefficients[10]),3)`, *p* \< 0.001), and 28% lower for those with some college or an associate's degree (OR = `r round(exp(summ.mb3$coefficients[5]),3)`, 95% CI: `r round(exp(summ.mb3$coefficients[5] - 1.96 * summ.mb3$coefficients[11]),3)`, `r round(exp(summ.mb3$coefficients[5] + 1.96 * summ.mb3$coefficients[11]),3)`, *p* \< 0.001) compared to those with a college degree. There was no significant difference in mind-body medicine use between those with a college degree and those with a graduate or professional degree. The marginal $R^2$ is 0.020 and the conditional $R^2$ is 0.466, indicating that income and education together explain approximately 2% of the variance explained in the model.

Model 4 adds control variables and logged income is no longer a significant predictor of the use of mind body approaches. The effects of education remain significant in this model, and the odds increase with each level of education. The odds of using mind-body approaches are 61% lower for those with less than a high school education (OR = `r round(exp(summ.mb4$coefficients[3]),3)`, 95% CI: `r round(exp(summ.mb4$coefficients[3] - 1.96 * summ.mb4$coefficients[19]),3)`, `r round(exp(summ.mb4$coefficients[3] + 1.96 * summ.mb4$coefficients[19]),3)`, *p* \< 0.001), 54% lower for those with a high school diploma or GED (OR = `r round(exp(summ.mb4$coefficients[4]),3)`, 95% CI: `r round(exp(summ.mb4$coefficients[4] - 1.96 * summ.mb4$coefficients[20]),3)`, `r round(exp(summ.mb4$coefficients[4] + 1.96 * summ.mb4$coefficients[20]),3)`, *p* \< 0.001), and 33% lower for those with some college or an associate's degree (OR = `r round(exp(summ.mb4$coefficients[5]),3)`, 95% CI: `r round(exp(summ.mb4$coefficients[5] - 1.96 * summ.mb4$coefficients[21]),3)`, `r round(exp(summ.mb4$coefficients[5] + 1.96 * summ.mb4$coefficients[21]),3)`, *p* \< 0.001) compared to those with a college degree. There was no significant difference in mind-body medicine use between those with a college degree and those with a graduate or professional degree. The marginal $R^2$ is 0.217 and the conditional $R^2$ is 0.453, indicating that income, education, and the control variables explain 21.7% of the variance in the model. Taken together, the model results in Table \ref{tab:mindbody-fixed-table} suggest a robust relationship between education and mind-body CAM use, and inconsistent effects of income.

Notably, the control variables make substantial contributions to the variance explained in the model. The odds of using mind-body approaches are higher for younger respondents, decreasing 3% for each additional year in age. The odds of using mind-body approaches are 50% higher for women, 46% higher for Whites compared to Blacks, 32% higher for Whites compared to other race, and 19% lower for married compared to not married respondents. The odds of using mind-body approaches are positively related to both spirituality and religiosity.

As Table \ref{tab:mindbody-random-table} illustrates, within-subject variance is 3.29 across all 4 models and between-subject variance ranges from 1.42 in Model 4 to 2.91 in Model 1. The Intraclass Correlation Coefficient also ranges from 0.30-0.47, indicating that 30-47% of the variation in use of mind-body approaches occurs between-subjects. Thus while income and education significantly predict the use of mind-body approaches, most of the variation in the use of these CAMs is explained by differences between-subjects that the data do not account for rather than by the predictors in the models. Also, it is notable that income has no significant effect without accounting for education, which suggests that income matters for mind-body CAM use within education groups but only when the larger effects of education are taken into account.

#### Alternative medicine

```{r model-glmm-altmed, echo = FALSE}
# baseline model 
baseline <- glmer(altmed ~ 1 + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Model with income
am1 <- glmer(altmed ~ LogIncome + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Model with education
am2 <- glmer(altmed ~ Education.2 + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Model with income and education
am3 <- glmer(altmed ~ LogIncome + Education.2 + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

#sjPlot::plot_model(am2, 
#                   show.values=TRUE, show.p=TRUE,
#                   title="Effect of Income and Education on Mind Body Medicine Use")


# Model with all covariates
am4 <- glmer(altmed ~ LogIncome + Education.2 + Age + Sex + Race + Marital + HlthInsurance + HLSelf + Spiritual + Religion + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Store results for reporting 
summ.am1 <- summary(am1)
summ.am2 <- summary(am2)
summ.am3 <- summary(am3)
summ.am4 <- summary(am4)

# Create table with results
altmed.tab.model <- tab_model(am1, am2, am3, am4,
          show.intercept = FALSE,
          pred.labels = c("Logged Income", "Less than high school", "High School Diploma/GED",
                          "Some College/AA", "Graduate/Professional Degree", "Age", 
                          "Female", "Non-Hispanic Black", "Other Race", 
                          "Mexican American and Other Hispanic", "Married", "Insured",
                          "Health Locus of Control", "Spirituality", "Religiosity"),
          dv.labels= c("Model 1", "Model 2", "Model 3", "Model 4"),
          title = "Logistic GLMM: Alternative Medical Systems")

 # first you parse the html, then you put it as table, taking the first
df <- data.frame(readHTMLTable(htmlParse(altmed.tab.model))[1])
# first row as colnames
colnames(df) <- df[1,]
# remove the fake first row
df <- df[-1,]
```

Table \ref{tab:altmed-fixed-table} shows results for level 1 predictors and Table \ref{tab:altmed-random-table} shows results for the level 2 predictors for all alternative medicine models. Model 1 assesses the relationship between logged income and the use of alternative medical systems. The odds of using alternative medicines at different levels of logged income do not differ significantly. The marginal $R^2$ is 0.00 and the conditional $R^2$ is 0.462, indicating that income explains 0% of the variance in the model.

```{r altmed-fixed-table, echo=FALSE}
# Fixed effects table
# Drop random effects rows
df.1 <- df[c(1:15),]
# Drop row names
rownames(df.1)<-NULL
kable(df.1,
      booktabs = T,
      caption = "Alternative Medical Systems: Fixed Effects")  %>%
  add_header_above(c(" " = 1, "Model 1" = 3, "Model 2" = 3, 
                     "Model 3" = 3, "Model 4" = 3)) %>%
  kable_classic(html_font = "CMU Serif") %>% 
  kableExtra::landscape() %>% 
  kable_styling(latex_options = c("striped", "scale_down"))


# Compare models
# anova(am1, am2, am3, am4)
```

```{r altmed-random-table, echo=FALSE}
# Random Effects Table 
# Drop fixed effects rows
df.1 <- df[c(17:22),]
# change column names
colnames(df.1) <- c(" ", "Model 1", "Model 2", "Model 3", "Model 4", "6", "7", "8", "9")
# Drop NA columns 
df.2 <- df.1 %>% 
  dplyr::select(-(6:13)) %>% 
  dplyr::select(-(1))

# Rename rows 
rownames(df.2) <- c("$\\sigma^2$", "$\\tau$00", "ICC", "N", "Observations", "Marginal R$^2$/Conditional R$^2$")

kable(df.2,
      booktabs = T,
      escape = F, 
      caption = "Alternative Medical Systems: Random Effects") %>%
   kable_classic(html_font = "CMU Serif") %>%
  kable_styling(latex_options = c("striped","scale_down")) %>% 
  footnote(general = c("$\\\\sigma^2$ = within subject variance. 
                       $\\\\tau$00 = between subject variance.
                       ICC = intraclass correlation coefficient"), 
           escape = F) 

```

Model 2 assesses the association between level of education and the use of alternative medical systems and reveals that the odds of using alternative medical systems tend to increase with education. The odds of using alternative medical systems are 62% lower for those with less than a high school diploma (OR = `r round(exp(summ.am2$coefficients[2]),3)`, 95% CI: `r round(exp(summ.am2$coefficients[2] - 1.96 * summ.am2$coefficients[7]),3)`, `r round(exp(summ.am2$coefficients[2] + 1.96 * summ.am2$coefficients[7]),3)`, *p* \< 0.001), 47% lower for those with a high school diploma or GED (OR = `r round(exp(summ.am2$coefficients[3]),3)`, 95% CI: `r round(exp(summ.am2$coefficients[3] - 1.96 * summ.am2$coefficients[8]),3)`, `r round(exp(summ.am2$coefficients[3] + 1.96 * summ.am2$coefficients[8]),3)`, *p* \< 0.001), and 28% lower for those with some college or an associate's degree (OR = `r round(exp(summ.am2$coefficients[4]),3)`, 95% CI: `r round(exp(summ.am2$coefficients[4] - 1.96 * summ.am2$coefficients[9]),3)`, `r round(exp(summ.am2$coefficients[4] + 1.96 * summ.am2$coefficients[9]),3)`, *p* \< 0.01) compared to those with a college degree. There is no significant difference in mind-body medicine use between those with a college degree and those with a graduate or professional degree. The marginal $R^2$ is 0.014 and the conditional $R^2$ is 0.462, indicating that education explains 1.4% of the variance in the model.

Model 3 includes both logged income and education as predictors of the use of alternative medical systems. When accounting for education, logged income does not significantly predict use of alternative medical systems. When accounting for the effect of logged income, the effects of education on the odds of using alternative medical systems are similar to Model 2. The marginal $R^2$ is 0.015 and the conditional $R^2$ is 0.460, indicating that education and income together explain approximately 1.5% of the variance in the model.

Model 4 adds control variables, and the effects of logged income and education remain consistent with the effects in Model 3. The marginal $R^2$ is 0.092 and the conditional $R^2$ is 0.468, indicating that education, income, and the controls explain 9.2% of the variance in the model.

The results of these models indicate that education significantly predicts the use of alternative medical systems, although education and demographic variables explain a much smaller amount of the variation than unmeasured differences between subjects. As shown in Table \ref{tab:altmed-random-table}, within-subject variance is 3.29 across all 4 models and between-subject variance ranges from 2.33 in Model 4 and 2.83 in Model 1. The ICC also ranges from 0.41-0.46, indicating that 41-46% of the variation in use of alternative medical systems occurs between-subjects. Thus, while education significantly predicts the use of alternative medical systems, most of the variation in the use of these CAMs is explained by differences between subjects that the data do not account for rather the by the predictors in the model.

#### Physical and nutritional approaches

```{r model-glmm-pna, echo = FALSE}
# baseline model 
baseline <- glmer(pna ~ 1 + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Model with income
pna1 <- glmer(pna ~ LogIncome + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Model with education
pna2 <- glmer(pna ~ Education.2 + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Model with income and education
pna3 <- glmer(pna ~ LogIncome + Education.2 + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

#sjPlot::plot_model(pna2, 
#                   show.values=TRUE, show.p=TRUE,
#                   title="Effect of Income and Education on Mind Body Medicine Use")


# Model with all covariates
pna4 <- glmer(pna ~ LogIncome + Education.2 + Age + Sex + Race + Marital + HlthInsurance + HLSelf + Spiritual + Religion + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Store results for reporting 
summ.pna1 <- summary(pna1)
summ.pna2 <- summary(pna2)
summ.pna3 <- summary(pna3)
summ.pna4 <- summary(pna4)

# Create table with results
pna.tab.model <- tab_model(pna1, pna2, pna3, pna4,
          show.intercept = FALSE,
          pred.labels = c("Logged Income", "Less than high school", "High School Diploma/GED",
                          "Some College/AA", "Graduate/Professional Degree", "Age", 
                          "Female", "Non-Hispanic Black", "Other Race", 
                          "Mexican American and Other Hispanic", "Married", "Insured",
                          "Health Locus of Control", "Spirituality", "Religiosity"),
          dv.labels= c("Model 1", "Model 2", "Model 3", "Model 4"),
          title = "Mixed Model: Physical and Nutritional Approaches")

 # first you parse the html, then you put it as table, taking the first
df <- data.frame(readHTMLTable(htmlParse(pna.tab.model))[1])
# first row as colnames
colnames(df) <- df[1,]
# remove the fake first row
df <- df[-1,]
```

Table \ref{tab:pna-fixed-table} shows results for the level 1 predictors and Table \ref{tab:pna-random-table} shows results for the level 2 predictor for all physical and nutritional approaches models. Model 1 assesses the relationship between logged income and the use of physical and nutritional approaches. Results show for each additional unit of logged income the odds of using physical and nutritional approaches increases by 5% (OR = `r round(exp(summ.pna1$coefficients[2]),3)`, 95% CI: `r round(exp(summ.pna1$coefficients[2] - 1.96 * summ.pna1$coefficients[4]),3)`, `r round(exp(summ.pna1$coefficients[2] + 1.96 * summ.pna1$coefficients[4]),3)`, *p* \< 0.05). The marginal $R^2$ is 0.001 and the conditional $R^2$ is 0.201, indicating that the variance explained in the model is due to the level 2 predictor rather than the level 1 predictors.

```{r pna-fixed-table, echo=FALSE}
# Fixed effects table
# Drop random effects rows
df.1 <- df[c(1:15),]
# Drop row names
rownames(df.1)<-NULL
kable(df.1,
      booktabs = T,
      caption = "Physical and Nutritional Approaches: Fixed Effects") %>%
  add_header_above(c(" " = 1, "Model 1" = 3, "Model 2" = 3, 
                     "Model 3" = 3, "Model 4" = 3)) %>%
  kable_classic(html_font = "CMU Serif") %>% 
  kableExtra::landscape() %>% 
  kable_styling(latex_options = c("striped", "scale_down"))
```

```{r pna-random-table, echo=FALSE}
# Random Effects Table 
# Drop fixed effects rows
df.1 <- df[c(17:22),]
# change column names
colnames(df.1) <- c(" ", "Model 1", "Model 2", "Model 3", "Model 4", "6", "7", "8", "9")
# Drop NA columns 
df.2 <- df.1 %>% 
  dplyr::select(-(6:13)) %>% 
  dplyr::select(-(1))

# Rename rows 
rownames(df.2) <- c("$\\sigma^2$", "$\\tau$00", "ICC", "N", "Observations", "Marginal R$^2$/Conditional R$^2$")

kable(df.2,
      booktabs = T,
      escape = F, 
      caption = "Physical and Nutritional Approaches: Random Effects") %>%
   kable_classic(html_font = "CMU Serif") %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>% 
  footnote(general = c("$\\\\sigma^2$ = within subject variance. 
                       $\\\\tau$00 = between subject variance.
                       ICC = intraclass correlation coefficient"), 
           escape = F) 


# Compare models
# anova(pna1, pna2, pna3, pna4)
```

Model 2 assesses the association between the use of physical and nutritional approaches and levels of education using college degree as a reference category. The odds of using physical and nutritional approaches increase with each level of education, so that the odds of using physical and nutritional approaches are 52% lower for those with less than a high school diploma (OR = `r round(exp(summ.pna2$coefficients[2]),3)`, 95% CI: `r round(exp(summ.pna2$coefficients[2] - 1.96 * summ.pna2$coefficients[7]),3)`, `r round(exp(summ.pna2$coefficients[2] + 1.96 * summ.pna2$coefficients[7]),3)`, *p* \< 0.001), 37% lower for those with a high school diploma or GED (OR = `r round(exp(summ.pna2$coefficients[3]),3)`, 95% CI: `r round(exp(summ.pna2$coefficients[3] - 1.96 * summ.pna2$coefficients[8]),3)`, `r round(exp(summ.pna2$coefficients[3] + 1.96 * summ.pna2$coefficients[8]),3)`, *p* \< 0.001), and 17% lower for those with some college or an associate's degree (OR = `r round(exp(summ.pna2$coefficients[4]),3)`, 95% CI: `r round(exp(summ.pna2$coefficients[4] - 1.96 * summ.pna2$coefficients[9]),3)`, `r round(exp(summ.pna2$coefficients[4] + 1.96 * summ.pna2$coefficients[9]),3)`, *p* \< 0.01) compared to those with a college degree. There is not a significant difference in use of physical and nutritional approaches between those with college degree and those with a graduate or professional degree. The marginal $R^2$ is 0.015 and the conditional $R^2$ is 0.205, indicating that education explains 1.5% of the variance in the model.

Model 3 includes both logged income and education as predictors of the use of physical and nutritional approaches. When accounting for education, logged income does not significantly predict the use of physical and nutritional approaches. The effects of education on the odds of using physical and nutritional approaches are similar to Model 2. The marginal $R^2$ is 0.015 and the conditional $R^2$ is 0.205, indicating that income and education together explain approximately 1.5% of the variance in the model.

Model 4 adds control variables. As in Model 3, logged income does not significantly predict the use of physical and nutritional approaches. The effects of education remain significant. The marginal $R^2$ is 0.065 and the conditional $R^2$ is 0.243, indicating that income, education, and the controls together explain 6.5% of the variance explained in the model.

As Table \ref{tab:pna-random-table} demonstrates, within-subject variance is 3.29 across all 4 models and between-subject variance ranges from 0.77 in Model 4 to 0.83 in Model 1. The ICC also ranges from 0.19-0.20, indicating that 19-20% of the variation in the use of physical and nutritional approaches occurs between-subjects. While income and education significantly predict the use of physical and nutritional approaches, most of the variation is explained by differences between-subjects that the data do not account for rather than the predictors in the model.

#### Manipulative

```{r model-glmm-manipulative, echo = FALSE}
# baseline model 
baseline <- glmer(manipulative ~ 1 + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Model with income
manipulative1 <- glmer(manipulative ~ LogIncome + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Model with education
manipulative2 <- glmer(manipulative ~ Education.2 + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Model with income and education
manipulative3 <- glmer(manipulative ~ LogIncome + Education.2 + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

#sjPlot::plot_model(manipulative2, 
#                   show.values=TRUE, show.p=TRUE,
#                   title="Effect of Income and Education on Mind Body Medicine Use")


# Model with all covariates
manipulative4 <- glmer(manipulative ~ LogIncome + Education.2 + Age + Sex + Race + Marital + HlthInsurance + HLSelf + Spiritual + Religion + (1|M2ID), data = waves123.long.complete, family = binomial, nAGQ = 0)

# Store results for reporting 
summ.manipulative1 <- summary(manipulative1)
summ.manipulative2 <- summary(manipulative2)
summ.manipulative3 <- summary(manipulative3)
summ.manipulative4 <- summary(manipulative4)

# Create table with results
manipulative.tab.model <- tab_model(manipulative1, manipulative2, manipulative3, manipulative4,
          show.intercept = FALSE,
          pred.labels = c("Logged Income", "Less than high school", "High School Diploma/GED",
                          "Some College/AA", "Graduate/Professional Degree", "Age", 
                          "Female", "Non-Hispanic Black", "Other Race", 
                          "Mexican American and Other Hispanic", "Married", "Insured",
                          "Health Locus of Control", "Spirituality", "Religiosity"),
          dv.labels= c("Model 1", "Model 2", "Model 3", "Model 4"),
          title = "Mixed Model: Manipulative Treatments")

 # first you parse the html, then you put it as table, taking the first
df <- data.frame(readHTMLTable(htmlParse(manipulative.tab.model))[1])
# first row as colnames
colnames(df) <- df[1,]
# remove the fake first row
df <- df[-1,]
```

Table \ref{tab:manipulative-fixed-table} shows results for level 1 predictors and Table \ref{tab:manipulative-random-table} shows results for all level 2 predictors for all manipulative treatment models. Model 1 reveals a positive relationship between logged income and the use of manipulative treatments. The odds of using manipulative treatments are 15% higher for each additional unit of logged income (OR = `r round(exp(summ.manipulative1$coefficients[2]),3)`, 95% CI: `r round(exp(summ.manipulative1$coefficients[2] - 1.96 * summ.manipulative1$coefficients[4]),3)`, `r round(exp(summ.manipulative1$coefficients[2] + 1.96 * summ.manipulative1$coefficients[4]),3)`, *p* \< 0.001). The marginal $R^2$ is 0.003 and the conditional $R^2$ is 0.426, indicating that the variance explained in the model is due to the level 2 rather than level 1 effects.

```{r manipulative-fixed-table, echo=FALSE}
# Fixed effects table
# Drop random effects rows
df.1 <- df[c(1:15),]
# Drop row names
rownames(df.1)<-NULL
kable(df.1,
      booktabs = T,
      caption = "Manipulative Treatments: Fixed Effects") %>%
  add_header_above(c(" " = 1, "Model 1" = 3, "Model 2" = 3, 
                     "Model 3" = 3, "Model 4" = 3)) %>%
  kable_classic(html_font = "CMU Serif") %>% 
  kableExtra::landscape() %>% 
  kable_styling(latex_options = c("striped", "scale_down"))
```

```{r manipulative-random-table, echo=FALSE}
# Random Effects Table 
# Drop fixed effects rows
df.1 <- df[c(17:22),]
# change column names
colnames(df.1) <- c(" ", "Model 1", "Model 2", "Model 3", "Model 4", "6", "7", "8", "9")
# Drop NA columns 
df.2 <- df.1 %>% 
  dplyr::select(-(6:13)) %>% 
  dplyr::select(-(1))

# Rename rows 
rownames(df.2) <- c("$\\sigma^2$", "$\\tau$00", "ICC", "N", "Observations", "Marginal R$^2$/Conditional R$^2$")

kable(df.2,
      booktabs = T,
      escape = F, 
      caption = "Manipulative Treatments: Random Effects") %>%
   kable_classic(html_font = "CMU Serif") %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>% 
  footnote(general = c("$\\\\sigma^2$ = within subject variance. 
                       $\\\\tau$00 = between subject variance.
                       ICC = intraclass correlation coefficient"), 
           escape = F) 


# Compare models
# anova(manipulative1, manipulative2, manipulative3, manipulative4)
```

Model 2 assesses the association between the use of manipulative treatments and levels of education using college degree as a reference category. The odds of using manipulative treatments are 50% lower for those with less than a high school diploma (OR = `r round(exp(summ.manipulative2$coefficients[2]),3)`, 95% CI: `r round(exp(summ.manipulative2$coefficients[2] - 1.96 * summ.manipulative2$coefficients[7]),3)`, `r round(exp(summ.manipulative2$coefficients[2] + 1.96 * summ.manipulative2$coefficients[7]),3)`, *p* \< 0.001) and 27% lower for those with a high school diploma or GED (OR = `r round(exp(summ.manipulative2$coefficients[3]),3)`, 95% CI: `r round(exp(summ.manipulative2$coefficients[3] - 1.96 * summ.manipulative2$coefficients[8]),3)`, `r round(exp(summ.manipulative2$coefficients[3] + 1.96 * summ.manipulative2$coefficients[8]),3)`, *p* \< 0.001) compared to those with a college degree. There is not a significant difference in use of physical and nutritional approaches between those with college degree and those with some college or an associate's degree or those with a graduate or professional degree. The marginal $R^2$ is 0.008 and the conditional $R^2$ is 0.429, education does not explain the variance in the model.

Model 3 includes both logged income and education as predictors of the use of manipulative treatments. When accounting for education, logged income significantly predicts the use of manipulative treatments. For each one unit increase in logged income the odds of using manipulative treatments increases by 10% (OR = `r round(exp(summ.manipulative3$coefficients[2]),3)`, 95% CI: `r round(exp(summ.manipulative3$coefficients[2] - 1.96 * summ.manipulative3$coefficients[8]),3)`, `r round(exp(summ.manipulative3$coefficients[2] + 1.96 * summ.manipulative3$coefficients[8]),3)` *p* \<0.010). When accounting for the effect of logged income, the effects of education on the odds of using manipulative treatments also change. The odds are 46% lower for those with less than a high school diploma (OR = `r round(exp(summ.manipulative3$coefficients[3]),3)`, 95% CI: `r round(exp(summ.manipulative3$coefficients[3] - 1.96 * summ.manipulative3$coefficients[9]),3)`, `r round(exp(summ.manipulative3$coefficients[3] + 1.96 * summ.manipulative3$coefficients[9]),3)`, *p* \< 0.001). However, there is no significant difference in use of manipulative treatments between those with a college degree and those with high school/GED, some college/AA, or those with a graduate or professional degree. The marginal $R^2$ is 0.009 and the conditional $R^2$ is 0.431, indicating that income and education do not explain the variance in the model.

Model 4 adds control variables. Income remains a significant predictor of manipulative treatment use. For each one unit increase in logged income the odds of using manipulative treatments increase by 16% (OR = `r round(exp(summ.manipulative4$coefficients[2]),3)`, 95% CI: `r round(exp(summ.manipulative4$coefficients[2] - 1.96 * summ.manipulative4$coefficients[18]),3)`, `r round(exp(summ.manipulative4$coefficients[2] + 1.96 * summ.manipulative4$coefficients[18]),3)`, *p* \< 0.001). The effects of education are similar to those in Model 3. The marginal $R^2$ is 0.034 and the conditional $R^2$ is 0.440, indicating that 3.4% of the variance explained in the model is due to income, education, and the controls.

As Table \ref{tab:manipulative-random-table} illustrates, within-in subject variance is 3.29 across all 4 models and between-subject variance ranges from 2.39 in Model 4 to 2.44 in Model 3. The ICC also ranges from 0.42-0.43, indicating that 42-43% of the variation in manipulative treatment use occurs between subjects.

Overall, the results in Tables \ref{tab:mindbody-fixed-table} - \ref{tab:manipulative-random-table} reveal that education has consistent, positive effects on the odds of using each type of CAM. While income has less consistent effects, the effect of education suggests a positive relationship between SES and CAM use. Next, I examine the effects of SES on the total number of CAMs that individuals used.

#### Total CAMs

```{r model-glmm-totcams, echo = FALSE}
# Model with income
tot1 <- glmer(totalCams ~ LogIncome + (1|M2ID), data = waves123.long.complete, family = poisson, nAGQ = 0)
# check_overdispersion(tot1)

# Model with education
tot2 <- glmer(totalCams ~ Education.2 + (1|M2ID), data = waves123.long.complete, family = poisson, nAGQ = 0)
# check_overdispersion(tot2)

# Model with income and education
tot3 <- glmer(totalCams ~ LogIncome + Education.2 + (1|M2ID), data = waves123.long.complete, family = poisson, nAGQ = 0)
# check_overdispersion(tot3)

# Model with all covariates
tot4 <- glmer(totalCams ~ LogIncome + Education.2 + Age + Sex + Race + Marital + HlthInsurance + HLSelf + Spiritual + Religion + (1|M2ID), data = waves123.long.complete, family = poisson, nAGQ = 0)

# Store results for reporting 
summ.tot1 <- summary(tot1)
summ.tot2 <- summary(tot2)
summ.tot3 <- summary(tot3)
summ.tot4 <- summary(tot4)

# Check for overdispersion 
# If the dispersion ratio is close to one, a Poisson model fits well to the data. Dispersion ratios larger than one indicate overdispersion, thus a negative binomial model or similar might fit better to the data. A p-value < .05 indicates overdispersion.

# check_overdispersion(tot4)

# Create table with results
tot.tab.model <- tab_model(tot1, tot2, tot3, tot4,
          show.intercept = FALSE,
          pred.labels = c("Logged Income", "Less than high school", "High School Diploma/GED",
                          "Some College/AA", "Graduate/Professional Degree", "Age", 
                          "Female", "Non-Hispanic Black", "Other Race", 
                          "Mexican American and Other Hispanic", "Married", "Insured",
                          "Health Locus of Control", "Spirituality", "Religiosity"),
          dv.labels= c("Model 1", "Model 2", "Model 3", "Model 4"),
          title = "Poisson GLMM: Total CAMs")

 # first you parse the html, then you put it as table, taking the first
df <- data.frame(readHTMLTable(htmlParse(tot.tab.model))[1])
# first row as colnames
colnames(df) <- df[1,]
# remove the fake first row
df <- df[-1,]
```

Table \ref{tab:total-fixed-table} shows results for the level 1 predictors for models of the total number of CAMs and Table \ref{tab:total-random-table} shows results for the level 2 predictors. Model 1 assesses the association between total CAMs used and income. In Model 1, the relationship between logged income and total CAMs used is statistically significant and positive (IRR = `r round(exp(summ.tot1$coefficients[2]),3)`, 95% CI: `r round(exp(summ.tot1$coefficients[2] - 1.96 * summ.tot1$coefficients[4]),3)`, `r round(exp(summ.tot1$coefficients[2] + 1.96 * summ.tot1$coefficients[4]),3)`, *p* \< 0.050). The marginal $R^2$ is 0.000 and the conditional $R^2$ is 0.559, indicating income does not explain the variance in the model.

```{r total-fixed-table, echo = FALSE}
# Fixed effects table
# Drop random effects rows
df.1 <- df[c(1:15),]
# Drop row names
rownames(df.1)<-NULL
kable(df.1,
      booktabs = T,
      caption = "Total CAMs: Fixed Effects") %>%
  add_header_above(c(" " = 1, "Model 1" = 3, "Model 2" = 3, 
                     "Model 3" = 3, "Model 4" = 3)) %>%
  kable_classic(html_font = "CMU Serif") %>% 
  kableExtra::landscape() %>% 
  kable_styling(latex_options = c("striped", "scale_down"))
```

```{r total-random-table, echo=FALSE}
# Random Effects Table 
# Drop fixed effects rows
df.1 <- df[c(17:22),]
# change column names
colnames(df.1) <- c(" ", "Model 1", "Model 2", "Model 3", "Model 4", "6", "7", "8", "9")
# Drop NA columns 
df.2 <- df.1 %>% 
  dplyr::select(-(6:13)) %>% 
  dplyr::select(-(1))

# Rename rows 
rownames(df.2) <- c("$\\sigma^2$", "$\\tau$00", "ICC", "N", "Observations", "Marginal R$^2$/Conditional R$^2$")

kable(df.2,
      booktabs = T,
      escape = F,
      caption = "Total CAMs: Random Effects") %>%
   kable_classic(html_font = "CMU Serif") %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>% 
  footnote(general = c("$\\\\sigma^2$ = within subject variance. 
                       $\\\\tau$00 = between subject variance.
                       ICC = intraclass correlation coefficient"), 
           escape = F) 

#anova(tot1, tot2, tot3,tot4)
```

Model 2 assesses the association between total CAMs used and levels of education. The expected count of total CAMs used increases with each level of education. The expected number of CAMs used decreases by 47% for those with less than a high school education(IRR = `r round(exp(summ.tot2$coefficients[2]),3)`, 95% CI: `r round(exp(summ.tot2$coefficients[2] - 1.96 * summ.tot2$coefficients[7]),3)`, `r round(exp(summ.tot2$coefficients[2] + 1.96 * summ.tot2$coefficients[7]),3)`, *p* \< 0.001), by 31% for those with a high school diploma or GED (OR = `r round(exp(summ.tot2$coefficients[3]),3)`, 95% CI: `r round(exp(summ.tot2$coefficients[3] - 1.96 * summ.tot2$coefficients[8]),3)`, `r round(exp(summ.tot2$coefficients[3] + 1.96 * summ.tot2$coefficients[8]),3)`, *p* \< 0.001), and by 13% for those with some college or an associate's degree (IRR = `r round(exp(summ.tot2$coefficients[4]),3)`, 95% CI: `r round(exp(summ.tot2$coefficients[4] - 1.96 * summ.tot2$coefficients[9]),3)`, `r round(exp(summ.tot2$coefficients[4] + 1.96 * summ.tot2$coefficients[9]),3)` *p* \< 0.001). The expected number of CAMs increases 9% for those with a graduate or professional degree (IRR = `r round(exp(summ.tot2$coefficients[5]),3)`, 95% CI: `r round(exp(summ.tot2$coefficients[5] - 1.96 * summ.tot2$coefficients[10]),3)`, `r round(exp(summ.tot2$coefficients[5] + 1.96 * summ.tot2$coefficients[10]),3)`, *p* \< 0.050) compared to those with a college degree. Within subject variance is 0.64 and between subject variance is 0.77. The ICC is 0.55, indicating that 55% of the variance in total CAMs used is between subjects. The marginal $R^2$ is 0.029 and the conditional $R^2$ is 0.557, indicating that 2.9% of the variance explained in the model is due to the fixed effects.

Model 3 includes both logged income and education as predictors of total CAMs used. When accounting for education, logged income does not significantly predict total CAMs used. The effects of education remain the same as in Model 2. The marginal $R^2$ in Model 3 is 0.029 and the conditional $R^2$ is 0.557, indicating that 2.9% of the variance explained in the model is due to the fixed effects.

Model 4 adds control variables, and the effect of income becomes significant. For each one unit increase in logged income the expected count of total CAMs used increases by 3% (IRR = `r round(exp(summ.tot4$coefficients[2]),3)`, 95% CI: `r round(exp(summ.tot4$coefficients[2] - 1.96 * summ.tot4$coefficients[18]),3)`, `r round(exp(summ.tot4$coefficients[2] + 1.96 * summ.tot4$coefficients[18]),3)`, *p* \< 0.010). The effects of education are significant and similar to those in Models 2 and 3. The marginal $R^2$ is 0.152 and the conditional $R^2$ is 0.554, indicating that 15.2% of the variance explained in the model is due to the level 1 predictors. Notably, the control variables make substantial contributions to the variance explained by the model. Again, control variables play a substantial role in explaining variation in the model. The expected number of CAMs use is negatively related to age, and positively related to health locus of control, spirituality, and religiosity. The expected count of total CAMs used is 41% higher for women, 18% lower for Blacks compared to Whites, 17% lower for other race compared to Whites, 10% lower for married respondents compared to not married, and 11% lower for the insured compared to the uninsured.

As shown in Table \ref{tab:total-random-table}, within-in subject variance is 3.29 across all 4 models and between-subject variance ranges from 0.58 in Model 4 to 0.82 in Model 1. The ICC also ranges from 0.47-0.56, indicating that 47-56% of the variation in total CAMs used occurs between subjects.

Results from the GLMM models provide weak support for *Hypothesis 1a* and strong support for *Hypothesis 1b*. These results support the predictions of HLT and Human Capital Theory. SES, particularly income, predicts greater overall use of CAM and use of more dimensions of AHL. Next, I examine whether greater use of CAMs among high SES groups contributes to socioeconomic health disparities.

### Mediation Models

In the following section, I will report results from several models testing whether CAM use mediates the relationship between SES and health as well as whether health mediates the relationship between SES and CAM use. I included controls as covariates of CAM and health variables in all path models, but excluded them from the path diagrams for ease of interpretation. I present standardized coefficients in the path diagrams (See Appendix A) and unstandardized coefficients in the text. Path diagrams present a visual representation of the causal paths tested in the model. Coefficients for count outcomes represent log of expected counts. Coefficients for binary outcomes represent log odds. Coefficients for ordinal outcomes represent the expected change in z-score for each one unit change in the predictor. Stata's gsem command does not provide standardized coefficients in output. Coefficients in the text are from GSEM analyses. First, I present model fit statistics from the SEM analyses. Second, I present results for the relationship between SES and CAMs (*Hypothesis 1*). Third, I present results for the relationship between SES and health (*Hypothesis 2*). Fourth, I present results for CAMs as a mediator of the SES-health relationship (*Hypothesis 3*). Finally, I present results for health as a mediator of the SES-CAM relationship (*Hypothesis 4*).

#### Model Fit

Tables presenting model fit statistics for all SEM models can be found in Appendix B. $\chi$^2^ tests were significant for all models, indicating poor model fit. However, this fit statistic is sensitive to sample size, and large sample sizes almost always produce a significant results @kyriazos2018.

Model fit for total CAM models was overall good. TLI was a bit below the cutoff in the chronic conditions model (CFI = 0.948; TLI = 0.872; RMSEA = 0.046; SRMR = 0.02). CFI and TLI were below acceptable cutoffs in the self-reported health model (CFI = 0.883; TLI = 0.716; RMSEA = 0.069; SRMR = 0.032). The diabetes (CFI = 0.972; TLI = 0.931; RMSEA = 0.033; SRMR = 0.014), chronic pain (CFI = 0.957; TLI = 0.895; RMSEA = 0.035; SRMR = 0.015), high blood pressure (CFI = 0.98; TLI = 0.951; RMSEA = 0.027; SRMR = 0.011) and depression/anxiety (CFI = 0.958; TLI = 0.897; RMSEA = 0.036; SRMR = 0.016) models presented no model fit concerns.

Most model fit statistics indicated good fit for the mind-body medicine models. The chronic pain (CFI = 0.961; TLI = 0.906; RMSEA = 0.03, SRMR = 0.014) and depression/anxiety (CFI = 0.96; TLI = 0.902; RMSEA = 0.033, SRMR = 0.015) models presented no problems. TLI was a bit below the cutoff in the chronic conditions model (CFI = 0.95; TLI = 0.877; RMSEA =0.043, SRMR = 0.02), and CFI and TLI were a bit below the cutoffs in the self-reported health model (CFI = 0.879; TLI = 0.705; RMSEA = 0.067, SRMR = 0.031).

Most model fit statistics indicated good model fit for the alternative medical system models. The high blood pressure (CFI = 0.981; TLI = 0.954; RMSEA = 0.02, SRMR = 0.01) and diabetes (CFI = 0.969; TLI = 0.924; RMSEA = 0.027; SRMR = 0.014) models presented no problems. TLI was below the cutoff for the chronic conditions (CFI = 0.937; TLI = 0.803;RMSEA = 0.049, SRMR = 0.023), chronic pain (CFI = 0.932; TLI = 0.833; RMSEA = 0.029; SRMR = 0.014), and depression/anxiety (CFI = 0.935; TLI = 0.841; RMSEA = 0.032, SRMR = 0.015) models. CFI and TLI were below the cutoff in the self-reported health model (CFI = 0.826; TLI = 0.577; RMSEA = 0.067, SRMR = 0.031).

While most fit statistics indicated good model fit for the physical and nutritional approaches models, multiple models had low CFI and TLI values. TLI was below the cutoff in the chronic conditions (CFI = 0.933; TLI = 0.792; RMSEA = 0.051, SRMR = 0.024), chronic pain (CFI = 0.933; TLI = 0.837; RMSEA = 0.029; SRMR = 0.014), and depression/anxiety models (CFI = 0.936; TLI = 0.844; RMSEA = 0.031, SRMR = 0.015). CFI and TLI were low in the self-reported health model (CFI = 0.827; TLI = 0.578; RMSEA = 0.066, SRMR = 0.031). All statistics indicated good fit in the high blood pressure model (CFI = 0.983; TLI = 0.958; RMSEA = 0.019; SRMR = 0.01).

At least one of the fit statistics was low for all of the manipulative treatment models. TLI was low in the chronic pain (CFI = 0.948; TLI = 0.873; RMSEA = 0.028; SRMR = 0.014), depression/anxiety (CFI = 0.947; TLI = 0.871; RMSEA = 0.031; SRMR = 0.015), and chronic conditions models (CFI = 0.941; TLI = 0.817; RMSEA = 0.049; SRMR = 0.023). CFI and TLI were low in the self-reported health model (CFI = 0.842; TLI = 0.614; RMSEA = 0.066; SRMR = 0.031). All other fit statistics indicated good model fit.

#### SES and CAMs

*Hypothesis 1* predicts that SES will be positively related to CAM use. Results do not support *Hypothesis 1a*. I find that income at wave 1 does not predict CAM use at waves 2 or 3. Results support *Hypothesis 1b*. Summarizing the effects of education on CAM use across models, I find that education at wave 1 is positively related to total CAMs (*b* = 0.139-0.145), use of mind-body medicines (*b* = 0.340), use of alternative medical systems (*b* = 0.275), and use of physical and nutritional approaches (*b* = 0.179) at waves 2. Similarly, education at wave 1 is positively related to total CAMs (*b* = 0.056-0.062), use of mind-body medicines (*b* = 0.227-0.230), use of alternative medical systems (*b* = 0.148-0.160), and use of physical and nutritional approaches (*b* = 0.140-0.162) at wave 3. These results provide support for Health Lifestyle Theory and Human Capital Theory.

#### SES and Health

*Hypothesis 2* predicts that SES is positively related to health. Results support *Hypothesis 2a*. Income at wave 1 is negatively related to chronic conditions at wave 2 (*b* = -0.064 - -0.070) and wave 3 (*b* = -0.059 - -0.062), and depression/anxiety at wave 2 (*b* = -0.148). Income at wave 1 is positively related to self-reported health at wave 2 (*b* = 0.105) and wave 3 (*b* = 0.171). Results support *Hypothesis 2b*. Education at wave 1 is negatively related to chronic conditions at wave 2 (*b* = -0.044 - -0.046) and wave 3 (*b* = -0.032 - -0.040), diabetes at wave 3 (*b* = -0.206 - -0.214), chronic pain at wave 2 (*b* = -0.096) and wave 3 (*b* = -0.091 - -0.095), high blood pressure at wave 2 (*b* = -0.115), and depression/anxiety at wave 3 (*b* = -0.115 - -0.123). Education is positively related to self-reported health at wave 2 (*b* = 0.176 - 0.177) and wave 3 (*b* = 0.068 - 0.073). SES is positively related to health as predicted by Fundamental Causes Theory.

#### CAM as a Mediator for the SES-Health Relationship

To test *Hypothesis 3*, I conducted SEM and GSEM models on the effects of SES in wave 1 on CAM use in wave 2 and health in wave 3. Results do not support *Hypothesis 3a*. CAM use does not mediate the relationship between income and health across all models. This is expected given the lack of association between income and CAM use reported in previous sections.

Mediation effects are indicated by a significant indirect effect coefficient. The indirect effect coefficient is the product of the path coefficients between the independent variable and mediator, and the mediator and outcome. Results from 6 models show a mediation effect for CAM use, though the effect of CAMs on health outcomes was not in the expected direction. For example, Total CAMs mediates the relationship between education and sum of chronic conditions. As stated in the previous sections, education has a negative effect on sum of chronic conditions at wave 3 and a positive effect on total CAMs used at wave 2. Total CAMs at wave 2 has a positive effect on chronic conditions, indicating that greater CAM use at wave 2 is associated with an increase in chronic conditions at wave 3 (GSEM: *b* = 0.031, 95% CI: 0.019, 0.043, p \< 0.001). This is in opposition with the Health Lifestyles Theory prediction that CAM use will be related to better health. GSEM results also show that the indirect (GSEM: *b* = 0.005, 95% CI: 0.003, 0.006 p \< 0.001) and total (GSEM: *b* = -0.034, 95% CI: -0.57, -0.011, p \< 0.010) effects of education on chronic conditions via total CAMs are significant. The overall effect of education on sum of chronic conditions is reduced via total CAMs effect on sum of chronic conditions.

Total CAMs mediate the relationship between education and chronic pain. As stated in previous sections, education is negatively related to chronic pain wave 3 and positively related to total CAMs used at wave 2. Total CAMs at wave 2 is positively related to chronic pain at wave 3 (GSEM: *b* = 0.058, 95% CI: 0.010, 0.106, p \< 0.050), indicating use of more CAMs at wave 2 is associated with higher odds of reporting chronic pain at wave 3. Note, however, that health lifestyles theory would predict a negative relationship between health behavior (CAM use) in wave 2 and health problems (chronic pain) in wave 3. Results from the GSEM show a significant positive indirect effect for education on chronic pain via total CAMs (*b* = 0.008, 95% CI: 0.001, 0.015, p \< 0.050). The total effect is not significant. The positive effect of total CAMs on chronic pain and negative effect of education on chronic pain appear to cancel each other out.

Total CAMs used mediates the relationship between education and depression/anxiety. The relationship between education and health and total CAMs is the same as observed in previous models, and similarly each additional CAM used at wave 2 is associated with an increase in the log odds of reporting depression/anxiety at wave 3. Total CAMs used at wave 2 has a positive effect on depression/anxiety at wave 3 (GSEM: b = 0.065, 95% CI: 0.008, 0.123, p \< 0.050). GSEM results show a significant positive indirect effect of education on depression and anxiety mediated by total CAMs used (b = 0.009, 95% CI: 0.009, 0.017, p \< 0.050) and a negative total effect (b = -0.114, 95% CI: -0.227, -0.001, p \< 0.050). While education has a small positive effect on depression/anxiety via total CAMs, the direct negative effect of education on depression/anxiety is stronger, producing an overall negative effect.

Use of alternative medical systems mediates the relationship between education and sum of chronic conditions, following the same pattern observed in previous models. Education is negatively related to sum of chronic conditions at wave 3, and positively related to use of alternative medical systems at wave 2. Use of alternative medical systems at wave 2 is positively associated with sum of chronic conditions at wave 3 (GSEM: b = 0.109, 95% CI: 0.035, 0.183, p \< 0.010). GSEM analyses show a significant, positive indirect effect for education on chronic conditions mediated by use of alternative medical systems (b = 0.030, 95% CI: 0.005, 0.055, p \< 0.050), though the total effect is not significant.

Use of physical and nutritional approaches mediates the relationship between education and sum of chronic conditions, as well as the relationship between education and chronic pain, following the same pattern observed in the models reviewed above. Education is negatively related to sum of chronic conditions and chronic pain at wave 3, and positively related to use of physical and nutritional approaches at wave 2. Use of physical and nutritional approaches has a positive effect on sum of chronic conditions at wave 3 (GSEM: *b* = 0.162, 95% CI: 0.111, 0.212, p \< 0.001) and education has a positive indirect effect on chronic conditions via use of physical and nutritional approaches (GSEM: *b* = 0.029, 95% CI: 0.013, 0.045, p \< 0.001). Results similarly show a positive relationship between using physical and nutritional approaches at wave 2 and reporting chronic pain at wave 3 (GSEM: b = 0.292, 95% CI: 0.093, 0.490, p \< 0.010) and a positive indirect effect of education on chronic pain mediated by use of physical and nutritional approaches (GSEM: b = 0.052, 95% CI: 0.009, 0.095, p \< 0.050). The total effects are not significant. The nonsignificant total effect indicates that the positive indirect effects and negative direct effects cancel each other out.

These results do not support *Hypothesis 3a* and contradict *Hypothesis 3b* generated by HLT. There are some mediating effects of CAMs, but these effects operate in the opposite direction HLT predicts. Education is positively related to better health and CAM use. People with higher levels of education are more likely to utilize CAMs. However, people who use CAMs are more likely to report poor health in the future. CAM use at wave 2 is consistently associated with worse health at wave 3. This pattern appears in models without a mediation effect as well. While previous research has found that people in poor health are more likely to use CAMs [@astin1998; @bell2006; @thorne2002], HLT predicts that higher CAM use among high SES groups will produce better health.

#### Health as a Mediator of the SES-CAM Relationship

Findings presented in the previous section indicate that CAMs are associated with worse health. It is possible that people are using CAMs to address health problems that they already have, rather than to prevent health problems or to improve well-being. Next, I review results for health as a mediator of the SES-CAM relationship. *Hypothesis 4* predicts that low SES will be associated with worse health at wave 2, and worse health at wave 2 will be associated with greater CAM use at wave 3. Results from 3 models found a mediation effect for health on the SES-CAM relationship indicated by a significant indirect effect.

Chronic conditions mediates the relationship between income and use of phsyical and nutritional approaches. As noted previously, income does not predict use of physical and nutritional approaches, and negatively predicts chronic conditions. Chronic conditions has a positive effect on use of physical and nutritional approaches at wave 3 (*b* = 0.162, 95% CI: 0.111, 0.212, *p* \< 0.001). GSEM analyses indicate a significant negative indirect effect (*b* = -0.007, 95% CI: -0.011, -0.002, p \< 0.010). The total effect is not significant. This indicates that there is a mediation effect of income on use of physical and nutritional approaches via income's effect on chronic conditions. Lower income is associated with more chronic conditions, and more chronic conditions is associated with greater odds of using physical and nutritional approaches. These results provide weak support for *Hypothesis 4a*.

Chronic conditions mediate the relationship between education and use of physical and nutritional approaches. As noted previously, education is negatively related to chronic conditions, and positively related to use of physical and nutritional approaches. Chronic conditions at wave 2 is positively related to use of physical and nutritional approaches at wave 3 (*b* = 0.093, 95% CI: 0.049, 0.138, *p* \< 0.001). Results show a negative indirect effect (*b* = -0.047, 95% CI: -0.008, -0.001, *p* \< 0.010) and a positive total effect (*b* = 0.154, 95% CI: 0.064, 0.243, *p* \< 0.001). This indicates that positive effect of education on physical and nutritional approaches is stronger than the negative indirect effect via chronic conditions. The indirect and direct effects indicate that chronic conditions may act as an alternative pathway to use of physical and nutritional approaches, which is higher among high SES respondents and among respondents with chronic conditions, who are less likely to be high SES. These results support *Hypothesis 4b*.

Self-reported health mediates the relationship between education and use of physical and nutritional approaches. AS noted previously, education is associated with higher odds of using physical and nutritional approaches, and better self-reported health. Self-reported health at wave 2 is negatively related to use of physical and nutritional approaches at wave 3 (*b* = -0.11, 95% CI: -0.226, -0.006, *p* \< 0.050), indicating that those with poor self-rated health are more likely to use physical and nutritional approaches at wave 3. There is a significant negative indirect effect of education on self-reported health via education's effect on use of physical and nutritional approaches (*b* = -0.022, 95% CI: -0.042, -0.002, p \< 0.050) and a positive total effect (*b* = 0.142, 95% CI: 0.052, 0.231, *p* \< 0.010). While education has a small negative effect on use of physical and nutritonal approaches via self-reported health, the total effect represents the strength of education's positive association with use of physical and nutritional approaches. These results support *Hypothesis 4b*.

These results suggest that poor health may be an alternative path to use of physical and nutritional approaches, which is higher among high SES respondents and among respondents with more chronic conditions and worse self-reported health, who are less likely to be high SES. These results indicate that people use CAMs to address poor health.

## Conclusion

This chapter examined the relationship between SES and CAM use, as well as possible mediation models for the relationships between SES, CAM use, and health. Results provided weak support for the *Hypothesis 1a* and strong support for *Hypothesis 1b*. The relationship between income and CAM use was inconsistent, most often becoming non-significant after accounting for the influence of education. SEM analyses did not find a relationship between income and future CAM use. Higher levels of education were consistently associated with greater levels of CAM use across analyses, specifically greater overall CAM use, use of mind-body medicines, use of alternative medical systems, and use of physical and nutritional approaches. These results support HLT, which predicts that high SES will be associated with healthier lifestyles. Individuals with similar socialization and life experiences related to their structural position develop similar internalized dispositions toward health and health behavior and are therefore likely to engage in health lifestyles characteristic of their status group. These results also speak to the significant role of education in producing health lifestyles. Human Capital Theory asserts that the human capital (skills and abilities) and psychosocial resources (learned effectiveness and sense of personal control) associated with higher education increase the effective agency of high SES individuals to construct healthy lifestyles.

Both income and education were related to better health outcomes across analyses, providing support for *Hypothesis 2* and confirming the well established SES gradient in health. This finding supports Fundamental Causes Theory, which states that SES is a fundamental social cause of health. Fundamental Causes Theory observes that the relationship between SES and health is positive and enduring. The flexible resources associated with high SES produce a multitude of mechanisms that favor better health. Therefore, Fundamental Causes Theory asserts that the association cannot be attributed to any one mechanism. While health behavior may represent one mechanism, the flexible resources associated with SES also operate at a contextual level, shaping access to living conditions that favor health through epi-genetic and stress pathways unrelated to an individual's effort to promote health.

Results from the mediation analysis do not support *Hypothesis 3*. CAM use does not mediate the SES-health relationship as predicted by HLT. It is possible that the SES-health relationship is mediated by other health lifestyles. Results support Fundamental Cause Theory, which states that SES is a fundamental cause of health. SES is associated with flexible resources that can be leveraged for better health, operating through individual and contextual level pathways.

Mediation analysis results revealed that CAM use is associated with worse health, contradicting the prediction of Health Lifestyle Theory that CAM use promotes health. This finding is in line with previous research that has shown that people with worse health tend to use CAM at higher rates. In combination with results providing evidence for health as a mediator of the SES-CAM relationship (which support *Hypothesis 4*), this indicates that people tend to use CAMs to address health problems rather than to prevent health problems. Analyses did not test the relationship between CAM use and current health. Instead, CAM use predicted worse future health. It is unlikely that CAMs produce negative health. Previous research does not support this. In the case of specific chronic conditions, this effect could be due to comorbid conditions not measured in the model. CAM use predicted higher odds of future depression/anxiety and chronic pain. It is possible that respondents were already using CAMs at wave 2 to address poor health, either unmeasured chronic conditions or unmeasured symptoms of chronic pain and depression/anxiety. In the case of sum of chronic conditions, it is possible that respondents were already experiencing and addressing symptoms for a condition that had not yet been diagnosed.

This points to an issue with the health measures in this study. Measures are limited to the presence or absence of an illness. While evidence from clinical trials do not suggest that CAMs "cure" or resolve chronic illness, evidence from trials and CAM user reports indicate that CAMs are useful tools for symptom management. I explore this point further in Chapter 5.
