---
title: "Factor Analysis"
author: "Hannah Andrews"
date: "4/30/2021"
output: html_document
---

Sample: There are 6157 complete cases on CAM variables at Wave 1. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(foreign)
library(expss)
library(Hmisc)
library(dplyr)
library(summarytools)
library(magrittr)
library(tidyverse)
library(mosaic)
library(lattice)
library(ggplot2)
library(scales)
library(ggthemes)
library(polycor)
library(psych)
library(REdaS)
library(haven)
library(labelled)
library(mice)
library(GPArotation)
library(foreach)
library(doParallel)
library(questionr)
library(gridExtra)
library(DAKS)
library(fdth)
library(lavaan)
library(validateR)
```



```{r, echo = FALSE, results = 'hide'}
# Read in MIDUS1 data and subset to complete cases on CAMs
# Read in stata file for MIDUS 1
path <- ("C:/Users/hanna/Documents/git/AHL/Stata/data-cleaning/MIDUS1.dta")
M1 <- read_dta(path)
acamsList <- c("acam1", "acam2", "acam3", "acam4", "acam5", "acam6", "acam7", 
               "acam8", "acam9", "acam10", "acam11", "acam12", "acam13", "acam14", "acam15")
acams <- M1[acamsList] # Subset MIDUS1 - only include CAMs.

# Remove attributes from acams data - created in Stata (labels and data notes). They're (were) producing an error with bart_spher.
acams <- remove_attributes(acams, "label")
acams <- remove_attributes(acams, "notes")

str(acams)

# Create column that is the sum of CAMs used 
totalCams <- rowSums(acams[,1:15])

acams[sapply(acams, is.numeric)] <- lapply(acams[sapply(acams, is.numeric)], as.factor) # convert all columns to factors

str(acams) 
head(acams)

# Rename columns 
acams <- dplyr::rename(acams, 
                          aAcupuncture = acam1, 
                          aBiofeedback = acam2,
                          aChiropractic = acam3,
                          aEnergyHeal = acam4,
                          aExerciseMove = acam5,
                          aHerbal = acam6,
                          aVitamins = acam7,
                          aHomeopathy = acam8,
                          aHypnosis = acam9,
                          aImageryTech = acam10,
                          aMassage = acam11,
                          aPrayer = acam12,
                          aRelaxMeditate = acam13,
                          aSpecialDiet = acam14,
                          aSpiritHeal = acam15)
head(acams)
# Keep only complete cases
acams=acams[complete.cases(acams),]
str(acams)
```

# Describe Data 
## Table: Frequencies and percentages on each variable 
```{r, echo = FALSE}
#apply(acams[, 1:15], 2 , questionr::freq)
freq_dframer <- function(data){
   quest_out   <- apply(data, 2, questionr::freq)
      templist <- list()
       for (i in 1:length(names(quest_out))) {
       temp <- data.frame(Item = rep(names(quest_out[i]), nrow(quest_out[[i]])),
                      Category = row.names(quest_out[[i]]),
                      quest_out[[i]])
       templist[[i]] <- temp
  }
      freqs <- dplyr::bind_rows(templist) 
      row.names(freqs) <- NULL
      names(freqs)[3:5] <- c("Frequency", "Percent", "Pcnt_of_nonMissing") 
      freqs$Item <- factor(freqs$Item, levels = unique(freqs$Item))
   freqs
}
freqout <- freq_dframer(acams)
freqout
```
## Visualization of frequencies
```{r, echo = FALSE}
# Relative frequencies 
ggplot(freqout, aes(x = Category, y = Frequency)) +
  geom_col() +
  facet_wrap(. ~ Item, nrow = 5 )
```
# Number of CAM treatments used in the past 12 months 
```{r, echo = FALSE}

# Convert column to a factor 
totalCams <- as.factor(totalCams)

# Relative frequencies 
prop.table(table(totalCams))

# Counts and percentages, relative and cumulative 
tb <- fdt_cat(totalCams)
tb

# create a histogram of total CAMs used 
hist(as.numeric(totalCams))
```

# Correlation matrix 
Create and plot the tetrachoric correlation matrix for all CAMs. 
```{r, echo = FALSE}
# Create correlation matrix 
het.mat <- hetcor(acams)$cor
cor.plot(het.mat, numbers=T, upper=FALSE, main = "Tetrachoric Correlations", show.legend = TRUE, xlas = 2)
```

Check that items are not perfectly correlated with each other. 
```{r, echo=FALSE}
# EFA following https://www.youtube.com/watch?v=C5RJvMaHJNo&ab_channel=StatisticsofDOOM
# Channel: Statistics of DOOM by Lecturer: Dr. Erin M. Buchanan

# Screen data 
# nrow(acams)
# n = 5326

# Check min and max scores by running summary 
# summary(acams)

# Additivity - make sure nothing is correlated at 1. 
# don't want any 1s on the off diagonal.
symnum(het.mat)
```
# Test for correlation adequacy 
I will test for correlation adequacy using Bartlett's Sphericity test. This test tests the hypothesis that correlations between variables are greater than would be expected by chance. The null hypothesis states that all off diagonal are 0. If the null hypothesis is rejected there is correlation adequacy. 
```{r, echo = FALSE}
# Correlation adequacy Bartlett's test
cortest.bartlett(het.mat, n = nrow(acams))
```
I reject the null hypothesis. The CAM items are adequately correlated. 

# Test for sampling adequacy 
I will test for sampling adequacy using the Kaiser-Meyer-Olkin (KMO) test.MSA refers to the overall measure of sampling adequacy. MSAi refer to the measure of sampling adequacy for each item. MSA is a measure of the proportion of variance among variables that might be common variance. The lower the proportion of variance that is common the more suited the data are for factor analysis.  

MSA cutoffs: >.9 marvelous, .8s meritorious, .7s middling, .6s mediocre, .5s miserable, less than .5 is unacceptable.
```{r, echo=FALSE}
KMO(het.mat)
```
Items that may be a concern with regard to sampling adequacy: prayer or other spiritual practices, relaxation or meditation, and spiritual healing. Overall MSA indicates sampling adequacy. 

# Determining the number of factors
First, I will run a parallel analysis. From RDocumentation: "``Parallel" analyis is a technique that compares the scree of factors of the observed data with that of a random data matrix of the same size as the original."  
```{r, echo = FALSE, message=FALSE, results='hide'}
# Convert acams to numerics for parallel analysis
acams[] <- lapply(acams, function(x) as.numeric(as.character(x)))
str(acams)
#acams2 <- acams + 1
#str(acams2)
```
```{r, echo = FALSE, message = FALSE}
#nofactors <- fa.parallel(acams, fm = "ml", fa = "fa", main = "Parallel Analysis Scree Plots", cor = "tet")
#nofactors
# Ran with both the acams data and the tetrachoric correlation matrix obtained from hetcor to be sure the analysis is the same. Eigenvalues are the same in both parallel analyses. 
nofactors <- fa.parallel(het.mat, n.obs = 6157, fm = "ml", fa = "fa", main = "Parallel Analysis Scree Plots- hetmat")
```
I received many warning messages stating "A cell entry of 0 was replaced with correct =  0.5.  Check your data!" This has to do with continuity when computing a tetrachoric correlation matrix. I added 1 to all values in the dataframe to check that the issue was not the 0/1 values. The results were the same. 

From Statistics of DOOM notes: 
i.	The dark line is set at one, which is part of the Kaiser criterion. This method is an older rule of thumb that is not well supported anymore.  You would look at the number of eigenvalues that are greater than 1 (or .70 in new literature).  This rule tends to overestimate the number of factors/components needed.
ii.	The red dotted line is the random data set used to test this analysis. Your data is randomly reordered to see how many factors are better than chance. 
iii.	The blue line and triangles are your eigenvalues from the real dataset. 
iv.	You want to look at where the blue and red lines cross. 

The parallel analysis suggests 6 factors. This is where the lines cross. Looking at the scree plot, none of the drop offs appear to be very large. Seems like there are maybe 2 factors. 

Note: Scree plots are a visual depiction of the eigenvalues. Look for the large drop off to figure out how many factors to use.

# Kaiser Criterion 
```{r}
# older kaiser criterion, number of eigenvalues greater than 1 
sum(nofactors$fa.values > 1.0)
# new kaiser criterion, number of eigenvalues greater than 0.7
sum(nofactors$fa.values > .7)
```
New kaiser criterion rule (eigenvalues greater than 0.7) suggests 2 factors. 

After readings Ayers and Kronenfeld (2010), I will test the old CAM domains posted by the National Center for Complementary and Alternative Medicine (NCCAM), now the National Center for Complementary and Integrative Medicine (NCCIM). Then I will test the domains found by Ayers and Kronenfeld (2010). Finally, I will test the NCCIM's current CAM domains before performing exploratory factor analysis on the MIDUS data. 

# Old NCCAM Domains 
NCCAM domains: 
- Mind-Body Treatments: hypnosis, meditation, guided imagery, relaxation, deep breathing, pray for self, others pray, prayer groups, healing rituals, biofeedback 
- Alternative Medical Systems: Acupuncture, ayurveda, folk medicine, naturopathy, homeopathy
- Manipulative Treatments: Chiropractic, massage
- Biologically based practices: herbs, vitamins. 
```{r, echo = FALSE}
# Specify the model
#NCCAM.model <- 'mindBody =~ aHypnosis + aRelaxMeditate + aImageryTech + aPrayer + aSpiritHeal + aBiofeedback
#                alternativeSystems =~ aAcupuncture + aHomeopathy + aExerciseMove 
#                manipulativeTreatments =~ aChiropractic + aMassage
#                biologicalBased =~ aVitamins + aHerbal + a SpecialDiet'

#acams[sapply(acams, is.numeric)] <- lapply(acams[sapply(acams, is.numeric)], as.factor) # convert all columns to factors


# Fit the model 
#fit <- cfa(NCCAM.model, data = acams, ordered = TRUE)

# display summary output
#summary(fit, fit.measures = TRUE)
```

# 1-6 Factor Models with all CAM variables 
All models run using oblique rotation (factors are allowed to correlate when rotated) and maximum likelihood estimation. 

## Factor loadings: 1 Factor Model with all CAM variables
```{r, echo = FALSE}
allvar_1 <- fa(r = het.mat, nfactors = 1, n.obs=nrow(acams), rotate = "oblimin", fm = "ml")
allvar_1$loadings
```
### Reliability: 1 Factor Model
```{r, echo = FALSE}
ML1 <- c(1:15)
allvar1_ML1 <- psych::alpha(acams[, ML1])
allvar1_ML1$total
```
Raw alpha is 0.695 - Acceptable? 

## Factor loadings: 2 Factor model with all CAM variables 
```{r, echo = FALSE}
allvar_2 <- fa(r = het.mat, nfactors = 2, n.obs=nrow(acams), rotate = "oblimin", fm = "ml")
allvar_2$loadings
```
All items only load onto one factor. 

Ml1 includes spiritual healing and prayer.
All other items load onto ML2. 

ML2 accounts for 40% of the variance. ML1 accounts for 12% of the variance. Cumulatively this factor model accounts for 52% of the variance. 

The factors are correlated (r = 0.51).

### Reliability: 2 Factor Model
```{r, echo = FALSE}
ML1 <- c(12, 15)
ML2 <- c(1:11, 13:14)
allvar2_ML1 <- psych::alpha(acams[, ML1])
allvar2_ML1$total
```
Raw alpha for ML1 (prayer and spiritual healing) is very low, 0.29
```{r, echo = FALSE}
allvar2_ML2 <- psych::alpha(acams[, ML2])
allvar2_ML2$total
```
Raw alpha for ML2 (all except prayer and spiritual healing) is acceptable(?), 0.68.

## Factor Loadings: 3 Factor model all CAM variables 

```{r, echo = FALSE}
allvar_3 <- fa(r = het.mat, nfactors = 3, n.obs=nrow(acams), rotate = "oblimin", fm = "ml")
allvar_3$loadings
```
Prayer loads onto two factors (ML1 and ML2). 
ML1: prayer, biofeedback, hypnosis, imagery techniques, meditation
ML2: prayer and spiritual healing 
ML3: special diet, acupuncture, chiropractice, energy healing, exercise/movement, herbal, vitamins, homeopathy, and massage

## Factor Loadings: 4 Factor model all CAM variables 

```{r, echo = FALSE}
allvar_4 <- fa(r = het.mat, nfactors = 4, n.obs=nrow(acams), rotate = "oblimin", fm = "ml")
allvar_4$loadings
```

Energy healing loads on three factors. 
Prayer loads on two factors. 
ML1: prayer, energy healing, spirit healing
ML2: energy healing, biofeedback, hypnosis, imagery techniques, meditation 
ML3: prayer 
ML4: Acupuncture, Chiropractic, Exercise, Herbal, Vitamins, Homeopathy, Massage, Special Diet

## Factor Loadings: 5 Factor model all CAM variables 

```{r, echo = FALSE}
allvar_5 <- fa(r = het.mat, nfactors = 5, n.obs=nrow(acams), rotate = "oblimin", fm = "ml")
allvar_5$loadings
```
Diet doesn't load on any factors 
Prayer loads on two factors. 
Energy healing loads on three factors. 

ML1: Prayer
ML2: Exercise and massage
ML3: Energy, Biofeedback, Hypnosis, ImageryTech, RelaxMeditate
ML4: prayer, energy, spiritual healing 
ML5: Acupuncture, Chiropractic, Herbal, Vitamins, Homeopathy

## Factor Loadings: 6 Factor model all CAM variables 
```{r, echo = FALSE}
allvar_6 <- fa(r = het.mat, nfactors = 6, n.obs=nrow(acams), rotate = "oblimin", fm = "ml")
allvar_6$loadings
```
Energy healing loads on 3 factors. 
Prayer and exercise are only items to load on their factors. 

## Diagrams of factor loadings 
```{r, echo = FALSE}
fa.diagram(allvar_2, cut = .299, sort = TRUE)
fa.diagram(allvar_3, cut = .299, sort = TRUE)
fa.diagram(allvar_4, cut = .299, sort = TRUE)
fa.diagram(allvar_5, cut = .299, sort = TRUE)
fa.diagram(allvar_6, cut = .299, sort = TRUE)
```

## Compare model fit for 1, 2, 3, 4, 5, 6 factor models including all variables 
```{r, echo=FALSE}
efa <- function(df,   #correlation matrix  
                k) {      #number of factors
  foreach(i=1:k, .packages="psych") %dopar% fa(df, 
                                               nfactors=i,
                                               n.obs=6157,
                                               rotate="oblimin",
                                               fm="ml"
  )
}

# Creates a table that compares the models. 
compare_models <- function(model) {
  table_efa <- data.frame(Model=0, TLI=0, BIC=0, RMSR=0, RMSEA=0, RMSEA_lower_bound=0, RMSEA_upper_bound=0, RMSEA_confidence=0, CFI=0)   #empty data.frame to prealocate memory. 
  for(i in 1:length(model)){
    table_efa [i,1] <- paste("Model", i)
    table_efa [i,2] <- model[[i]]$TLI
    table_efa [i,3] <- model[[i]]$BIC
    table_efa [i,4] <- model[[i]]$rms
    table_efa [i,5:8] <- model[[i]]$RMSEA
    table_efa [i, 9] <- 1 - ((model[[i]]$STATISTIC-model[[i]]$dof)/(model[[i]]$null.chisq-model[[i]]$null.dof))
  }
  return(table_efa)
}

test <- efa(het.mat, 6)
model_fit_all_vars <- compare_models(test)
rownames(model_fit_all_vars) <- c('1 Factor Mode', '2 Factor Model', '3 Factor Model', '4 Factor Model', '5 Factor Model', '6 Factor Model')
model_fit_all_vars
```


```{r, echo = FALSE, results = 'hide'}
# Get CFI
#finalmodel <- round1 
#1 - ((finalmodel$STATISTIC-finalmodel$dof)/(finalmodel$null.chisq-finalmodel$null.dof))
```
# Different rotations, all CAMs

## 1 Factor model, Varimax rotation 
```{r, echo = FALSE}
allvar_varimax1 <- fa(r = het.mat, nfactors = 1, n.obs=nrow(acams), rotate = "varimax", fm = "ml")
allvar_varimax1$loadings
```
## Four factor, varimax rotation
```{r, echo = FALSE}
allvar_varimax4 <- fa(r = het.mat, nfactors = 4, n.obs=nrow(acams), rotate = "varimax", fm = "ml")
allvar_varimax4$loadings
```
## Four factor, promax rotation
```{r, echo = FALSE}
allvar_promax4 <- fa(r = het.mat, nfactors = 4, n.obs=nrow(acams), rotate = "promax", fm = "ml")
allvar_promax4$loadings
```

## Three factor, promax rotation 
```{r, echo = FALSE}
allvar_promax3 <- fa(r = het.mat, nfactors = 3, n.obs=nrow(acams), rotate = "promax", fm = "ml")
allvar_promax3$loadings
allvar_promax3
```
# Principal Axis Factoring Method 
## Four factor with promax (oblique rotation) 
```{r, echo = FALSE}
allvar_PApromax4 <- fa(r = het.mat, nfactors = 4, n.obs=nrow(acams), rotate = "promax", fm = "pa")
allvar_PApromax4
```
## Four factor with oblimin (oblique rotation)
```{r, echo = FALSE}
allvar_PAoblimin4 <- fa(r = het.mat, nfactors = 4, n.obs=nrow(acams), rotate = "oblimin", fm = "pa")
allvar_PAoblimin4
```

# Principal components factor analysis
## Four factor with varimax
```{r, echo = false}
allvar_PCApromax4 <- principal(r = het.mat, nfactors = 4, rotate="varimax",n.obs=nrow(acams))
allvar_PCApromax4
```



# Run factor model without prayer and spiritual healing 
```{r, echo = FALSE, results = 'hide'}
# Remove aPrayer and aSpiritHeal from acams
acams13 <- subset(acams, select = -c(aPrayer, aSpiritHeal))
str(acams13)
# convert all columns to factors. This is important for the tetrachoric correlations. They are much smaller when run as numeric. 
# Checked against tetrachoric matrix produced by Stata; they are the same when hetcor is run on factors in R.
# Will have to convert back to numeric for parallel analysis. 
acams13[sapply(acams13, is.numeric)] <- lapply(acams13[sapply(acams13, is.numeric)], as.factor)
str(acams13)
# Create correlation matrix 
het.mat13 <- hetcor(acams13)$cor
```
## Tetrachoric correlations (excluding prayer and spiritual healing)
```{r, echo = FALSE}
cor.plot(het.mat13, numbers=T, upper=FALSE, main = "Tetrachoric Correlations excluding prayer and spiritual healing ", show.legend = TRUE, xlas = 2)
```

## Test for correlation adequacy 

Bartlett's test for sphericity. 
```{r, echo = FALSE}
# Correlation adequacy Bartlett's test
cortest.bartlett(het.mat13, n = nrow(acams))
```
Items are adequately correlated. 

## Test for sampling adequacy 

MSA cutoffs: >.9 marvelous, .8s meritorious, .7s middling, .6s mediocre, .5s miserable, less than .5 is unacceptable.
```{r, echo=FALSE}
KMO(het.mat13)
```
All variables meet requirements for sampling adequacy. 

## Determining number of factors 

### Parallel Analysis 
```{r, echo = FALSE, message=FALSE, results='hide'}
# Convert acams13 to numeric for parallel analysis
acams13[] <- lapply(acams13, function(x) as.numeric(as.character(x)))
str(acams13)
#acams2 <- acams + 1
#str(acams2)
```
```{r, echo = FALSE, message = FALSE}
nofactors13 <- fa.parallel(acams13, fm = "ml", fa = "fa", main = "Parallel Analysis Scree Plots excluding prayer and spiritual healing", cor = "tet")
```
The parallel analysis suggests 5 factors. There is a drop off on the scree plot after 4. 

### Kaiser Criterion 
```{r}
# older kaiser criterion, number of eigenvalues greater than 1 
sum(nofactors13$fa.values > 1.0)
# new kaiser criterion, number of eigenvalues greater than 0.7
sum(nofactors13$fa.values > .7)
```
The Kaiser criterion suggests there is only 1 factor. The Kaiser criterion tends to over estimate factors. Doesn't seem good that it's only estimating 1. 

## Factor Loadings: 1 factor model excluding prayer and spiritual healing 
```{r, echo = FALSE}
exprayheal_1 <- fa(r = het.mat13, nfactors = 1, n.obs=6157, rotate = "oblimin", fm = "ml")
exprayheal_1$loadings
```
### Reliability: 1 Factor Model excluding prayer and spiritual healing 
```{r, echo = FALSE}
ML1 <- c(1:13)
exprayheal1_ML1 <- psych::alpha(acams13[, ML1])
exprayheal1_ML1$total
```
Removing prayer and spiritual healing does not improve the reliability at all. Raw alpha is almost exactly the same, 0.68. Lowers reliability slightly (from 0.695). 

## Factor Loadings: 3 factor model excluding prayer and spiritual healing 
```{r, echo = FALSE}
exprayheal_3 <- fa(r = het.mat13, nfactors = 3, n.obs=6157, rotate = "oblimin", fm = "ml")
exprayheal_3$loadings
```

Energy healing loads on two factors. 

### Reliability: 3 Factor Model excluding prayer and spiritual healing 
```{r, echo = FALSE}
ML3 <- c(1, 6:8, 13)
ML1 <- c(2, 9:10, 12)
ML2 <- c(3, 5, 11)
exprayheal3_ML3 <- psych::alpha(acams13[, ML3])
exprayheal3_ML3$total
```

```{r, echo = FALSE}
exprayheal3_ML1 <- psych::alpha(acams13[, ML1])
exprayheal3_ML1$total
```

```{r, echo = FALSE}
exprayheal3_ML2 <- psych::alpha(acams13[, ML2])
exprayheal3_ML2$total
```

## Factor Loadings: 4 factor model excluding prayer and spiritual healing 
```{r, echo = FALSE}
exprayheal_4 <- fa(r = het.mat13, nfactors = 4, n.obs=6157, rotate = "oblimin", fm = "ml")
exprayheal_4$loadings
```
Massage and energy heal load on two factors. 

## Factor loadings: 5 factor model excluding prayer and spiritual healing
```{r, echo = FALSE}
exprayheal_5 <- fa(r = het.mat13, nfactors = 5, n.obs=nrow(acams), rotate = "oblimin", fm = "ml")
exprayheal_5$loadings
```
Exercise/movement and energy healing load on two factors 

## Compare model fit for 3, 4, and 5 factor models excluding prayer and spiritual healing
```{r, echo=FALSE}
efa <- function(df,   #correlation matrix  
                k) {      #number of factors
  foreach(i=1:k, .packages="psych") %dopar% fa(df, 
                                               nfactors=i,
                                               n.obs=6157,
                                               rotate="oblimin",
                                               fm="ml"
  )
}

# Creates a table that compares the models. 
compare_models <- function(model) {
  table_efa <- data.frame(Model=0, TLI=0, BIC=0, RMSR=0, RMSEA=0, RMSEA_lower_bound=0, RMSEA_upper_bound=0, RMSEA_confidence=0, CFI=0)   #empty data.frame to prealocate memory. 
  for(i in 1:length(model)){
    table_efa [i,1] <- paste("Model", i)
    table_efa [i,2] <- model[[i]]$TLI
    table_efa [i,3] <- model[[i]]$BIC
    table_efa [i,4] <- model[[i]]$rms
    table_efa [i,5:8] <- model[[i]]$RMSEA
    table_efa [i, 9] <- 1 - ((model[[i]]$STATISTIC-model[[i]]$dof)/(model[[i]]$null.chisq-model[[i]]$null.dof))
  }
  return(table_efa)
}

test <- efa(het.mat13, 5)
model_fit_exprayheal <- compare_models(test)
rownames(model_fit_exprayheal) <- c('1 Factor Model', '2 Factor Model', '3 Factor Model', '4 Factor Model', '5 Factor Model')
model_fit_exprayheal
```
# Principal Axis Factoring excluding Prayer and Spiritual Healing
## Three factor model, oblimin rotation 
```{r, echo = FALSE}
exprayheal_PAoblimin3 <- fa(r = het.mat13, nfactors = 3, n.obs=nrow(acams), rotate = "oblimin", max.iter = 100, fm = "pa")
exprayheal_PAoblimin3
exprayheal_PAoblimin3$loadings
```
## Three factor model, promax rotation 
```{r, echo = FALSE}
exprayheal_PApromax3 <- fa(r = het.mat13, nfactors = 3, n.obs=nrow(acams), rotate = "promax", max.iter = 100, fm = "pa")
exprayheal_PApromax3
exprayheal_PApromax3$loadings
```
# Minimum residual factoring method excluding Prayer and Spiritual Healing
## Three factor model, oblimin rotation 
```{r, echo = FALSE}
exprayheal_MINRESoblimin3 <- fa(r = het.mat13, nfactors = 3, n.obs=nrow(acams), rotate = "oblimin", max.iter = 100, fm = "minres")
exprayheal_MINRESoblimin3
exprayheal_MINRESoblimin3$loadings
```
## Three factor model, promax rotation 
```{r, echo = FALSE}
exprayheal_MINRESpromax3 <- fa(r = het.mat13, nfactors = 3, n.obs=nrow(acams), rotate = "promax", max.iter = 100, fm = "minres")
exprayheal_MINRESpromax3
exprayheal_MINRESpromax3$loadings
```

## Three factor model, promax rotation 
```{r, echo = FALSE}
exprayheal_MINRESpromax3 <- fa(r = het.mat13, nfactors = 3, n.obs=nrow(acams), rotate = "promax", max.iter = 100, fm = "minres")
exprayheal_MINRESpromax3
exprayheal_MINRESpromax3$loadings
```
# Exclude prayer, spiritual healing, and energy healing 
## Tetrachoric correlation matrix
```{r, echo = FALSE, results= 'hide'}
acams12 <- subset(acams, select = -c(aPrayer, aSpiritHeal, aEnergyHeal))
acams12
# Converts all columns to factor 
acams12[sapply(acams12, is.numeric)] <- lapply(acams12[sapply(acams12, is.numeric)], as.factor) # convert all columns to factors
# Create correlation matrix 
het.mat12 <- hetcor(acams12)$cor
```
```{r, echo = FALSE}
cor.plot(het.mat12, numbers=T, upper=FALSE, main = "Tetrachoric Correlations excluding pray, spirit, energy", show.legend = TRUE, xlas = 2)
```
## Bartlett's Test of Spherecity
I will test for correlation adequacy using Bartlett's Sphericity test. This test tests the hypothesis that correlations between variables are greater than would be expected by chance. The null hypothesis states that all off diagonal are 0. If the null hypothesis is rejected there is correlation adequacy. 
```{r, echo = FALSE}
# Correlation adequacy Bartlett's test
cortest.bartlett(het.mat12, n = nrow(acams))
```
## KMO Test for sampling adequacy 
I will test for sampling adequacy using the Kaiser-Meyer-Olkin (KMO) test.MSA refers to the overall measure of sampling adequacy. MSAi refer to the measure of sampling adequacy for each item. MSA is a measure of the proportion of variance among variables that might be common variance. The lower the proportion of variance that is common the more suited the data are for factor analysis.  

MSA cutoffs: >.9 marvelous, .8s meritorious, .7s middling, .6s mediocre, .5s miserable, less than .5 is unacceptable.
```{r, echo=FALSE}
KMO(het.mat12)
```
Great.  

## Parallel Analysis 
### Maximum Likelihood
First, I will run a parallel analysis. From RDocumentation: "``Parallel" analyis is a technique that compares the scree of factors of the observed data with that of a random data matrix of the same size as the original."  
```{r, echo = FALSE, message=FALSE, results='hide'}
# Convert acams to numerics for parallel analysis
#acams[] <- lapply(acams, function(x) as.numeric(as.character(x)))
#str(acams)
#acams2 <- acams + 1
#str(acams2)
```
```{r, echo = FALSE, message = FALSE}
#nofactors <- fa.parallel(acams, fm = "ml", fa = "fa", main = "Parallel Analysis Scree Plots", cor = "tet")
#nofactors
# Ran with both the acams data and the tetrachoric correlation matrix obtained from hetcor to be sure the analysis is the same. Eigenvalues are the same in both parallel analyses. 
acams12_parallelMl <- fa.parallel(het.mat12, n.obs = 6157, fm = "ml", fa = "fa", main = "Parallel Analysis Scree Plots- hetmat12, maximum likelihood")
```

### Principal axis 
```{r, echo = FALSE}
acams12_parallelPa <- fa.parallel(het.mat12, n.obs = 6157, fm = "pa", fa = "fa", main = "Parallel Analysis Scree Plots- hetmat12, principal axis")
```
### Minimum Residual
```{r, echo = FALSE}
acams12_parallelMinRes <- fa.parallel(het.mat12, n.obs = 6157, fm = "minres", fa = "fa", main = "Parallel Analysis Scree Plots- hetmat12, minimum residual")
```
From Statistics of DOOM notes: 
i.	The dark line is set at one, which is part of the Kaiser criterion. This method is an older rule of thumb that is not well supported anymore.  You would look at the number of eigenvalues that are greater than 1 (or .70 in new literature).  This rule tends to overestimate the number of factors/components needed.
ii.	The red dotted line is the random data set used to test this analysis. Your data is randomly reordered to see how many factors are better than chance. 
iii.	The blue line and triangles are your eigenvalues from the real dataset. 
iv.	You want to look at where the blue and red lines cross. 

The parallel analysis suggests 5 factors. This is where the lines cross. Looking at the scree plot, none of the drop offs appear to be very large. Seems like there are maybe 2 factors. 

Note: Scree plots are a visual depiction of the eigenvalues. Look for the large drop off to figure out how many factors to use.

## Kaiser Criterion 
Old Kaiser Criterion
```{r}
# older kaiser criterion, number of eigenvalues greater than 1 
sum(acams12_parallelMl$fa.values > 1.0)
```
New Kaiser Criterion
```{r, echo = FALSE}
# new kaiser criterion, number of eigenvalues greater than 0.7
sum(acams12_parallelMl$fa.values > .7)
```

## Create table for model comparisons of all models excluding prayer, spiritual healing, and energy 
Factor alphas input in order from factor loadings matrix. For example, if ML3 is the first column of factors presented in the factor loadings table when called, the alpha will be input under reliabilityF1. 
Reliability calculate using Kuder-Richardson Formula 20 (KR-20), a measure for testing the reliability of binary items. If an item loads equally on multiple factors, I include it in the reliability calculation for all factors it loads on. In notes, I refer to factors in order of reliability columns. For example, "Massage loads onto factors 1 and 2" refers to reliabilityF1 and reliabilityF2.  
```{r, echo = FALSE}
acams12_modelcomparison <- data.frame(Model = character(0), TLI = numeric(0), CFI = numeric(0), BIC = numeric(0), RMSEA = numeric(0), RMSEA_lower_bound = numeric(0), RMSEA_upper_bound = numeric(0), RMSEA_confidence = numeric(0), RMSR = numeric(0), reliabilityF1 = numeric(0), reliabilityF2 = numeric(0), reliabilityF3 = numeric(0), reliabilityF4 = numeric(0), reliabilityF5 = numeric(0), notes = character(0))   #empty data.frame to prealocate memory.
```

## Maximum likelihood, Three factor 
### Orthogonal rotation: Varimax
```{r, echo = FALSE}
# Run model 
acams12_MlVarimax3 <- fa(r = het.mat12, nfactors = 3, n.obs=nrow(acams), rotate = "varimax", max.iter = 100, fm = "ml")
# Input fit statistics for model comparisons table
acams12_modelcomparison [1,1] <- "Three Factor, Maximum Likelihood, Varimax"
acams12_modelcomparison [1,2] <- acams12_MlVarimax3$TLI
acams12_modelcomparison [1,3] <- 1 - ((acams12_MlVarimax3$STATISTIC-acams12_MlVarimax3$dof)/(acams12_MlVarimax3$null.chisq-acams12_MlVarimax3$null.dof))
acams12_modelcomparison [1,4] <- acams12_MlVarimax3$BIC
acams12_modelcomparison [1,5:8] <- acams12_MlVarimax3$RMSEA
acams12_modelcomparison [1,9] <- acams12_MlVarimax3$rms
# Review factors loadings 
acams12_MlVarimax3$loadings
fa.diagram(acams12_MlVarimax3, cut = .299, sort = TRUE)
```
Massage and special diet load equally onto multiple factors. When calculating alphas, I include them for all factors they load onto.
```{r, echo = FALSE, results='hide'}
acams12_modelcomparison [1, 15] <- "Massage loads on all three factors fairly equally. Special diet loads on factors 1 and 3."
```
```{r, echo = FALSE, results= 'hide'}
# Convert acams12 columns to numeric for reliability calculations 
acams12[] <- lapply(acams12, function(x) as.numeric(as.character(x)))
str(acams12)
```
```{r, echo = FALSE}
acams12_modelcomparison [1, 10] <- kr20(acams12[, c("aAcupuncture", "aChiropractic", "aHerbal", "aVitamins", "aHomeopathy", "aMassage", "aSpecialDiet")])
acams12_modelcomparison [1, 11] <- kr20(acams12[, c("aBiofeedback", "aHypnosis","aImageryTech", "aMassage", "aRelaxMeditate")])
acams12_modelcomparison [1, 12] <- kr20(acams12[, c("aExerciseMove", "aMassage", "aSpecialDiet")])
```
### Oblique rotation: promax 
```{r, echo = FALSE}
# Run model
acams12_MlPromax3 <- fa(r = het.mat12, nfactors = 3, n.obs=nrow(acams), rotate = "Promax", max.iter = 100, fm = "ml")
# Input fit statistics to model comparison table
acams12_modelcomparison [2,1] <- "Three Factor, Maximum Likelihood, Promax"
acams12_modelcomparison [2,2] <- acams12_MlPromax3$TLI
acams12_modelcomparison [2,3] <- 1 - ((acams12_MlPromax3$STATISTIC-acams12_MlPromax3$dof)/(acams12_MlPromax3$null.chisq-acams12_MlPromax3$null.dof))
acams12_modelcomparison [2,4] <- acams12_MlPromax3$BIC
acams12_modelcomparison [2,5:8] <- acams12_MlPromax3$RMSEA
acams12_modelcomparison [2,9] <- acams12_MlPromax3$rms
# Review factor loadings
acams12_MlPromax3$loadings
fa.diagram(acams12_MlPromax3, cut = .299, sort = TRUE)
```
Special diet does not load on any factors. 
Massage loads equally on two factors.
```{r, echo = FALSE, results = 'hide'}
acams12_modelcomparison [2, 15] <- "Special diet does not load on any factors. Massage loads on factors 1 and 3."
```
```{r, echo = FALSE}
# Reliability
acams12_modelcomparison [2, 10] <- kr20(acams12[, c("aAcupuncture", "aChiropractic", "aHerbal", "aVitamins", "aHomeopathy", "aMassage")])
acams12_modelcomparison [2, 11] <- kr20(acams12[, c("aBiofeedback", "aHypnosis","aImageryTech", "aRelaxMeditate")])
acams12_modelcomparison [2, 12] <- kr20(acams12[, c("aExerciseMove", "aMassage")])
```

### Oblique rotation: oblimin
```{r, echo = FALSE}
# Run model 
acams12_Mloblimin3 <- fa(r = het.mat12, nfactors = 3, n.obs=nrow(acams), rotate = "oblimin", max.iter = 100, fm = "ml")

# Input fit statistics to model comparison table
acams12_modelcomparison [3,1] <- "Three Factor, Maximum Likelihood, oblimin"
acams12_modelcomparison [3,2] <- acams12_Mloblimin3$TLI
acams12_modelcomparison [3,3] <- 1 - ((acams12_Mloblimin3$STATISTIC-acams12_Mloblimin3$dof)/(acams12_Mloblimin3$null.chisq-acams12_Mloblimin3$null.dof))
acams12_modelcomparison [3,4] <- acams12_Mloblimin3$BIC
acams12_modelcomparison [3,5:8] <- acams12_Mloblimin3$RMSEA
acams12_modelcomparison [3,9] <- acams12_Mloblimin3$rms

# Review factor loadings
acams12_Mloblimin3$loadings
fa.diagram(acams12_Mloblimin3, cut = .299, sort = TRUE)
```
Massage loads fairly equally on factors 1 and 3. Special diet does not load on any factors. 
```{r, echo = FALSE, results = 'hide'}
acams12_modelcomparison [3, 15] <- "Massage loads fairly equally on factors 1 and 3. Special diet does not load on any factors."
```
```{r, echo = FALSE}
# Reliability
acams12_modelcomparison [3, 10] <- kr20(acams12[, c("aAcupuncture", "aChiropractic", "aHerbal", "aVitamins", "aHomeopathy", "aMassage")])
acams12_modelcomparison [3, 11] <- kr20(acams12[, c("aBiofeedback", "aHypnosis","aImageryTech", "aRelaxMeditate")])
acams12_modelcomparison [3, 12] <- kr20(acams12[, c("aExerciseMove", "aMassage")])
```
### Oblique rotation: cluster
```{r, echo = FALSE}
# Run model
acams12_Mlcluster3 <- fa(r = het.mat12, nfactors = 3, n.obs=nrow(acams), rotate = "cluster", max.iter = 100, fm = "ml")

# Input fit statistics to model comparison table
acams12_modelcomparison [4,1] <- "Three Factor, Maximum Likelihood, cluster"
acams12_modelcomparison [4,2] <- acams12_Mlcluster3$TLI
acams12_modelcomparison [4,3] <- 1 - ((acams12_Mlcluster3$STATISTIC-acams12_Mlcluster3$dof)/(acams12_Mlcluster3$null.chisq-acams12_Mlcluster3$null.dof))
acams12_modelcomparison [4,4] <- acams12_Mlcluster3$BIC
acams12_modelcomparison [4,5:8] <- acams12_Mlcluster3$RMSEA
acams12_modelcomparison [4,9] <- acams12_Mlcluster3$rms

# Review factor loadings
acams12_Mlcluster3$loadings
fa.diagram(acams12_Mlcluster3, cut = .299, sort = TRUE)
```
All items load on one factor. Special diet is questionable - less than .1 difference factor loading between factors 1 and 3. About .12 difference in loadings on factors 1 and 3 for massage.  

```{r, echo = FALSE, results = 'hide'}
acams12_modelcomparison [4, 15] <- "All items load on one factor. Special diet is questionable - less than .1 difference factor loading between factors 1 and 2. About .12 difference in loadings on factors 1 and 3 for massage."
```
```{r, echo = FALSE}
# Reliability
acams12_modelcomparison [4, 10] <- kr20(acams12[, c("aAcupuncture", "aChiropractic", "aHerbal", "aVitamins", "aHomeopathy")])
acams12_modelcomparison [4, 11] <- kr20(acams12[, c("aBiofeedback", "aHypnosis","aImageryTech", "aRelaxMeditate")])
acams12_modelcomparison [4, 12] <- kr20(acams12[, c("aExerciseMove", "aMassage", "aSpecialDiet")])
```
## Principal Axis, Three factor 
### Orthogonal rotation: varimax
```{r, echo = FALSE}
# Run model 
acams12_PAVarimax3 <- fa(r = het.mat12, nfactors = 3, n.obs=nrow(acams), rotate = "varimax", max.iter = 100, fm = "pa")

# Review factor loadings
acams12_PAVarimax3$loadings
fa.diagram(acams12_PAVarimax3, cut = .299, sort = TRUE)
```
Principal axis factoring produces an ultra-Heywood case. Warning message states factor scorings are probably incorrect. None of the principal axis solution are included in the table. 
### Oblique rotation: promax
```{r, echo = FALSE}
acams12_PApromax3 <- fa(r = het.mat12, nfactors = 3, n.obs=nrow(acams), rotate = "promax", max.iter = 100, fm = "pa")
acams12_PApromax3
acams12_PApromax3$loadings
```
### Oblique rotation: oblimin
```{r, echo = FALSE}
acams12_PAoblimin3 <- fa(r = het.mat12, nfactors = 3, n.obs=nrow(acams), rotate = "oblimin", max.iter = 100, fm = "pa")
acams12_PAoblimin3
acams12_PAoblimin3$loadings
```
### Oblique rotation: cluster 
```{r, echo = FALSE}
acams12_PAcluster3 <- fa(r = het.mat12, nfactors = 3, n.obs=nrow(acams), rotate = "cluster", max.iter = 100, fm = "pa")
acams12_PAcluster3
acams12_PAcluster3$loadings
```

## Minimum Residual, Three Factors
Minimum residual solutions produce error message warning about ultra-Heywood case. Not included in model comparisons table.
### Orthogonal rotation: varimax 
```{r, echo = FALSE}
# Run model
acams12_MRvarimax3 <- fa(r = het.mat12, nfactors = 3, n.obs=nrow(acams), rotate = "varimax", max.iter = 100, fm = "minres")

# Review factor loadings
acams12_MRvarimax3$loadings
fa.diagram(acams12_MRvarimax3, cut = .299, sort = TRUE)
```
### Oblique rotation: promax
```{r, echo = FALSE}
acams12_MRpromax3 <- fa(r = het.mat12, nfactors = 3, n.obs=nrow(acams), rotate = "Promax", max.iter = 100, fm = "minres")
acams12_MRpromax3
acams12_MRpromax3$loadings
```
### Oblique rotation: oblimin
```{r, echo = FALSE}
acams12_MRoblimin3 <- fa(r = het.mat12, nfactors = 3, n.obs=nrow(acams), rotate = "oblimin", max.iter = 100, fm = "minres")
acams12_MRoblimin3
acams12_MRoblimin3$loadings
```
### Oblique rotation: cluster 
```{r, echo = FALSE}
acams12_MRcluster3 <- fa(r = het.mat12, nfactors = 3, n.obs=nrow(acams), rotate = "cluster", max.iter = 100, fm = "minres")
acams12_MRcluster3
acams12_MRcluster$loadings
```

## Maximum Likelihood, Four Factor 
### Orthogonal rotation: Varimax
```{r, echo = FALSE}
# Run model 
acams12_MLvarimax4 <- fa(r = het.mat12, nfactors = 4, n.obs=nrow(acams), rotate = "varimax", max.iter = 100, fm = "ml")
# Input fit statistics for model comparisons table
acams12_modelcomparison [5,1] <- "Four Factor, Maximum Likelihood, Varimax"
acams12_modelcomparison [5,2] <- acams12_MLvarimax4$TLI
acams12_modelcomparison [5,3] <- 1 - ((acams12_MLvarimax4$STATISTIC-acams12_MLvarimax4$dof)/(acams12_MLvarimax4$null.chisq-acams12_MLvarimax4$null.dof))
acams12_modelcomparison [5,4] <- acams12_MLvarimax4$BIC
acams12_modelcomparison [5,5:8] <- acams12_MLvarimax4$RMSEA
acams12_modelcomparison [5,9] <- acams12_MLvarimax4$rms
# Review factors loadings 
acams12_MLvarimax4$loadings
fa.diagram(acams12_MLvarimax4, cut = .299, sort = TRUE)
```
Acupuncture loads on 1, 2, 4. Special diet loads on 2, 3. 
```{r, echo = FALSE, results = 'hide'}
acams12_modelcomparison [5, 15] <- "Acupuncture loads on 1, 2, 4. Special diet loads on 2, 3."
```
```{r, echo = FALSE}
# Reliability
acams12_modelcomparison [5, 10] <- kr20(acams12[, c("aAcupuncture", "aBiofeedback", "aImageryTech", "aHypnosis", "aRelaxMeditate")])
acams12_modelcomparison [5, 11] <- kr20(acams12[, c("aAcupuncture", "aHerbal", "aHomeopathy", "aSpecialDiet", "aVitamins")])
acams12_modelcomparison [5, 12] <- kr20(acams12[, c("aSpecialDiet", "aExerciseMove")])
acams12_modelcomparison [5, 13] <- kr20(acams12[, c("aMassage", "aAcupuncture", "aChiropractic")])
```
### Oblique rotation: promax
```{r, echo = FALSE}
# Run model 
acams12_MLpromax4 <- fa(r = het.mat12, nfactors = 4, n.obs=nrow(acams), rotate = "Promax", max.iter = 100, fm = "ml")
# Input fit statistics for model comparisons table
acams12_modelcomparison [6,1] <- "Four Factor, Maximum Likelihood, Promax"
acams12_modelcomparison [6,2] <- acams12_MLpromax4$TLI
acams12_modelcomparison [6,3] <- 1 - ((acams12_MLpromax4$STATISTIC-acams12_MLpromax4$dof)/(acams12_MLpromax4$null.chisq-acams12_MLpromax4$null.dof))
acams12_modelcomparison [6,4] <- acams12_MLpromax4$BIC
acams12_modelcomparison [6,5:8] <- acams12_MLpromax4$RMSEA
acams12_modelcomparison [6,9] <- acams12_MLpromax4$rms
# Review factors loadings 
acams12_MLpromax4$loadings
fa.diagram(acams12_MLpromax4, cut = .299, sort = TRUE)
```
```{r, echo = FALSE, results = 'hide'}
acams12_modelcomparison [6, 15] <- "Exercise only loads on one factor, factor 3"
```
```{r, echo = FALSE}
# Reliability
acams12_modelcomparison [6, 10] <- kr20(acams12[, c("aBiofeedback", "aHypnosis", "aImageryTech", "aRelaxMeditate")])
acams12_modelcomparison [6, 11] <- kr20(acams12[, c("aAcupuncture", "aHerbal", "aVitamins", "aHomeopathy", "aSpecialDiet")])
acams12_modelcomparison [6, 12] <- kr20(acams12[, c("aExerciseMove")])
acams12_modelcomparison [6, 13] <- kr20(acams12[, c("aChiropractic", "aMassage")])
```
### Oblique rotation: oblimin
```{r, echo = FALSE}
# Run model 
acams12_MLoblimin4 <- fa(r = het.mat12, nfactors = 4, n.obs=nrow(acams), rotate = "oblimin", max.iter = 100, fm = "ml")
# Input fit statistics for model comparisons table
acams12_modelcomparison [7,1] <- "Four Factor, Maximum Likelihood, Oblimin"
acams12_modelcomparison [7,2] <- acams12_MLoblimin4$TLI
acams12_modelcomparison [7,3] <- 1 - ((acams12_MLoblimin4$STATISTIC-acams12_MLoblimin4$dof)/(acams12_MLoblimin4$null.chisq-acams12_MLoblimin4$null.dof))
acams12_modelcomparison [7,4] <- acams12_MLoblimin4$BIC
acams12_modelcomparison [7,5:8] <- acams12_MLoblimin4$RMSEA
acams12_modelcomparison [7,9] <- acams12_MLoblimin4$rms
# Review factors loadings 
acams12_MLoblimin4$loadings
fa.diagram(acams12_MLoblimin4, cut = .299, sort = TRUE)
```
```{r, echo = FALSE, results = 'hide'}
acams12_modelcomparison [7, 15] <- "Exercise only loads on one factor, factor 3"
```
```{r, echo = FALSE}
# Reliability
acams12_modelcomparison [7, 10] <- kr20(acams12[, c("aAcupuncture", "aHerbal", "aVitamins", "aHomeopathy", "aSpecialDiet")])
acams12_modelcomparison [7, 11] <- kr20(acams12[, c("aBiofeedback", "aHypnosis", "aImageryTech", "aRelaxMeditate")])
acams12_modelcomparison [7, 12] <- kr20(acams12[, c("aExerciseMove")])
acams12_modelcomparison [7, 13] <- kr20(acams12[, c("aChiropractic", "aMassage")])
```
### Oblique rotation: cluster
```{r, echo = FALSE}
# Run model 
acams12_MLcluster4 <- fa(r = het.mat12, nfactors = 4, n.obs=nrow(acams), rotate = "cluster", max.iter = 100, fm = "ml")
# Input fit statistics for model comparisons table
acams12_modelcomparison [8,1] <- "Four Factor, Maximum Likelihood, Cluster"
acams12_modelcomparison [8,2] <- acams12_MLcluster4$TLI
acams12_modelcomparison [8,3] <- 1 - ((acams12_MLcluster4$STATISTIC-acams12_MLcluster4$dof)/(acams12_MLcluster4$null.chisq-acams12_MLcluster4$null.dof))
acams12_modelcomparison [8,4] <- acams12_MLcluster4$BIC
acams12_modelcomparison [8,5:8] <- acams12_MLcluster4$RMSEA
acams12_modelcomparison [8,9] <- acams12_MLcluster4$rms
# Review factors loadings 
acams12_MLcluster4$loadings
fa.diagram(acams12_MLcluster4, cut = .299, sort = TRUE)
```
```{r, echo = FALSE, results = 'hide'}
acams12_modelcomparison [8, 15] <- "Exercise only loads on one factor, factor 3"
```
```{r, echo = FALSE}
# Reliability
acams12_modelcomparison [8, 10] <- kr20(acams12[, c("aAcupuncture", "aHerbal", "aVitamins", "aHomeopathy", "aSpecialDiet")])
acams12_modelcomparison [8, 11] <- kr20(acams12[, c("aBiofeedback", "aHypnosis", "aImageryTech", "aRelaxMeditate")])
acams12_modelcomparison [8, 12] <- kr20(acams12[, c("aExerciseMove")])
acams12_modelcomparison [8, 13] <- kr20(acams12[, c("aChiropractic", "aMassage")])
```
# Four factor model dropping prayer, spiritual healing, biofeedback, and hypnosis. 
```{r, echo = FALSE, results = 'hide'}
# Remove aPrayer and aSpiritHeal from acams
acams11 <- subset(acams, select = -c(aPrayer, aSpiritHeal, aBiofeedback, aHypnosis))
acams11
# Create correlation matrix 
het.mat3 <- hetcor(acams11)$cor
```

```{r, echo = FALSE}
# Run four factor model without prayer, spiritual healing, biofeedback, or hypnosis.
round3 <- fa(r = het.mat3, nfactors = 4, n.obs=nrow(acams11), rotate = "oblimin", fm = "ml")
round3
round3$loadings
```

### Factor Loadings 
All items now load on a factor. All items only load onto one factor (great!). Energy healing is now loading onto the same factor as acupuncture, herbal therapy, vitamins, and homeopathy. I am not sure why an item would change factors after dropping items that did not load. 

### Model fit 

#### Goodness of fit statistics
TLI equals 0.985 - great! 
The comparitive fit index (CFI) equals
```{r, echo = FALSE}
# Get CFI
finalmodel <- round3 
1 - ((finalmodel$STATISTIC-finalmodel$dof)/(finalmodel$null.chisq-finalmodel$null.dof))
```

Also great!

#### Residual fit statistics 
(RMSR) is  0.01, indicating excellent fit. RMSEA is 0.019, indicating excellent fit. 

### Reliability Assessment 
```{r, echo = FALSE}
factor1 <- c(1, 3, 5:7)
factor2 <- c(8, 10)
factor3 <- c(2, 9)
factor4 <- c(4, 11)
```
```{r, echo = FALSE}
psych::alpha(acams11[ , factor1])
```
Raw alpha for factor 1 (acupuncture, energy healing, herbal, vitamins, and homeopathy) is 0.59 . 

```{r, echo = FALSE}
psych::alpha(acams11[ , factor2])
```
Raw alpha for factor 2 (imagery techniques and meditation) is 0.48 . 
```{r, echo = FALSE}
psych::alpha(acams11[ , factor3])
```
Raw alpha for factor 3 ( chiropractic and massage) is 0.40 . 

```{r, echo = FALSE }
psych::alpha(acams11[ , factor4])
```
Raw alpha for factor 4 (exercise and diet) is 0.40 .

 
