---
title: "Change Models"
author: "Hannah Andrews"
date: '2022-05-05'
output:
  pdf_document: default
  html_document: default
header-includes:
   - \usepackage{rotating, graphicx}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE}
library(lme4)
library(lattice)
library(sjstats)
library(jtools)
library(ROCR)
library(sjPlot)
library(performance)
library(XML)
library(kableExtra)
library(tidyverse)
library(caTools)
library(fastDummies)
library(semptools)
library(semoutput)
library(lavaan)
library(semPlot)
library(stargazer)
library(semTable)
setwd("C:/Users/hanna/Documents/git/AHL/R/")
load("waves123.long.rda")
load("waves123.long.complete.rda")
load("waves123.wide.rda")
load("waves12.wide.complete.rda")
load("waves23.wide.complete.rda")
load("waves13.wide.complete.rda")
setwd("C:/Users/hanna/Documents/git/AHL/R/Mixed model")
```

# Chron/Income Waves 1 and 2: Difference in cams predicted by difference in chronic conditions moderated by income

```{r models-waves12-chron-mod-income, echo = FALSE, results='hide'}

# Run diff cams model
waves12.change.diffcams <- lm(formula = DiffCams ~ DiffChron, data = waves12.wide.complete)

# Center diff chron 
waves12.wide.complete$DiffChronCenter <- scale(waves12.wide.complete$DiffChron, scale = FALSE)
# Center log income 
waves12.wide.complete$LogIncomeCenter <- scale(waves12.wide.complete$LogIncome_W1, scale = FALSE)
# Create interaction term 
waves12.wide.complete$DiffChronxLogIncome <- waves12.wide.complete$DiffChronCenter*waves12.wide.complete$LogIncomeCenter

# Run income model 
waves12.change.income <- lm(formula = DiffCams ~ DiffChron + LogIncome_W1 + DiffChronxLogIncome, data = waves12.wide.complete)

# Add Controls 
waves12.change.controls <- lm(formula = DiffCams ~ DiffChron + LogIncome_W1 + DiffChronxLogIncome + Age_W2 + Sex_W1 + Race_W2 + Marital_W2 + HlthInsurance_W2 + HLSelf_W2 + Spiritual_W2 + Religion_W2, data = waves12.wide.complete)
```

```{r waves12-chron-mod-income, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves12.change.diffcams, waves12.change.income, waves12.change.controls, 
          title = "Waves 1 and 2 Difference in CAMs Regressed on Difference in Chronic Conditions Moderated by Income", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in Chronic Conditions", "Logged Income", 
                               "Difference in Chronic Conditions x Logged Income",
                               "Age at Wave 2", "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE, 
          notes = c("White is the reference category for race.",
                    "Standard errors shown in parentheses."), 
          notes.align = "r") 

```

# Chron/Income Waves 2 and 3: Difference in cams predicted by difference in chronic conditions moderated by income

```{r models-waves23-chron-mod-income, echo = FALSE, results='hide'}

# Run diff cams model
waves23.change.diffcams <- lm(formula = DiffCams ~ DiffChron, data = waves23.wide.complete)

# Center diff chron 
waves23.wide.complete$DiffChronCenter <- scale(waves23.wide.complete$DiffChron, scale = FALSE)
# Center log income 
waves23.wide.complete$LogIncomeCenter <- scale(waves23.wide.complete$LogIncome_W2, scale = FALSE)
# Create interaction term 
waves23.wide.complete$DiffChronxLogIncome <- waves23.wide.complete$DiffChronCenter*waves23.wide.complete$LogIncomeCenter

# Run income model 
waves23.change.income <- lm(formula = DiffCams ~ DiffChron + LogIncome_W2 + DiffChronxLogIncome, data = waves23.wide.complete)

# Add Controls 
waves23.change.controls <- lm(formula = DiffCams ~ DiffChron + LogIncome_W2 + DiffChronxLogIncome + Age_W3 + Sex_W2 + Race_W2 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves23.wide.complete)
```

```{r waves23-chron-mod-income, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves23.change.diffcams, waves23.change.income, waves23.change.controls, 
          title = "Waves 2 and 3 Difference in CAMs Regressed on Difference in Chronic Conditions Moderated by Income", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in Chronic Conditions", "Logged Income", 
                               "Difference in Chronic Conditions x Logged Income",
                               "Age at Wave 3", "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("White is the reference category for race.", 
                    "Standard errors shown in parentheses."), 
          notes.align = "r") 

```

# Chron/Income Waves 1 and 3 Income as moderator 

```{r models-waves13-chron-mod-income, echo = FALSE, results='hide'}
# Run diff cams model
waves13.change.diffcams <- lm(formula = DiffCams ~ DiffChron, data = waves13.wide.complete)

# Center diff chron 
waves13.wide.complete$DiffChronCenter <- scale(waves13.wide.complete$DiffChron, scale = FALSE)
# Center log income 
waves13.wide.complete$LogIncomeCenter <- scale(waves13.wide.complete$LogIncome_W1, scale = FALSE)
# Create interaction term 
waves13.wide.complete$DiffChronxLogIncome <- waves13.wide.complete$DiffChronCenter*waves13.wide.complete$LogIncomeCenter

# Run income model 
waves13.change.income <- lm(formula = DiffCams ~ DiffChron + LogIncome_W1 + DiffChronxLogIncome, data = waves13.wide.complete)

# Add Controls
waves13.change.controls <- lm(formula = DiffCams ~ DiffChron + LogIncome_W1 + DiffChronxLogIncome + Age_W3 + Sex_W1 + Race_W3 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves13.wide.complete)
```

```{r waves13-chron-mod-income, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves13.change.diffcams, waves13.change.income, waves13.change.controls, 
          title = "Waves 1 and 3 Difference in CAMs Regressed on Difference in Chronic Conditions Moderated by Income", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in Chronic Conditions", "Logged Income", 
                               "Difference in Chronic Conditions x Logged Income",
                               "Age at Wave 3", "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("White is the reference category for race.",
                    "Standard errors shown in parentheses."),
          notes.align = "r")

```

# Chron/Educ Waves 1 and 2 Education as moderator

```{r models-waves12-chron-mod-educ, echo = FALSE, results='hide'}

# Run diff cams model
waves12.change.diffcams <- lm(formula = DiffCams ~ DiffChron, data = waves12.wide.complete)

# Run income model 
waves12.change.education <- lm(formula = DiffCams ~ DiffChron + Education_W1 + DiffChron*Education_W1, data = waves12.wide.complete)

# Add Controls
waves12.change.controls <- lm(formula = DiffCams ~ DiffChron + Education_W1 + DiffChron*Education_W1 + Age_W2 + Sex_W1 + Race_W2 + Marital_W2 + HlthInsurance_W2 + HLSelf_W2 + Spiritual_W2 + Religion_W2, data = waves12.wide.complete)
```

```{r waves12-chron-mod-educ, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves12.change.diffcams, waves12.change.education, waves12.change.controls, 
          title = "Waves 1 and 2 Difference in CAMs Regressed on Difference in Chronic Conditions Moderated by Education", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in Chronic Conditions", "HS Diploma/GED", 
                               "Some College/AA Degree",
                               "College Graduate", "Graduate or Professional Degree", "Age at Wave 2",
                               "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity", 
                               "Difference in Chronic Conditions x HS Diploma/GED", 
                               "Difference in Chronic Conditions x Some College/AA Degree", 
                               "Difference in Chronic Conditions x College Graduate", 
                               "Difference in Chronic Conditions x Professional or Graduate Degree"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("Less than HS is the reference category for education.",
                    "White is the reference category for race.",
                    "Standard errors shown in parentheses."), 
          notes.align = "r", 
          float.env = "sidewaystable")
```

# Chron/Educ Waves 2 and 3 Education as moderator

```{r models-waves23-chron-mod-educ, echo = FALSE, results='hide'}

# Run diff cams model
waves23.change.diffcams <- lm(formula = DiffCams ~ DiffChron, data = waves23.wide.complete)

# Run income model 
waves23.change.education <- lm(formula = DiffCams ~ DiffChron + Education_W2 + DiffChron*Education_W2, data = waves23.wide.complete)

# Add Controls
waves23.change.controls <- lm(formula = DiffCams ~ DiffChron + Education_W2 + DiffChron*Education_W2 + Age_W3 + Sex_W2 + Race_W2 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves23.wide.complete)
```

```{r waves23-chron-mod-educ, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves23.change.diffcams, waves23.change.education, waves23.change.controls, 
          title = "Waves 2 and 3 Difference in CAMs Regressed on Difference in Chronic Conditions Moderated by Education", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in Chronic Conditions", "HS Diploma/GED", 
                               "Some College/AA Degree",
                               "College Graduate", "Graduate or Professional Degree", "Age at Wave 3",
                               "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity", 
                               "Difference in Chronic Conditions x HS Diploma/GED", 
                               "Difference in Chronic Conditions x Some College/AA Degree", 
                               "Difference in Chronic Conditions x College Graduate", 
                               "Difference in Chronic Conditions x Professional or Graduate Degree"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("Less than HS is the reference category for education.",
                    "White is the reference category for race.", 
                    "Standard errors shown in parentheses."),
          notes.align = "r", 
          float.env = "sidewaystable")
```

# Chron/Educ Waves 1 and 3 Education as moderator 

```{r models-waves13-chron-mod-educ, echo = FALSE, results='hide'}

# Run diff cams model
waves13.change.diffcams <- lm(formula = DiffCams ~ DiffChron, data = waves13.wide.complete)

# Run income model 
waves13.change.education <- lm(formula = DiffCams ~ DiffChron + Education_W1 + DiffChron*Education_W1, data = waves13.wide.complete)

# Add Controls
waves13.change.controls <- lm(formula = DiffCams ~ DiffChron + Education_W1 + DiffChron*Education_W1 + Age_W3 + Sex_W1 + Race_W3 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves13.wide.complete)
```

```{r waves13-chron-mod-educ, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves13.change.diffcams, waves13.change.education, waves13.change.controls, 
          title = "Waves 1 and 3 Difference in CAMs Regressed on Difference in Chronic Conditions Moderated by Education", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in Chronic Conditions", "HS Diploma/GED", 
                               "Some College/AA Degree",
                               "College Graduate", "Graduate or Professional Degree", "Age at Wave 3",
                               "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity", 
                               "Difference in Chronic Conditions x HS Diploma/GED", 
                               "Difference in Chronic Conditions x Some College/AA Degree", 
                               "Difference in Chronic Conditions x College Graduate", 
                               "Difference in Chronic Conditions x Professional or Graduate Degree"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("Less than HS is the reference category for education.", 
                    "White is the reference category for race.", 
                    "Standard errors shown in parentheses."), 
          notes.align = "r", 
          float.env = "sidewaystable")
```

# SRH/Income Waves 1 and 2: Difference in cams predicted by difference in SRH moderated by income

```{r model-waves12-SRH-mod-income, echo = FALSE, results='hide'}

# Run diff cams model
waves12.change.diffcams <- lm(formula = DiffCams ~ DiffSRH, data = waves12.wide.complete)

# Center diff SRH 
waves12.wide.complete$DiffSRHCenter <- scale(waves12.wide.complete$DiffSRH, scale = FALSE)
# Create interaction term 
waves12.wide.complete$DiffSRHxLogIncome <- NA
waves12.wide.complete$DiffSRHxLogIncome <- waves12.wide.complete$DiffSRHCenter*waves12.wide.complete$LogIncomeCenter

# Run income model 
waves12.change.income <- lm(formula = DiffCams ~ DiffSRH + LogIncome_W1 + DiffSRHxLogIncome, data = waves12.wide.complete)

# Add controls 
waves12.change.controls <- lm(formula = DiffCams ~ DiffSRH + LogIncome_W1 + DiffSRHxLogIncome + Age_W2 + Sex_W1 + Race_W2 + Marital_W2 + HlthInsurance_W2 + HLSelf_W2 + Spiritual_W2 + Religion_W2, data = waves12.wide.complete)

```

```{r waves12-SRH-mod-income, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves12.change.diffcams, waves12.change.income, waves12.change.controls, 
          title = "Waves 1 and 2 Difference in CAMs Regressed on Difference in SRH Moderated by Income", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in SRH", "Logged Income", 
                               "Difference in SRH x Logged Income",
                               "Age at Wave 2", "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE, 
          notes = c("White is the reference category for race.", 
                    "Standard errors shown in parentheses.", 
                    "SRH: Self-Reported Health."), 
          notes.align = "r") 

```

# SRH/Income Waves 2 and 3: Difference in cams predicted by difference in chronic conditions moderated by income

```{r models-waves23-SRH-mod-income, echo = FALSE, results='hide'}

# Run diff cams model
waves23.change.diffcams <- lm(formula = DiffCams ~ DiffSRH, data = waves23.wide.complete)

# Center diff SRH 
waves23.wide.complete$DiffSRHCenter <- scale(waves23.wide.complete$DiffSRH, scale = FALSE)
# Create interaction term 
waves23.wide.complete$DiffSRHxLogIncome <- NA
waves23.wide.complete$DiffSRHxLogIncome <- waves23.wide.complete$DiffSRHCenter*waves23.wide.complete$LogIncomeCenter


# Run income model 
waves23.change.income <- lm(formula = DiffCams ~ DiffSRH + LogIncome_W2 + DiffSRHxLogIncome, data = waves23.wide.complete)

# Add Controls 
waves23.change.controls <- lm(formula = DiffCams ~ DiffSRH + LogIncome_W2 + DiffSRHxLogIncome + Age_W3 + Sex_W2 + Race_W2 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves23.wide.complete)
```

```{r waves23-SRH-mod-income, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves23.change.diffcams, waves23.change.income, waves23.change.controls, 
          title = "Waves 2 and 3 Difference in CAMs Regressed on Difference in SRH Moderated by Income", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in SRH", "Logged Income", 
                               "Difference in SRH x Logged Income",
                               "Age at Wave 3", "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE, 
          notes = c("White is the reference category for race.", 
                    "Standard errors shown in parentheses.", 
                    "SRH: Self-Reported Health."), 
          notes.align = "r") 

```

# SRH/Income Waves 1 and 3 Income as moderator 

```{r models-waves13-SRH-mod-income, echo = FALSE, results='hide'}
# Run diff cams model
waves13.change.diffcams <- lm(formula = DiffCams ~ DiffSRH, data = waves13.wide.complete)

# Center diff chron 
waves13.wide.complete$DiffSRHCenter <- scale(waves13.wide.complete$DiffSRH, scale = FALSE)
# Center log income 
waves13.wide.complete$LogIncomeCenter <- scale(waves13.wide.complete$LogIncome_W1, scale = FALSE)
# Create interaction term 
waves13.wide.complete$DiffSRHxLogIncome <- waves13.wide.complete$DiffSRHCenter*waves13.wide.complete$LogIncomeCenter

# Run income model 
waves13.change.income <- lm(formula = DiffCams ~ DiffSRH + LogIncome_W1 + DiffSRHxLogIncome, data = waves13.wide.complete)


# Add Controls
waves13.change.controls <- lm(formula = DiffCams ~ DiffSRH + LogIncome_W1 + DiffSRHxLogIncome + Age_W3 + Sex_W1 + Race_W3 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves13.wide.complete)
```

```{r waves13-SRH-mod-income, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves13.change.diffcams, waves13.change.income, waves13.change.controls, 
          title = "Waves 1 and 3 Difference in CAMs Regressed on Difference in SRH Moderated by Income", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in SRH", "Logged Income", 
                               "Difference in SRH x Logged Income",
                               "Age at Wave 3", "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("White is the reference category for race.", 
                    "Standard errors shown in parentheses.", 
                    "SRH: Self-Reported Health"), 
          notes.align = "r")

```

# SRH/Educ Waves 1 and 2 Education as moderator

```{r models-waves12-SRH-mod-educ, echo = FALSE, results='hide'}

# Run diff cams model
waves12.change.diffcams <- lm(formula = DiffCams ~ DiffSRH, data = waves12.wide.complete)

# Run income model 
waves12.change.education <- lm(formula = DiffCams ~ DiffSRH + Education_W1 + DiffChron*Education_W1, data = waves12.wide.complete)

# Add Controls
waves12.change.controls <- lm(formula = DiffCams ~ DiffSRH + Education_W1 + DiffChron*Education_W1 + Age_W2 + Sex_W1 + Race_W2 + Marital_W2 + HlthInsurance_W2 + HLSelf_W2 + Spiritual_W2 + Religion_W2, data = waves12.wide.complete)
```

```{r waves12-SRH-mod-educ, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves12.change.diffcams, waves12.change.education, waves12.change.controls, 
          title = "Waves 1 and 2 Difference in CAMs Regressed on Difference in SRH Moderated by Education", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in SRH", "HS Diploma/GED", 
                               "Some College/AA Degree",
                               "College Graduate", "Graduate or Professional Degree", "Age at Wave 2",
                               "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity", 
                               "Difference in SRH x HS Diploma/GED", 
                               "Difference in SRH x Some College/AA Degree", 
                               "Difference in SRH x College Graduate", 
                               "Difference in SRH x Professional or Graduate Degree"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("Less than HS is the reference category for education.", 
                    "White is the reference category for race.", 
                    "Standard errors shown in parentheses.", 
                    "SRH: Self-Reported Health."), 
          notes.align = "r", 
          float.env = "sidewaystable")
```

# SRH/Educ Waves 2 and 3 Education as moderator

```{r models-waves23-SRH-mod-educ, echo = FALSE, results='hide'}

# Run diff cams model
waves23.change.diffcams <- lm(formula = DiffCams ~ DiffSRH, data = waves23.wide.complete)

# Run income model 
waves23.change.education <- lm(formula = DiffCams ~ DiffSRH + Education_W2 + DiffSRH*Education_W2, data = waves23.wide.complete)

# Add Controls 
waves23.change.controls <- lm(formula = DiffCams ~ DiffSRH + Education_W2 + DiffSRH*Education_W2 + Age_W3 + Sex_W2 + Race_W2 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves23.wide.complete)

```

```{r waves23-SRH-mod-educ, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves23.change.diffcams, waves23.change.education, waves23.change.controls, 
          title = "Waves 2 and 3 Difference in CAMs Regressed on Difference in SRH Moderated by Education", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in SRH", "HS Diploma/GED", 
                               "Some College/AA Degree",
                               "College Graduate", "Graduate or Professional Degree", "Age at Wave 3",
                               "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity", 
                               "Difference in SRH x HS Diploma/GED", 
                               "Difference in SRH x Some College/AA Degree", 
                               "Difference in SRH x College Graduate", 
                               "Difference in SRH x Professional or Graduate Degree"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("Less than HS is the reference category for education.", 
                    "White is the reference category for race.", 
                    "Standard errors shown in parentheses.", 
                    "SRH: Self-Reported Health."), 
          notes.align = "r", 
          float.env = "sidewaystable")
```

# SRH/Educ Waves 1 and 3 Education as moderator 

```{r models-waves13-SRH-mod-educ, echo = FALSE, results='hide'}
# Run diff cams model
waves13.change.diffcams <- lm(formula = DiffCams ~ DiffSRH, data = waves13.wide.complete)

# Run income model 
waves13.change.education <- lm(formula = DiffCams ~ DiffSRH + Education_W1 + DiffChron*Education_W1, data = waves13.wide.complete)

# Add Controls 
waves13.change.controls <- lm(formula = DiffCams ~ DiffSRH + Education_W1 + DiffChron*Education_W1 + Age_W3 + Sex_W1 + Race_W3 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves13.wide.complete)
```

```{r waves13-SRH-mod-educ, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves13.change.diffcams, waves13.change.education, waves13.change.controls, 
          title = "Waves 1 and 3 Difference in CAMs Regressed on Difference in SRH Moderated by Education", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in SRH", "HS Diploma/GED", 
                               "Some College/AA Degree",
                               "College Graduate", "Graduate or Professional Degree", "Age at Wave 3",
                               "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity", 
                               "Difference in SRH x HS Diploma/GED", 
                               "Difference in SRH x Some College/AA Degree", 
                               "Difference in SRH x College Graduate", 
                               "Difference in SRH x Professional or Graduate Degree"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("Less than HS is the reference category for education.", 
                    "White is the reference category for race.", 
                    "Standard errors shown in parentheses.", 
                    "SRH: Self-Reported Health."), 
          notes.align = "r", 
          float.env = "sidewaystable")
```