---
title: "Ch. 4 Socioeconomic Status as Moderator of the CAM-Health Relationship"
output:
  pdf_document: default
  bookdown::pdf_document2: default
header-includes:
- \usepackage{float}
- \usepackage{sectsty}
- \usepackage{paralist}
- \usepackage{setspace}\spacing{2}
- \usepackage{fancyhdr}
- \usepackage{lastpage}
- \usepackage{dcolumn}
- \usepackage{rotating, graphicx, longtable}
- \extrafloats{200}
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

```{r, echo = FALSE}
library(lme4)
library(lattice)
library(sjstats)
library(jtools)
library(ROCR)
library(sjPlot)
library(performance)
library(XML)
library(kableExtra)
library(tidyverse)
library(caTools)
library(fastDummies)
library(semptools)
library(semoutput)
library(lavaan)
library(semPlot)
library(stargazer)
library(semTable)
library(rstatix)
library(gtsummary)
library(forcats)
setwd("C:/Users/hanna/Documents/git/AHL/R/")
load("waves123.long.rda")
load("waves123.long.complete.rda")
load("waves123.wide.rda")
load("waves12.wide.complete.rda")
load("waves23.wide.complete.rda")
load("waves13.wide.complete.rda")
setwd("C:/Users/hanna/Documents/git/AHL/R/Mixed model")
# Tutorials
# https://lmudge13.github.io/sample_code/mixed_effects.html
# Converting html to pdf https://github.com/gorkang/html2latex/
```

```{r wave-comp-educ-recodes, echo=FALSE}
waves12.wide.complete$Education_W1.2 <- factor(waves12.wide.complete$Education_W1, 
                                              levels = c("College Grad", 
                                                         "Less than High School", 
                                                         "HS Diploma/GED", 
                                                         "Some College/AA Degree", 
                                                         "Graduate or Professional Degree"))


waves23.wide.complete$Education_W2.2 <- factor(waves23.wide.complete$Education_W2, 
                                              levels = c("College Grad", 
                                                         "Less than High School", 
                                                         "HS Diploma/GED", 
                                                         "Some College/AA Degree", 
                                                         "Graduate or Professional Degree"))

waves13.wide.complete$Education_W1.2 <- factor(waves13.wide.complete$Education_W1, 
                                              levels = c("College Grad", 
                                                         "Less than High School", 
                                                         "HS Diploma/GED", 
                                                         "Some College/AA Degree", 
                                                         "Graduate or Professional Degree"))

```

```{r models-waves12-chron-mod-income, echo = FALSE, results='hide'}

# Center diff chron 
waves12.wide.complete$DiffChronCenter <- scale(waves12.wide.complete$DiffChron, scale = FALSE)
waves12.wide.complete$DiffChronCenter <- as.numeric(waves12.wide.complete$DiffChronCenter)
# Center log income 
waves12.wide.complete$LogIncomeCenter <- scale(waves12.wide.complete$LogIncome_W1, scale = FALSE)
waves12.wide.complete$LogIncomeCenter <- as.numeric(waves12.wide.complete$LogIncomeCenter)
# Create interaction term 
waves12.wide.complete$DiffChronxLogIncome <- waves12.wide.complete$DiffChronCenter*waves12.wide.complete$LogIncomeCenter

# Run income model 
waves12.chron.income.1 <- lm(formula = DiffCams ~ DiffChronCenter + LogIncomeCenter, data = waves12.wide.complete)

# Run income model 
waves12.chron.income.2 <- lm(formula = DiffCams ~ DiffChronCenter + LogIncomeCenter + DiffChronCenter*LogIncomeCenter, data = waves12.wide.complete)

# Add Controls 
waves12.chron.income.3 <- lm(formula = DiffCams ~ DiffChronCenter + LogIncomeCenter + DiffChronCenter*LogIncomeCenter + Education_W1.2 + Age_W2 + Sex_W1 + Race_W2 + Marital_W2 + HlthInsurance_W2 + HLSelf_W2 + Spiritual_W2 + Religion_W2, data = waves12.wide.complete)

# Extract standard errors
waves12.chron.income.1.se <- summary(waves12.chron.income.1)$coef[,"Std. Error"]
waves12.chron.income.2.se <- summary(waves12.chron.income.2)$coef[,"Std. Error"]
waves12.chron.income.3.se <- summary(waves12.chron.income.3)$coef[,"Std. Error"]
```

```{r models-waves23-chron-mod-income, echo = FALSE, results='hide'}

# Center diff chron 
waves23.wide.complete$DiffChronCenter <- scale(waves23.wide.complete$DiffChron, scale = FALSE)
waves23.wide.complete$DiffChronCenter <- as.numeric(waves23.wide.complete$DiffChronCenter)
# Center log income 
waves23.wide.complete$LogIncomeCenter <- scale(waves23.wide.complete$LogIncome_W2, scale = FALSE)
# Create interaction term 
waves23.wide.complete$DiffChronxLogIncome <- waves23.wide.complete$DiffChronCenter*waves23.wide.complete$LogIncomeCenter

# Run income model 
waves23.chron.income.1 <- lm(formula = DiffCams ~ DiffChronCenter + LogIncomeCenter, data = waves23.wide.complete)

# Run income model 
waves23.chron.income.2 <- lm(formula = DiffCams ~ DiffChronCenter + LogIncomeCenter + DiffChronCenter*LogIncomeCenter, data = waves23.wide.complete)

# Add Controls 
waves23.chron.income.3 <- lm(formula = DiffCams ~ DiffChronCenter + LogIncomeCenter + DiffChronCenter*LogIncomeCenter + Education_W2.2 + Age_W3 + Sex_W2 + Race_W2 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves23.wide.complete)

# Extract standard errors
waves23.chron.income.1.se <- summary(waves23.chron.income.1)$coef[,"Std. Error"]
waves23.chron.income.2.se <- summary(waves23.chron.income.2)$coef[,"Std. Error"]
waves23.chron.income.3.se <- summary(waves23.chron.income.3)$coef[,"Std. Error"]
```

```{r models-waves13-chron-mod-income, echo = FALSE, results='hide'}

# Center diff chron 
waves13.wide.complete$DiffChronCenter <- scale(waves13.wide.complete$DiffChron, scale = FALSE)
waves13.wide.complete$DiffChronCenter <- as.numeric(waves13.wide.complete$DiffChronCenter)
# Center log income 
waves13.wide.complete$LogIncomeCenter <- scale(waves13.wide.complete$LogIncome_W1, scale = FALSE)
# Create interaction term 
waves13.wide.complete$DiffChronxLogIncome <- waves13.wide.complete$DiffChronCenter*waves13.wide.complete$LogIncomeCenter

# Run income model 
waves13.chron.income.1 <- lm(formula = DiffCams ~ DiffChronCenter + LogIncomeCenter, data = waves13.wide.complete)

# Run income model 
waves13.chron.income.2 <- lm(formula = DiffCams ~ DiffChronCenter + LogIncomeCenter + DiffChronCenter*LogIncomeCenter, data = waves13.wide.complete)

# Add Controls
waves13.chron.income.3 <- lm(formula = DiffCams ~ DiffChronCenter + LogIncomeCenter + DiffChronCenter*LogIncomeCenter + Education_W1.2 + Age_W3 + Sex_W1 + Race_W3 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves13.wide.complete)

waves13.chron.income.1.se <- summary(waves13.chron.income.1)$coef[,"Std. Error"]
waves13.chron.income.2.se <- summary(waves13.chron.income.2)$coef[,"Std. Error"]
waves13.chron.income.3.se <- summary(waves13.chron.income.3)$coef[,"Std. Error"]
```

```{r models-waves12-chron-mod-educ, echo = FALSE, results='hide'}

# Run income model 
waves12.chron.educ.1 <- lm(formula = DiffCams ~ DiffChronCenter + Education_W1.2, data = waves12.wide.complete)

# Run income model 
waves12.chron.educ.2 <- lm(formula = DiffCams ~ DiffChronCenter + Education_W1.2 + DiffChronCenter*Education_W1.2, data = waves12.wide.complete)

# Add Controls
waves12.chron.educ.3 <- lm(formula = DiffCams ~ DiffChronCenter + Education_W1.2 + DiffChronCenter*Education_W1.2 + LogIncome_W1 + Age_W2 + Sex_W1 + Race_W2 + Marital_W2 + HlthInsurance_W2 + HLSelf_W2 + Spiritual_W2 + Religion_W2, data = waves12.wide.complete)

# Extract standard errors
waves12.chron.educ.1.se <- summary(waves12.chron.educ.1)$coef[,"Std. Error"]
waves12.chron.educ.2.se <- summary(waves12.chron.educ.2)$coef[,"Std. Error"]
waves12.chron.educ.3.se <- summary(waves12.chron.educ.3)$coef[,"Std. Error"]
```

```{r models-waves23-chron-mod-educ, echo = FALSE, results='hide'}

# Run income model 
waves23.chron.educ.1 <- lm(formula = DiffCams ~ DiffChronCenter + Education_W2.2, data = waves23.wide.complete)

# Run income model 
waves23.chron.educ.2 <- lm(formula = DiffCams ~ DiffChronCenter + Education_W2.2 + DiffChronCenter*Education_W2.2, data = waves23.wide.complete)

# Add Controls
waves23.chron.educ.3 <- lm(formula = DiffCams ~ DiffChronCenter + Education_W2.2 + DiffChronCenter*Education_W2.2 + LogIncome_W2 + Age_W3 + Sex_W2 + Race_W2 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves23.wide.complete)

# Extract standard errors
waves23.chron.educ.1.se <- summary(waves23.chron.educ.1)$coef[,"Std. Error"]
waves23.chron.educ.2.se <- summary(waves23.chron.educ.2)$coef[,"Std. Error"]
waves23.chron.educ.3.se <- summary(waves23.chron.educ.3)$coef[,"Std. Error"]
```

```{r models-waves13-chron-mod-educ, echo = FALSE, results='hide'}


# Run income model 
waves13.chron.educ.1 <- lm(formula = DiffCams ~ DiffChronCenter + Education_W1.2, data = waves13.wide.complete)

# Run income model 
waves13.chron.educ.2 <- lm(formula = DiffCams ~ DiffChronCenter + Education_W1.2 + DiffChronCenter*Education_W1.2, data = waves13.wide.complete)

# Add Controls
waves13.chron.educ.3 <- lm(formula = DiffCams ~ DiffChronCenter + Education_W1.2 + DiffChronCenter*Education_W1.2 + LogIncome_W1 + Age_W3 + Sex_W1 + Race_W3 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves13.wide.complete)

# Extract standard errors
waves13.chron.educ.1.se <- summary(waves13.chron.educ.1)$coef[,"Std. Error"]
waves13.chron.educ.2.se <- summary(waves13.chron.educ.2)$coef[,"Std. Error"]
waves13.chron.educ.3.se <- summary(waves13.chron.educ.3)$coef[,"Std. Error"]

```

```{r model-waves12-SRH-mod-income, echo = FALSE, results='hide'}

# Center diff SRH 
waves12.wide.complete$DiffSRHCenter <- scale(waves12.wide.complete$DiffSRH, scale = FALSE)
waves12.wide.complete$DiffSRHCenter <- as.numeric(waves12.wide.complete$DiffSRHCenter)
# Create interaction term 
waves12.wide.complete$DiffSRHxLogIncome <- NA
waves12.wide.complete$DiffSRHxLogIncome <- waves12.wide.complete$DiffSRHCenter*waves12.wide.complete$LogIncomeCenter

# Run income model 
waves12.SRH.income.1 <- lm(formula = DiffCams ~ DiffSRHCenter + LogIncomeCenter, data = waves12.wide.complete)

# Run income model 
waves12.SRH.income.2 <- lm(formula = DiffCams ~ DiffSRHCenter + LogIncomeCenter + DiffSRHCenter*LogIncomeCenter, data = waves12.wide.complete)

# Add controls 
waves12.SRH.income.3 <- lm(formula = DiffCams ~ DiffSRHCenter + LogIncomeCenter + DiffSRHCenter*LogIncomeCenter + Education_W1.2 + Age_W2 + Sex_W1 + Race_W2 + Marital_W2 + HlthInsurance_W2 + HLSelf_W2 + Spiritual_W2 + Religion_W2, data = waves12.wide.complete)

# Extract standard errors
waves12.SRH.income.1.se <- summary(waves12.SRH.income.1)$coef[,"Std. Error"]
waves12.SRH.income.2.se <- summary(waves12.SRH.income.2)$coef[,"Std. Error"]
waves12.SRH.income.3.se <- summary(waves12.SRH.income.3)$coef[,"Std. Error"]

```

```{r models-waves23-SRH-mod-income, echo = FALSE, results='hide'}

# Center diff SRH 
waves23.wide.complete$DiffSRHCenter <- scale(waves23.wide.complete$DiffSRH, scale = FALSE)
waves23.wide.complete$DiffSRHCenter <- as.numeric(waves23.wide.complete$DiffSRHCenter)
# Create interaction term 
# waves23.wide.complete$DiffSRHxLogIncome <- NA
waves23.wide.complete$DiffSRHxLogIncome <- waves23.wide.complete$DiffSRHCenter*waves23.wide.complete$LogIncomeCenter


# Run income model 
waves23.SRH.income.1 <- lm(formula = DiffCams ~ DiffSRHCenter + LogIncomeCenter, data = waves23.wide.complete)

# Run income model 
waves23.SRH.income.2 <- lm(formula = DiffCams ~ DiffSRHCenter + LogIncomeCenter + DiffSRHCenter*LogIncomeCenter, data = waves23.wide.complete)

# Add Controls 
waves23.SRH.income.3 <- lm(formula = DiffCams ~ DiffSRHCenter + LogIncomeCenter + DiffSRHCenter*LogIncomeCenter + Education_W2.2 + Age_W3 + Sex_W2 + Race_W2 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves23.wide.complete)

# Extract standard errors
waves23.SRH.income.1.se <- summary(waves23.SRH.income.1)$coef[,"Std. Error"]
waves23.SRH.income.2.se <- summary(waves23.SRH.income.2)$coef[,"Std. Error"]
waves23.SRH.income.3.se <- summary(waves23.SRH.income.3)$coef[,"Std. Error"]
```

```{r models-waves13-SRH-mod-income, echo = FALSE, results='hide'}

# Center diff chron 
waves13.wide.complete$DiffSRHCenter <- scale(waves13.wide.complete$DiffSRH, scale = FALSE)
waves13.wide.complete$DiffSRHCenter <- as.numeric(waves13.wide.complete$DiffSRHCenter)
# Center log income 
waves13.wide.complete$LogIncomeCenter <- scale(waves13.wide.complete$LogIncome_W1, scale = FALSE)
# Create interaction term 
waves13.wide.complete$DiffSRHxLogIncome <- waves13.wide.complete$DiffSRHCenter*waves13.wide.complete$LogIncomeCenter

# Run income model 
waves13.SRH.income.1 <- lm(formula = DiffCams ~ DiffSRHCenter + LogIncomeCenter, data = waves13.wide.complete)

# Run income model 
waves13.SRH.income.2 <- lm(formula = DiffCams ~ DiffSRHCenter + LogIncomeCenter + DiffSRHCenter*LogIncomeCenter, data = waves13.wide.complete)


# Add Controls
waves13.SRH.income.3 <- lm(formula = DiffCams ~ DiffSRHCenter + LogIncomeCenter + DiffSRHCenter*LogIncomeCenter + Education_W1.2 + Age_W3 + Sex_W1 + Race_W3 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves13.wide.complete)

# Extract standard errors
waves13.SRH.income.1.se <- summary(waves13.SRH.income.1)$coef[,"Std. Error"]
waves13.SRH.income.2.se <- summary(waves13.SRH.income.2)$coef[,"Std. Error"]
waves13.SRH.income.3.se <- summary(waves13.SRH.income.3)$coef[,"Std. Error"]
```

```{r models-waves12-SRH-mod-educ, echo = FALSE, results='hide'}

# Run income model 
waves12.SRH.educ.1 <- lm(formula = DiffCams ~ DiffSRHCenter + Education_W1.2, data = waves12.wide.complete)

# Run income model 
waves12.SRH.educ.2 <- lm(formula = DiffCams ~ DiffSRHCenter + Education_W1.2 + DiffSRHCenter*Education_W1.2, data = waves12.wide.complete)

# Add Controls
waves12.SRH.educ.3 <- lm(formula = DiffCams ~ DiffSRHCenter + Education_W1.2 + DiffSRHCenter*Education_W1.2 + LogIncome_W1 + Age_W2 + Sex_W1 + Race_W2 + Marital_W2 + HlthInsurance_W2 + HLSelf_W2 + Spiritual_W2 + Religion_W2, data = waves12.wide.complete)

# Extract standard errors
waves12.SRH.educ.1.se <- summary(waves12.SRH.educ.1)$coef[,"Std. Error"]
waves12.SRH.educ.2.se <- summary(waves12.SRH.educ.2)$coef[,"Std. Error"]
waves12.SRH.educ.3.se <- summary(waves12.SRH.educ.3)$coef[,"Std. Error"]
```

```{r models-waves23-SRH-mod-educ, echo = FALSE, results='hide'}

# Run income model 
waves23.SRH.educ.1 <- lm(formula = DiffCams ~ DiffSRHCenter + Education_W2.2, data = waves23.wide.complete)

# Run income model 
waves23.SRH.educ.2 <- lm(formula = DiffCams ~ DiffSRHCenter + Education_W2.2 + DiffSRHCenter*Education_W2.2, data = waves23.wide.complete)

# Add Controls 
waves23.SRH.educ.3 <- lm(formula = DiffCams ~ DiffSRHCenter + Education_W2.2 + DiffSRHCenter*Education_W2.2 + LogIncome_W2 + Age_W3 + Sex_W2 + Race_W2 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves23.wide.complete)

# Extract standard errors
waves23.SRH.educ.1.se <- summary(waves23.SRH.educ.1)$coef[,"Std. Error"]
waves23.SRH.educ.2.se <- summary(waves23.SRH.educ.2)$coef[,"Std. Error"]
waves23.SRH.educ.3.se <- summary(waves23.SRH.educ.3)$coef[,"Std. Error"]
```

```{r models-waves13-SRH-mod-educ, echo = FALSE, results='hide'}

# Run income model 
waves13.SRH.educ.1 <- lm(formula = DiffCams ~ DiffSRHCenter + Education_W1.2, data = waves13.wide.complete)

# Run income model 
waves13.SRH.educ.2 <- lm(formula = DiffCams ~ DiffSRHCenter + Education_W1.2 + DiffSRHCenter*Education_W1.2, data = waves13.wide.complete)

# Add Controls 
waves13.SRH.educ.3 <- lm(formula = DiffCams ~ DiffSRHCenter + Education_W1.2 + DiffSRHCenter*Education_W1.2 + LogIncome_W1 + Age_W3 + Sex_W1 + Race_W3 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3, data = waves13.wide.complete)

# Extract standard errors
waves13.SRH.educ.1.se <- summary(waves13.SRH.educ.1)$coef[,"Std. Error"]
waves13.SRH.educ.2.se <- summary(waves13.SRH.educ.2)$coef[,"Std. Error"]
waves13.SRH.educ.3.se <- summary(waves13.SRH.educ.3)$coef[,"Std. Error"]
```

# Questions and Hypotheses

In this chapter, I examine income and education as moderators of the relationship between changes in health and changes in CAM use. First, I examine whether change in health status, as measured by changes in chronic conditions and self-reported health, predicts a change in CAM use. I expect that individuals who experience a decline in health status are more likely to adopt new CAMs (*Hypothesis 5*). Second, I examine whether income and education moderate this relationship by influencing the likelihood of using CAMs in response to changes in health status. I expect that individuals with higher incomes are more likely to adopt new CAMs in response to a decline in health compared to those with lower incomes (*Hypothesis 6a*). I also expect that individuals with higher levels of education are more likely to adopt new CAMs in response to a decline in health compared to those with lower levels of education (*Hypothesis 6b*).

There is little existing research on SES as a moderator of the relationship between change in health and change in individual health behavior. Previous research has shown that education is an important moderator of behavior change following a new diagnosis (Margolis 2013) or a heart attack (Wray et al. 1998). There is also evidence that socioeconomic status moderates behavioral change in cancer survivors (Naik et al. 2016). A new diagnosis has been found to increase the likelihood of behavioral change (Clark and Etilé 2002; Falba 2005; Keenan 2009), although this research typically includes socioeconomic status as a control variable and does not consider socioeconomic status as a possible moderator of this relationship. To my knowledge, research has yet to examine socioeconomic status as a moderator of the relationship between change in health and change in health lifestyles, AHL or THL. Given that social position shapes the structural context within which indiviuals make health lifestyles choices (Cockerham 2005; Link and Phelan 2005), I expect that SES will condition change in AHL following a change in health.

In sum, I will be testing the following hypotheses: 

* *Hypothesis 5*  A decline in health status will result in higher CAM use.
* *Hypothesis 6*  SES will moderate the relationship between change in health status and change in health behavior.
  + *Hypothesis 6a* Those with higher incomes are more likely to adopt new CAMs in the event of a health decline compared to those with lower incomes.
  + *Hypothesis 6b* Those with higher levels of education are more likely to adopt new CAMs in the event of a health decline compared to those with lower levels of education.

# Data and Methods

In the following analyses, I compared values across waves in the following combinations: waves 1 and 2, waves 2 and 3, and waves 1 and 3. The *n* for each analysis varies depending on whether respondents participated in both of the included waves on all variables. The *n* for wave 1 and 2 comparisons is `r nrow(waves12.wide.complete)`, for wave 2 and 3 comparisons is `r nrow(waves23.wide.complete)`, and for wave 1 and 3 comparisons is `r nrow(waves13.wide.complete)`.

This chapter includes three new variables in the analysis: difference in CAMs, difference in chronic conditions, and difference in SRH across waves. For each wave comparison, I subtracted the value from the first wave from the value at the second wave to create the difference variable. For example, I created the difference in CAMs variable for the wave 1 and 2 comparison by subtracting number of CAMs used at wave 1 from the number of CAMs used at wave 2. A positive value indicates that the respondent used more CAMs, developed more chronic conditions, or reported better SRH. Negative values indicate that the respondent used fewer CAMs, had fewer chronic conditions, or had poorer SRH compared to the previous wave. A value of zero indicates no difference between waves. 

Continuous variables that I used to generate interaction terms (difference in chronic conditions, difference in self-reported health, difference in log income) are centered around the mean for ease of interpretation. 

All other variables are coded in the same way as described in the previous chapter. I used ordinary least squares regression to analyze the data.

# Results

Descriptive statistics for each wave comparison sample can be found in Tables \ref{tab:waves12-descriptives-table} through \ref{tab:waves13-descriptives-table}.

```{r waves12-descriptives-table, echo = FALSE}
waves12.wide.complete %>% 
  dplyr::select(DiffCams, DiffChron, DiffSRH, Income_W1, Education_W1, Age_W2, Sex_W1, Race_W2, Marital_W2, HlthInsurance_W2, HLSelf_W2, Spiritual_W2, Religion_W2) %>%
  gtsummary::tbl_summary(
    type = list(DiffCams ~ "continuous2",
                DiffChron ~ "continuous2",
                DiffSRH ~ "continuous2", 
                Income_W1 ~ "continuous2", 
                Age_W2 ~ "continuous2",
                HLSelf_W2 ~ "continuous2",
                Spiritual_W2 ~ "continuous2", 
                Religion_W2 ~ "continuous2"),
    statistic = list(all_continuous() ~ c("{N_nonmiss}/{N_obs}", 
                                          "{mean} ({sd})",
                                          "{min}, {max}"),
                     all_categorical() ~ "{n}/{N_nonmiss} ({p})%"),
    digits = all_continuous() ~ 2, 
    label = list(DiffCams ~ "Difference in Total CAMs Used", 
                 DiffChron ~ "Difference in Chronic Conditions",
                 DiffSRH ~ "Difference in Self-Reported Health", 
                 Income_W1 ~ "Wave 1 Income", 
                 Education_W1 ~ "Wave 1 Education",
                 Age_W2 ~ "Wave 2 Age", 
                 Sex_W1 ~ "Sex", 
                 Race_W2 ~ "Race",
                 Marital_W2 ~ "Wave 2 Marital Status",
                 HlthInsurance_W2 ~ "Wave 2 Health Insurance Status",
                 HLSelf_W2 ~ "Wave 2 Health Locus of Control",
                 Spiritual_W2 ~ "Wave 2 Spirituality",
                 Religion_W2 ~ "Wave 2 Religiosity"),
    missing_text = "(Missing)") %>% 
    bold_labels() %>% 
    as_kable_extra(
      longtable = TRUE,
      booktabs=TRUE,
      caption="Descriptive Statistics Wave 1 2 Comparison",
      escape = F) %>%
  kable_classic(html_font = "CMU Serif") %>%
  kable_styling(latex_options = c("striped", "repeat_header"),
                font_size = 7) 

```

```{r waves23-descriptives-table, echo = FALSE}
waves23.wide.complete %>% 
  dplyr::select(DiffCams, DiffChron, DiffSRH, Income_W2, Education_W2, Age_W3, Sex_W2, Race_W2, Marital_W3, HlthInsurance_W3, HLSelf_W3, Spiritual_W3, Religion_W3) %>%
  gtsummary::tbl_summary(
    type = list(DiffCams ~ "continuous2",
                DiffChron ~ "continuous2",
                DiffSRH ~ "continuous2", 
                Income_W2 ~ "continuous2", 
                Age_W3 ~ "continuous2",
                HLSelf_W3 ~ "continuous2",
                Spiritual_W3 ~ "continuous2", 
                Religion_W3 ~ "continuous2"),
    statistic = list(all_continuous() ~ c("{N_nonmiss}/{N_obs}", 
                                          "{mean} ({sd})",
                                          "{min}, {max}"),
                     all_categorical() ~ "{n}/{N_nonmiss} ({p})%"),
    digits = all_continuous() ~ 2, 
    label = list(DiffCams ~ "Difference in Total CAMs Used", 
                 DiffChron ~ "Difference in Chronic Conditions",
                 DiffSRH ~ "Difference in Self-Reported Health", 
                 Income_W2 ~ "Wave 2 Income", 
                 Education_W2 ~ "Wave 2 Education",
                 Age_W3 ~ "Wave 3 Age", 
                 Sex_W2 ~ "Sex", 
                 Race_W2 ~ "Race",
                 Marital_W3 ~ "Wave 3 Marital Status",
                 HlthInsurance_W3 ~ "Wave 3 Health Insurance Status",
                 HLSelf_W3 ~ "Wave 3 Health Locus of Control",
                 Spiritual_W3 ~ "Wave 3 Spirituality",
                 Religion_W3 ~ "Wave 3 Religiosity"),
    missing_text = "(Missing)") %>% 
    bold_labels() %>% 
    as_kable_extra(
      longtable = TRUE,
      booktabs=TRUE,
      caption="Descriptive Statistics Wave 2 3 Comparison",
      escape = F) %>%
  kable_classic(html_font = "CMU Serif") %>%
  kable_styling(latex_options = c("striped", "repeat_header"),
                font_size = 7) 

```

```{r waves13-descriptives-table, echo = FALSE}
waves13.wide.complete %>% 
  dplyr::select(DiffCams, DiffChron, DiffSRH, Income_W1, Education_W1, Age_W3, Sex_W1, Race_W3, Marital_W3, HlthInsurance_W3, HLSelf_W3, Spiritual_W3, Religion_W3) %>%
  gtsummary::tbl_summary(
    type = list(DiffCams ~ "continuous2",
                DiffChron ~ "continuous2", 
                DiffSRH ~ "continuous2", 
                Income_W1 ~ "continuous2", 
                Age_W3 ~ "continuous2",
                HLSelf_W3 ~ "continuous2",
                Spiritual_W3 ~ "continuous2", 
                Religion_W3 ~ "continuous2"),
    statistic = list(all_continuous() ~ c("{N_nonmiss}/{N_obs}", 
                                          "{mean} ({sd})",
                                          "{min}, {max}"),
                     all_categorical() ~ "{n}/{N_nonmiss} ({p})%"),
    digits = all_continuous() ~ 2, 
    label = list(DiffCams ~ "Difference in Total CAMs Used", 
                 DiffChron ~ "Difference in Chronic Conditions",
                 DiffSRH ~ "Difference in Self-Reported Health", 
                 Income_W1 ~ "Wave 1 Income", 
                 Education_W1 ~ "Wave 1 Education",
                 Age_W3 ~ "Wave 3 Age", 
                 Sex_W1 ~ "Sex", 
                 Race_W3 ~ "Race",
                 Marital_W3 ~ "Wave 3 Marital Status",
                 HlthInsurance_W3 ~ "Wave 3 Health Insurance Status",
                 HLSelf_W3 ~ "Wave 3 Health Locus of Control",
                 Spiritual_W3 ~ "Wave 3 Spirituality",
                 Religion_W3 ~ "Wave 3 Religiosity"),
    missing_text = "(Missing)") %>% 
    bold_labels() %>% 
    as_kable_extra(
      longtable = TRUE,
      booktabs=TRUE,
      caption="Descriptive Statistics Wave 1 3 Comparison",
      escape = F) %>%
  kable_classic(html_font = "CMU Serif") %>%
  kable_styling(latex_options = c("striped", "repeat_header"),
                font_size = 7) 

```

## Sum of Chronic Conditions 

### Income as Moderator

Tables \@ref{tab:waves12-chron-mod-income} through \ref{tab:waves13-chron-mod-income} present the OLS results for the sum of chronic conditions, and the results do not support *Hypothesis 6a*. Income does not moderate the relationship between changes in total CAMs used and differences in chronic conditions in any combination of waves. Figures \ref{fig:plot-waves12-chron-income-3} through \ref{fig:plot-waves13-chron-income-3} show plots of the regression lines for average income and 1 standard deviation above and below the mean to give a general sense of the interaction. There is a main effect for difference in chronic conditions on difference in CAMs in the wave 1 and 2 comparison, which suggests more CAM use in response to an increase in chronic health conditions among individuals with higher income (*b* = `r round(waves12.chron.income.1$coefficients[2],3)`, SE = `r round(waves12.chron.income.1.se[2],3)`, *p* \< 0.001). This offers some support for *Hypothesis 5*, which posited that individuals would use more CAMs in response to a new diagnosis or worsening health status. However, this effect does not appear in other wave comparisons.  

```{r waves12-chron-mod-income, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves12.chron.income.1, waves12.chron.income.2, waves12.chron.income.3, 
          title = "Waves 1 and 2 Difference in CAMs Regressed on Difference in Chronic Conditions Moderated by Income", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in Chronic Conditions", "Logged Income", 
                               "Less than High School", "HS Diploma/GED", 
                               "Some College/AA Degree", "Graduate/Professional Degree",
                               "Age at Wave 2", "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race",
                               "Married", "Insured", "Health Locus of Control",
                               "Spirituality", "Religiosity",
                               "Difference in Chronic Conditions x Logged Income"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE, 
          notes = c("White is the reference category for race.",
                    "Standard errors shown in parentheses.", 
                    "College degree is the reference category for education."), 
          notes.align = "r", 
          label = "tab:waves12-chron-mod-income",
          star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001)) 

```

```{r plot-waves12-chron-income-3, echo = FALSE, fig.cap= "Income Moderation Effects on Relationship Between Difference in Chronic Conditions and Difference in Total CAMs Used Between Waves 1 and 2"}
library(interactions)
interact_plot(waves12.chron.income.3, pred = DiffChronCenter, modx = LogIncomeCenter, 
              x.label = "Difference in Chronic Conditions", 
              y.label = "Difference in Total CAMs",
              legend.main = "Logged Income")
```

```{r waves23-chron-mod-income, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves23.chron.income.1, waves23.chron.income.2, waves23.chron.income.3, 
          title = "Waves 2 and 3 Difference in CAMs Regressed on Difference in Chronic Conditions Moderated by Income", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in Chronic Conditions", "Logged Income", 
                               "Less than High School", "HS Diploma/GED", 
                               "Some College/AA Degree", "Graduate/Professional Degree",
                               "Age at Wave 3", "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race",
                               "Married", "Insured", "Health Locus of Control", 
                               "Spirituality", "Religiosity",  
                               "Difference in Chronic Conditions x Logged Income"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("White is the reference category for race.", 
                    "Standard errors shown in parentheses.", 
                    "College degree is the referecne category for education."), 
          notes.align = "r", 
          label = "tab:waves23-chron-mod-income",
          star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001)) 

```

```{r plot-waves23-chron-income-3, echo = FALSE, fig.cap= "Income Moderation Effects on Relationship Between Difference in Chronic Conditions and Difference in Total CAMs Used Between Waves 2 and 3"}
library(interactions)
interact_plot(waves23.chron.income.3, pred = DiffChronCenter, modx = LogIncomeCenter, 
              x.label = "Difference in Chronic Conditions", 
              y.label = "Difference in Total CAMs",
              legend.main = "Logged Income")
```

```{r waves13-chron-mod-income, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves13.chron.income.1, waves13.chron.income.2, waves13.chron.income.3,
          title = "Waves 1 and 3 Difference in CAMs Regressed on Difference in Chronic Conditions Moderated by Income", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in Chronic Conditions", "Logged Income", 
                               "Less than High School", "HS Diploma/GED", 
                               "Some College/AA Degree", "Graduate/Professional Degree",
                               "Age at Wave 3", "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race",
                               "Married", "Insured", "Health Locus of Control",
                               "Spirituality", "Religiosity", 
                               "Difference in Chronic Conditions x Logged Income"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("White is the reference category for race.",
                    "Standard errors shown in parentheses.", 
                    "College degree is the referecne category for education."),
          notes.align = "r", 
          label = "tab:waves13-chron-mod-income",
          star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001))

```

```{r plot-waves13-chron-income-3, echo = FALSE, fig.cap= "Income Moderation Effects on Relationship Between Difference in Chronic Conditions and Difference in Total CAMs Used Between Waves 1 and 3"}
library(interactions)
interact_plot(waves13.chron.income.3, pred = DiffChronCenter, modx = LogIncomeCenter, 
              x.label = "Difference in Chronic Conditions", 
              y.label = "Difference in Total CAMs",
              legend.main = "Logged Income")
```

### Education as Moderator

Findings on whether education moderates the relationship between difference in chronic conditions and difference in total CAMs used were mixed between wave comparisons. Looking at differences between waves 1 and 2, we see that difference in chronic conditions has a significant positive relationship with difference in number of CAMs used (See Table \ref{tab:waves12-chron-mod-educ}; *b* = `r round (waves12.chron.educ.1$coefficients[2],3)`, SE = `r round(waves12.chron.educ.1.se[2],3)`, *p* \< 0.001). This supports *Hypothesis 5*. There is no significant difference in moderation effects for different levels of education compared to those with a college degree, though all slopes are positive indicating greater CAM use when new chronic conditions are developed (See Figure \ref{fig:plot-waves12-chron-educ-3}).

Table \ref{tab:waves23-chron-mod-educ} shows results for the wave 2 and 3 comparison. Model 1 includes difference in chronic conditions and education as predictors. Model 2 adds the interaction term. Model 3 adds controls. Looking at Model 1, we see that there is no significant main effect for change in chronic conditions on change in number of CAMs used. In Model 3, we see that the effect of difference in chronic conditions on difference in number of CAMs used is higher by 0.130 for respondents with some college/AA degree compared to those with a college degree (*b* = `r round(waves23.chron.educ.3$coefficients[20],3)`, SE = `r round(waves23.chron.educ.3.se[20], 3)`, *p* \< 0.050). For those with a graduate/professional degree compared to those with a college degree, the effect of difference in chronic conditions on difference in CAMs is higher by 0.174 (*b* = `r round(waves23.chron.educ.3$coefficients[21],3)`, SE = `r round(waves23.chron.educ.3.se[21],3)`, *p* \< 0.010). Figure \ref{fig:plot-waves23-chron-educ-3} shows the slopes for each level of education. The figure shows that among college graduates, the relationship between difference in chronic conditions and difference in CAMs is negative. College graduates who develop new chronic conditions use fewer CAMs. With the exception of college graduates, it does appear that as education level increases, the relationship between a change in chronic conditions and a change in CAMs becomes more steeply positive, providing some support for *Hypothesis 6b*. However, the impact of education as a moderator is non-linear.  

Table \ref{tab:waves13-chron-mod-educ} shows results for the wave 1 and 3 comparison. 
There is no main effect for difference in chronic conditions on difference in CAMs used between waves 1 and 3. Model 3 includes main effects, the interaction term, and controls. Results show significant differences in the magnitude of the effect of difference in chronic conditions on difference in CAMs used for respondents with less than a HS diploma (*b* = `r round(waves13.chron.educ.3$coefficients[18],3)`, SE = `r round(waves13.chron.educ.3.se[18],3)`, *p* \< 0.050) and respondents with a graduate/professional degree compared to those with a college degree (*b* = `r round(waves13.chron.educ.3$coefficients[21],3)`, SE = `r round(waves13.chron.educ.3.se[21],3)`, *p* \< 0.050) compared to respondents with a college degree. A plot of the slopes for each education level does not support *Hypothesis 6b* (See Figure \ref{fig:plot-waves13-chron-educ-3}). Again, the slope for college graduates is negative, indicating that college graduates used fewer CAMs when they developed a new chronic condition. The slopes for respondents with less than a high school diploma and respondents with a graduate or professional degree are the steepest slopes and nearly identical, indicating these groups are the most likely to adopt new CAMs when developing a chronic condition. This suggests a curvilinear relationship between SES and adoption of CAM use in response to a changes in chronic conditions.   

```{r waves12-chron-mod-educ, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves12.chron.educ.1, waves12.chron.educ.2, waves12.chron.educ.3, 
          title = "Waves 1 and 2 Difference in CAMs Regressed on Difference in Chronic Conditions Moderated by Education", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in Chronic Conditions", "Less than high school", "HS Diploma/GED", 
                               "Some College/AA Degree","Graduate or Professional Degree", 
                               "Logged Income", "Age at Wave 2",
                               "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity", 
                               "Difference in Chronic Conditions x Less than HS Diploma", 
                               "Difference in Chronic Conditions x High School Diploma/GED", 
                               "Difference in Chronic Conditions x Some College/AA Degree", 
                               "Difference in Chronic Conditions x Professional or Graduate Degree"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("College graduate is the reference category for education.",
                    "White is the reference category for race.",
                    "Standard errors shown in parentheses."), 
          notes.align = "r", 
          float.env = "sidewaystable", 
          label = "tab:waves12-chron-mod-educ",
          star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001))
```

```{r plot-waves12-chron-educ-3, echo=FALSE, fig.cap= "Education Moderation Effects on Relationship Between Difference in Chronic Conditions and Difference in Total CAMs Used Between Waves 1 and 2"}
plot_model(waves12.chron.educ.3, 
           type = "pred", 
           terms = c("DiffChronCenter", "Education_W1.2"), 
           title = "",
           axis.title = c("Difference in Chronic Conditions", "Difference in CAMs"),
           legend.title = "Education")

```

```{r waves23-chron-mod-educ, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves23.chron.educ.1, waves23.chron.educ.2, waves23.chron.educ.3, 
          title = "Waves 2 and 3 Difference in CAMs Regressed on Difference in Chronic Conditions Moderated by Education", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in Chronic Conditions", "Less than high school", "HS Diploma/GED", 
                               "Some College/AA Degree","Graduate or Professional Degree",
                               "Logged Income", "Age at Wave 3",
                               "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity", 
                               "Difference in Chronic Conditions x Less than HS Diploma", 
                               "Difference in Chronic Conditions x High School Diploma/GED", 
                               "Difference in Chronic Conditions x Some College/AA Degree", 
                               "Difference in Chronic Conditions x Professional or Graduate Degree"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("College graduate is the reference category for education.",
                    "White is the reference category for race.", 
                    "Standard errors shown in parentheses."),
          notes.align = "r", 
          float.env = "sidewaystable", 
          label = "tab:waves23-chron-mod-educ",
          star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001))
```

```{r plot-waves23-chron-educ-3, echo=FALSE, fig.cap= "Education Moderation Effects on Relationship Between Difference in Chronic Conditions and Difference in Total CAMs Used Between Waves 2 and 3"}
plot_model(waves23.chron.educ.3, 
           type = "pred", 
           terms = c("DiffChronCenter", "Education_W2.2"), 
           title = "",
           axis.title = c("Difference in Chronic Conditions", "Difference in CAMs"),
           legend.title = "Education")

```

```{r waves13-chron-mod-educ, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves13.chron.educ.1, waves13.chron.educ.2, waves13.chron.educ.3,
          title = "Waves 1 and 3 Difference in CAMs Regressed on Difference in Chronic Conditions Moderated by Education", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in Chronic Conditions", "Less than high school", "HS Diploma/GED", 
                               "Some College/AA Degree","Graduate or Professional Degree", 
                               "Logged Income", "Age at Wave 3",
                               "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity", 
                               "Difference in Chronic Conditions x Less than HS Diploma", 
                               "Difference in Chronic Conditions x High School Diploma/GED", 
                               "Difference in Chronic Conditions x Some College/AA Degree", 
                               "Difference in Chronic Conditions x Professional or Graduate Degree"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("College graduate is the reference category for education.", 
                    "White is the reference category for race.", 
                    "Standard errors shown in parentheses."), 
          notes.align = "r", 
          float.env = "sidewaystable", 
          label = "tab:waves13-chron-mod-educ",
          star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001))
```

```{r plot-waves13-chron-educ-3, echo=FALSE, fig.cap= "Education Moderation Effects on Relationship Between Difference in Chronic Conditions and Difference in Total CAMs Used Between Waves 1 and 3"}
plot_model(waves13.chron.educ.3, 
           type = "pred", 
           terms = c("DiffChronCenter", "Education_W1.2"), 
           title = "",
           axis.title = c("Difference in Chronic Conditions", "Difference in CAMs"),
           legend.title = "Education")

```

## Self-Reported Health

No relationships between changes in SRH and changes in total CAMs used were found in any of the between wave comparisons (See Tables \ref{tab:waves12-SRH-mod-income}, \ref{tab:waves23-SRH-mod-income}, \ref{tab:waves13-SRH-mod-income}, \ref{tab:waves12-SRH-mod-educ}, \ref{tab:waves23-SRH-mod-educ}, \ref{tab:waves13-SRH-mod-educ}).

### Income as Moderator

```{r waves12-SRH-mod-income, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves12.SRH.income.1, waves12.SRH.income.2, waves12.SRH.income.3,
          title = "Waves 1 and 2 Difference in CAMs Regressed on Difference in SRH Moderated by Income", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in SRH", "Logged Income", 
                               "Less than High School", "HS Diploma/GED", 
                               "Some College/AA Degree", "Graduate/Professional Degree",
                               "Age at Wave 2", "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race",
                               "Married", "Insured", "Health Locus of Control",
                               "Spirituality", "Religiosity", 
                               "Difference in SRH x Logged Income"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE, 
          notes = c("White is the reference category for race.", 
                    "Standard errors shown in parentheses.", 
                    "SRH: Self-Reported Health."), 
          notes.align = "r", 
          label = "tab:waves12-SRH-mod-income",
          star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001)) 

```

```{r plot-waves12-SRH-income-3, echo = FALSE, fig.cap= "Income Moderation Effects on Relationship Between Difference in Self-Reported Health and Difference in Total CAMs Used Between Waves 1 and 2"}
library(interactions)
interact_plot(waves12.SRH.income.3, pred = DiffSRHCenter, modx = LogIncomeCenter, 
              x.label = "Difference in Self-Reported Health", 
              y.label = "Difference in Total CAMs",
              legend.main = "Logged Income")
```

```{r waves23-SRH-mod-income, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves23.SRH.income.1, waves23.SRH.income.2, waves23.SRH.income.3, 
          title = "Waves 2 and 3 Difference in CAMs Regressed on Difference in SRH Moderated by Income", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in SRH", "Logged Income", 
                               "Less than High School", "HS Diploma/GED", 
                               "Some College/AA Degree", "Graduate/Professional Degree",
                               "Age at Wave 3", "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race",
                               "Married", "Insured", "Health Locus of Control",
                               "Spirituality", "Religiosity", 
                               "Difference in SRH x Logged Income"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE, 
          notes = c("White is the reference category for race.", 
                    "Standard errors shown in parentheses.", 
                    "SRH: Self-Reported Health."), 
          notes.align = "r", 
          label = "tab:waves23-SRH-mod-income",
          star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001)) 

```

```{r plot-waves23-SRH-income-3, echo = FALSE, fig.cap= "Income Moderation Effects on Relationship Between Difference in Self-Reported Health and Difference in Total CAMs Used Between Waves 2 and 3"}
library(interactions)
interact_plot(waves23.SRH.income.3, pred = DiffSRHCenter, modx = LogIncomeCenter, 
              x.label = "Difference in Self-Reported Health", 
              y.label = "Difference in Total CAMs",
              legend.main = "Logged Income")
```

```{r waves13-SRH-mod-income, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves13.SRH.income.1, waves13.SRH.income.2, waves13.SRH.income.3,
          title = "Waves 1 and 3 Difference in CAMs Regressed on Difference in SRH Moderated by Income", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in SRH", "Logged Income", 
                               "Less than High School", "HS Diploma/GED", 
                               "Some College/AA Degree", "Graduate/Professional Degree",
                               "Age at Wave 3", "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race",
                               "Married", "Insured", "Health Locus of Control",
                               "Spirituality", "Religiosity", 
                               "Difference in SRH x Logged Income"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("White is the reference category for race.", 
                    "Standard errors shown in parentheses.", 
                    "SRH: Self-Reported Health"), 
          notes.align = "r", 
          label = "tab:waves13-SRH-mod-income",
          star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001))

```

```{r plot-waves13-SRH-income-3, echo = FALSE, fig.cap= "Income Moderation Effects on Relationship Between Difference in Self-Reported Health and Difference in Total CAMs Used Between Waves 1 and 3"}
library(interactions)
interact_plot(waves13.SRH.income.3, pred = DiffSRHCenter, modx = LogIncomeCenter, 
              x.label = "Difference in Self-Reported Health", 
              y.label = "Difference in Total CAMs",
              legend.main = "Logged Income")
```

### Education as Moderator

```{r plot-waves12-SRH-educ-3, echo=FALSE, fig.cap= "Education Moderation Effects on Relationship Between Difference in SRH and Difference in Total CAMs Used Between Waves 1 and 2"}
plot_model(waves12.SRH.educ.3, 
           type = "pred", 
           terms = c("DiffSRHCenter", "Education_W1.2"), 
           title = "",
           axis.title = c("Difference in SRH", "Difference in CAMs"),
           legend.title = "Education")

```

```{r waves12-SRH-mod-educ, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves12.SRH.educ.1, waves12.SRH.educ.2, waves12.SRH.educ.3,  
          title = "Waves 1 and 2 Difference in CAMs Regressed on Difference in SRH Moderated by Education", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in SRH", "Less than high school", "HS Diploma/GED", 
                               "Some College/AA Degree","Graduate or Professional Degree",
                               "Logged Income", "Age at Wave 2",
                               "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity", 
                               "Difference in Chronic Conditions x Less than HS Diploma", 
                               "Difference in Chronic Conditions x High School Diploma/GED", 
                               "Difference in Chronic Conditions x Some College/AA Degree", 
                               "Difference in Chronic Conditions x Professional or Graduate Degree"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("College graduate is the reference category for education.", 
                    "White is the reference category for race.", 
                    "Standard errors shown in parentheses.", 
                    "SRH: Self-Reported Health."), 
          notes.align = "r", 
          float.env = "sidewaystable", 
          label = "tab:waves12-SRH-mod-educ",
          star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001))
```

```{r plot-waves23-SRH-educ-3, echo=FALSE, fig.cap= "Education Moderation Effects on Relationship Between Difference in SRH and Difference in Total CAMs Used Between Waves 2 and 3"}
plot_model(waves23.SRH.educ.3, 
           type = "pred", 
           terms = c("DiffSRHCenter", "Education_W2.2"), 
           title = "",
           axis.title = c("Difference in SRH", "Difference in CAMs"),
           legend.title = "Education")

```

```{r waves23-SRH-mod-educ, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves23.SRH.educ.1, waves23.SRH.educ.2, waves23.SRH.educ.3,  
          title = "Waves 2 and 3 Difference in CAMs Regressed on Difference in SRH Moderated by Education", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in SRH", "Less than high school", "HS Diploma/GED", 
                               "Some College/AA Degree", "Graduate or Professional Degree",
                               "Logged Income", "Age at Wave 2",
                               "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity", 
                               "Difference in Chronic Conditions x Less than HS Diploma", 
                               "Difference in Chronic Conditions x High School Diploma/GED", 
                               "Difference in Chronic Conditions x Some College/AA Degree", 
                               "Difference in Chronic Conditions x Professional or Graduate Degree"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("College graduate is the reference category for education.", 
                    "White is the reference category for race.", 
                    "Standard errors shown in parentheses.", 
                    "SRH: Self-Reported Health."), 
          notes.align = "r", 
          float.env = "sidewaystable", 
          label = "tab:waves12-SRH-mod-educ",
          star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001))
```

```{r plot-waves13-SRH-educ-3, echo=FALSE, fig.cap= "Education Moderation Effects on Relationship Between Difference in SRH and Difference in Total CAMs Used Between Waves 1 and 3"}
plot_model(waves12.SRH.educ.3, 
           type = "pred", 
           terms = c("DiffSRHCenter", "Education_W1.2"), 
           title = "",
           axis.title = c("Difference in SRH", "Difference in CAMs"),
           legend.title = "Education")

```

```{r waves13-SRH-mod-educ, echo = FALSE, results='asis'}
# Create regression results table
stargazer(waves13.SRH.educ.1, waves13.SRH.educ.2, waves13.SRH.educ.3, 
          title = "Waves 1 and 3 Difference in CAMs Regressed on Difference in SRH Moderated by Education", 
          column.labels = c("Model 1", "Model 2", "Model 3"), 
          covariate.labels = c("Difference in SRH", "Less than high school", "HS Diploma/GED", 
                               "Some College/AA Degree","Graduate or Professional Degree",
                               "Logged Income", "Age at Wave 2",
                               "Female", "Non-Hispanic Black", 
                               "Mexican American and Other Hispanic", "Other Race", "Married", 
                               "Insured", "Health Locus of Control", "Spirituality", "Religiosity", 
                               "Difference in Chronic Conditions x Less than HS Diploma", 
                               "Difference in Chronic Conditions x High School Diploma/GED", 
                               "Difference in Chronic Conditions x Some College/AA Degree", 
                               "Difference in Chronic Conditions x Professional or Graduate Degree"),
          dep.var.caption  = "", 
          dep.var.labels.include = FALSE,
          model.numbers          = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"), 
          column.sep.width = "3pt", # to reduce column width
          single.row = TRUE, # to put coefficients and standard errors on same line
          font.size = "small",
          header = FALSE,
          notes = c("College graduate is the reference category for education.", 
                    "White is the reference category for race.", 
                    "Standard errors shown in parentheses.", 
                    "SRH: Self-Reported Health."), 
          notes.align = "r", 
          float.env = "sidewaystable", 
          label = "tab:waves12-SRH-mod-educ",
          star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001))
```

# Conclusion 

This chapter examined the relationship between change in health status and change in use of CAM. Results provided some support for *Hypothesis 5*. A change in chronic conditions was positively related to a change in number of CAMs used in some wave comparisons. Results did not support *Hypothesis 6a*. Income does not moderate the relationship between change in health and change in CAM use. Support for *Hypothesis 6b* was mixed. The relationship between difference in chronic conditions and change in CAMs use varied across levels of education; however, this was not always in the expected direction. 