---
title: "SEM"
author: "Hannah Andrews"
date: '2022-03-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```
```{r, echo = FALSE}
library(lme4)
library(lattice)
library(sjstats)
library(jtools)
library(ROCR)
library(sjPlot)
library(performance)
library(XML)
library(kableExtra)
library(tidyverse)
library(caTools)
library(lavaan)
library(semPlot)
library(fastDummies)
setwd("C:/Users/hanna/Documents/git/AHL/R/")
load("waves123.long.complete.rda")
load("waves12.wide.complete.rda")
load("waves23.wide.complete.rda")
load("waves13.wide.complete.rda")
load("waves123.wide.rda")
setwd("C:/Users/hanna/Documents/git/AHL/R/Mixed model")
# Tutorials
# https://lmudge13.github.io/sample_code/mixed_effects.html
# Converting html to pdf https://github.com/gorkang/html2latex/
```
```{r, echo = FALSE}
model1 <- '
SumChron_W3 ~ LogIncome_W1 + Education_W1 + totalCams_W2 #SES and CAMs predict health
totalCams_W2 ~ LogIncome_W1 + Education_W1
'
waves123.wide$Education_W1 <- factor(waves123.wide$Education_W1, 
                                     levels = c("Less than High School", "HS Diploma/GED", 
                                                "Some College/AA Degree", "College Grad",
                                                "Graduate or Professional Degree"), 
                                     ordered = TRUE)

model1.summ <- sem(model1, data = waves123.wide)

fit <- sem(model, data = waves123.wide)
summary(model1.summ, standardized = TRUE)
fitMeasures(model1.summ, c("cfi", "rmsea", "srmr"))


model2.summ <- sem(model1, data = waves123.wide, missing = "ML", do.fit = TRUE)
summary(model2.summ, standardized = TRUE)
fitMeasures(model2.summ, c("cfi", "rmsea", "srmr"))

semPaths(model2.summ, what = "paths", whatLabels = "stand", rotation = 1)

m <- matrix(NA, 3, 4)
m[1, 1] <- "LogIncome_W1"
m[3, 1] <- "Education_W1"
m[2, 2] <- "totalCams_W2"
m[2, 4] <- "SumChron_W3"
m

m2 <- matrix(NA, 5, 3)
m2[2, 1] <- "LogIncome_W1"
m2[4, 1] <- "Education_W1"
m2[3, 2] <- "totalCams_W2"
m2[3, 3] <- "SumChron_W3" 

semPaths(model1.summ, whatLabels = "std",
           sizeMan = 10,
           edge.label.cex = 1.15,
           style = "ram", 
           nCharNodes = 0, nCharEdges = 0,
           layout = m)

semPaths(model2.summ, what = "paths", whatLabels = "std",
           sizeMan = 10,
           edge.label.cex = 1.15,
           style = "ram", 
           nCharNodes = 0, 
           nCharEdges = 0)
```

```{r, echo = FALSE}

         
fit <- sem(model, data = waves123.wide)
summary(fit)
fitMeasures(fit, c("cfi", "rmsea", "srmr"))

```

```{r, echo = FALSE}
# Create dummy variables for race at wave 1
library(fastDummies)
path.waves123.wide <- dummy_cols(path.waves123.wide, select_columns = 'Race_W2')
path.waves123.wide <- rename(path.waves123.wide, 
                             MexicanHispanic_W2 = "Race_W2_Mexican American and Other Hispanic",
                             Black_W2 = "Race_W2_Non-Hispanic Black", 
                             OtherRace_W2 = "Race_W2_Other Race",
                             White_W2 = "Race_W2_Non-Hispanic White")

model2 <- '
# Regressions
totalCams_W2 ~ LogIncome_W1 + Education_W1 
SumChron_W2 ~ LogIncome_W1 + Education_W1 
totalCams_W3 ~ LogIncome_W1 + Education_W1 + SumChron_W2 + totalCams_W2 + Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
SumChron_W3 ~ LogIncome_W1 + Education_W1 + totalCams_W2 + SumChron_W2 + Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
'


model1 <- '
# Regressions
totalCams_W2 ~ LogIncome_W1 + Education_W1  
SumChron_W2 ~ LogIncome_W1 + Education_W1 
totalCams_W3 ~ LogIncome_W1 + Education_W1 + SumChron_W2 + totalCams_W2
SumChron_W3 ~ LogIncome_W1 + Education_W1 + totalCams_W2 + SumChron_W2
'

path.waves123.wide <- waves123.wide %>%
  mutate(waves123.wide = factor(Education_W1, levels = c("Less than High School", "HS Diploma/GED", 
                                                "Some College/AA Degree", "College Grad",
                                                "Graduate or Professional Degree"), labels = 1:5, ordered = TRUE))

path.waves123.wide$Education_W1 <- as.numeric(path.waves123.wide$Education_W1)

path.waves123.wide <- path.waves123.wide %>%
  mutate(path.waves123.wide = factor(Education_W2, levels = c("Less than High School", "HS Diploma/GED", 
                                                "Some College/AA Degree", "College Grad",
                                                "Graduate or Professional Degree"), labels = 1:5, ordered = TRUE))

path.waves123.wide$Education_W2 <- as.numeric(path.waves123.wide$Education_W2)

path.waves123.wide <- path.waves123.wide %>%
  mutate(path.waves123.wide = factor(Education_W3, levels = c("Less than High School", "HS Diploma/GED", 
                                                "Some College/AA Degree", "College Grad",
                                                "Graduate or Professional Degree"), labels = 1:5, ordered = TRUE))

path.waves123.wide$Education_W3 <- as.numeric(path.waves123.wide$Education_W3)




model1.summ <- sem(model1, data = path.waves123.wide)
summary(model1.summ, standardized = TRUE)
fitMeasures(model1.summ, c("cfi", "rmsea", "srmr"))

model2.summ <- sem(model2, data = path.waves123.wide)
summary(model2.summ, standardized = TRUE)
fitMeasures(model2.summ, c("cfi", "rmsea", "srmr"))


m <- matrix(NA, 19, 4)
m[1, 1] <- "LogIncome_W1"
m[4, 1] <- "Education_W1"
m[2, 2] <- "totalCams_W2"
m[3, 2] <- "SumChron_W2"
m[1, 3] <- "totalCams_W3"
m[4, 3] <- "SumChron_W3"

test <- semptools::drop_nodes(
                object = semPlotModel(model2.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2"))

path_dia <- semPaths(test, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram",
           layout = m,
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

my_position_list <- c("totalCams_W3 ~ Education_W1" = .20, 
                      "SumChron_W3 ~ LogIncome_W1" = .75)
library(semptools)
path_dia2 <- set_edge_label_position(path_dia, my_position_list)
plot(path_dia2)

m[1, 4] <- "Age_W3" 
m[3, 4] <- "Sex_W1" 
m[5, 4] <- "Marital_W3" 
m[7, 4] <- "HlthInsurance_W3"
m[9, 4] <- "HLSelf_W3" 
m[11, 4] <- "Spiritual_W3"
m[13, 4] <- "Religion_W3" 
m[15, 4] <- "MexicanHispanic_W2"
m[17, 4] <- "Black_W2" 
m[19, 4] <- "OtherRace_W2"

m <- matrix(NA, 7, 10)
m[1, 1] <- "LogIncome_W1"
m[1, 10] <- "Education_W1"
m[3, 4] <- "totalCams_W2"
m[3, 7] <- "SumChron_W2"
m[5, 1] <- "totalCams_W3"
m[5, 10] <- "SumChron_W3"
m[7, 1] <- "Age_W3" 
m[7, 2] <- "Sex_W1" 
m[7, 3] <- "Marital_W3" 
m[7, 4] <- "HlthInsurance_W3"
m[7, 5] <- "HLSelf_W3" 
m[7, 6] <- "Spiritual_W3"
m[7, 7] <- "Religion_W3" 
m[7, 8] <- "MexicanHispanic_W2"
m[7, 9] <- "Black_W2" 
m[7, 10] <- "OtherRace_W2"







path.waves123.wide$diffeducation <- path.waves123.wide$Education_W3 - path.waves123.wide$Education_W1
hist(path.waves123.wide$diffeducation)
str(waves123.wide$diffeducation)

```

# Education and race recodes 
```{r, echo = FALSE, results='hide'}
# Order Education levels
path.waves123.wide <- waves123.wide %>%
  mutate(waves123.wide = factor(Education_W1, levels = c("Less than High School", "HS Diploma/GED", 
                                                "Some College/AA Degree", "College Grad",
                                                "Graduate or Professional Degree"), labels = 1:5, ordered = TRUE))
# Convert to numeric to be used as ordinal variable
path.waves123.wide$Education_W1 <- as.numeric(path.waves123.wide$Education_W1)
# Order education levels
path.waves123.wide <- path.waves123.wide %>%
  mutate(path.waves123.wide = factor(Education_W2, levels = c("Less than High School", "HS Diploma/GED", 
                                                "Some College/AA Degree", "College Grad",
                                                "Graduate or Professional Degree"), labels = 1:5, ordered = TRUE))
# Convert to numeric to be used as ordinal variable
path.waves123.wide$Education_W2 <- as.numeric(path.waves123.wide$Education_W2)
# Order Education levels
path.waves123.wide <- path.waves123.wide %>%
  mutate(path.waves123.wide = factor(Education_W3, levels = c("Less than High School", "HS Diploma/GED", 
                                                "Some College/AA Degree", "College Grad",
                                                "Graduate or Professional Degree"), labels = 1:5, ordered = TRUE))
# Convert to numeric to be used as ordinal variable
path.waves123.wide$Education_W3 <- as.numeric(path.waves123.wide$Education_W3)

# Create dummy variables for race
path.waves123.wide <- dummy_cols(path.waves123.wide, select_columns = 'Race_W2', ignore_na = TRUE)
path.waves123.wide <- rename(path.waves123.wide, 
                             MexicanHispanic_W2 = "Race_W2_Mexican American and Other Hispanic",
                             Black_W2 = "Race_W2_Non-Hispanic Black", 
                             OtherRace_W2 = "Race_W2_Other Race",
                             White_W2 = "Race_W2_Non-Hispanic White")

# Order SRH levels
path.waves123.wide <- path.waves123.wide %>%
  mutate(waves123.wide = factor(SRH_W2, levels = c("Poor", "Fair", "Good",
                                                   "Very Good", "Excellent"), 
                                labels = 1:5, 
                                ordered = TRUE))
# Convert to numeric to be used as ordinal variable
path.waves123.wide$SRH_W2 <- as.numeric(path.waves123.wide$SRH_W2)
# Order SRH levels
path.waves123.wide <- path.waves123.wide %>%
  mutate(waves123.wide = factor(SRH_W3, levels = c("Poor", "Fair", "Good",
                                                   "Very Good", "Excellent"), 
                                labels = 1:5, 
                                ordered = TRUE))
# Convert to numeric to be used as ordinal variable
path.waves123.wide$SRH_W3 <- as.numeric(path.waves123.wide$SRH_W3)
```

# Full model as in chapter 05/03

```{r, echo = FALSE}
model <- '
# Regressions
totalCams_W2 ~ LogIncome_W1 + Education_W1 + Sex_W1 + Age_W2 + Marital_W2 + HlthInsurance_W2 + HLSelf_W2 + Spiritual_W2 + Religion_W2 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
SumChron_W2 ~ LogIncome_W1 + Education_W1 + Sex_W1 + Age_W2 + Marital_W2 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
totalCams_W3 ~ LogIncome_W1 + Education_W1 + SumChron_W2 + totalCams_W2 + Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
SumChron_W3 ~ LogIncome_W1 + Education_W1 + totalCams_W2 + SumChron_W2 + Age_W3 + Sex_W1 + Marital_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
# Covariances
totalCams_W2 ~~ SumChron_W2
totalCams_W3 ~~ SumChron_W3
'

# Run model and get summary and fit statistics
model.summ <- sem(model, data = path.waves123.wide)
summary(model.summ, standardized = TRUE)
fitMeasures(model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 4, 3)
m[1, 1] <- "LogIncome_W1"
m[4, 1] <- "Education_W1"
m[2, 2] <- "totalCams_W2"
m[3, 2] <- "SumChron_W2"
m[1, 3] <- "totalCams_W3"
m[4, 3] <- "SumChron_W3"

# Drop controls from path diagram
path_dia <- semptools::drop_nodes(
                object = semPlotModel(model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
path_dia2 <- semPaths(path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram",
           layout = m, 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

# Create list to rearrange positions of node edges that were blocking each other
my_position_list <- c("totalCams_W3 ~ Education_W1" = .20, 
                      "SumChron_W3 ~ LogIncome_W1" = .75)
# Change label position
path_dia3 <- set_edge_label_position(path_dia2, my_position_list)

path_dia4 <- change_node_label(path_dia3,
                           c(LogIncome_W1 = "Income W1",
                             Education_W1 = "Education W1",
                             totalCams_W2 = "Total CAMs W2",
                             SumChron_W2 = "Sum Chron W2", 
                             totalCams_W3 = "Total CAMs W3",
                             SumChron_W3 = "Sum Chron W3"),
                           label.cex = 1.1)

path_dia5 <- mark_sig(path_dia4, model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))





```
```{r, echo = FALSE}
# Create final plot
plot(path_dia5)
```

# Simple Model

```{r, echo = FALSE}
sim.model <- '
# Regressions
totalCams_W2 ~ LogIncome_W1 + Education_W1 
SumChron_W2 ~ LogIncome_W1 + Education_W1 
totalCams_W3 ~ LogIncome_W1 + Education_W1 + SumChron_W2 + totalCams_W2 
SumChron_W3 ~ LogIncome_W1 + Education_W1 + totalCams_W2 + SumChron_W2 
# Covariances
totalCams_W2 ~~ SumChron_W2
totalCams_W3 ~~ SumChron_W3
'

# Run model and get summary and fit statistics
sim.model.summ <- sem(sim.model, data = path.waves123.wide)
summary(sim.model.summ, standardized = TRUE)
fitMeasures(sim.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 4, 3)
m[1, 1] <- "LogIncome_W1"
m[4, 1] <- "Education_W1"
m[2, 2] <- "totalCams_W2"
m[3, 2] <- "SumChron_W2"
m[1, 3] <- "totalCams_W3"
m[4, 3] <- "SumChron_W3"

# Create path diagram
sim.path <- semPaths(sim.model.summ, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram",
           layout = m, 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

# Create list to rearrange positions of node edges that were blocking each other
my_position_list <- c("totalCams_W3 ~ Education_W1" = .20, 
                      "SumChron_W3 ~ LogIncome_W1" = .75)
# Change label position
sim.path.2 <- set_edge_label_position(sim.path, my_position_list)

sim.path.3 <- change_node_label(sim.path.2,
                           c(LogIncome_W1 = "Income W1",
                             Education_W1 = "Education W1",
                             totalCams_W2 = "Total CAMs W2",
                             SumChron_W2 = "Sum Chron W2", 
                             totalCams_W3 = "Total CAMs W3",
                             SumChron_W3 = "Sum Chron W3"),
                           label.cex = 1.1)

sim.path.4 <- mark_sig(sim.path.3, model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))


```
# Simple and Full Path Diagrams

```{r, echo = FALSE}
# Create final plot
plot(sim.path.4)
plot(path_dia5)
```

# Fit Indices Simply and Full 

```{r, echo = FALSE}
fitMeasures(sim.model.summ, c("cfi", "rmsea", "srmr"))
fitMeasures(model.summ, c("cfi", "rmsea", "srmr"))
```



# Self Reported Health

```{r, echo = FALSE}
srh.model <- '
# Regressions
totalCams_W2 ~ LogIncome_W1 + Education_W1 + Sex_W1 + Age_W2 + Marital_W2 + HlthInsurance_W2 + HLSelf_W2 + Spiritual_W2 + Religion_W2 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
SRH_W2 ~ LogIncome_W1 + Education_W1 + Sex_W1 + Age_W2 + Marital_W2 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
totalCams_W3 ~ LogIncome_W1 + Education_W1 + SRH_W2 + totalCams_W2 + Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
SRH_W3 ~ LogIncome_W1 + Education_W1 + totalCams_W2 + SRH_W2 + Age_W3 + Sex_W1 + Marital_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
# Covariances
totalCams_W2 ~~ SRH_W2
totalCams_W3 ~~ SRH_W3
'

# Run model and get summary and fit statistics
srh.model.summ <- sem(srh.model, data = path.waves123.wide)
summary(srh.model.summ, standardized = TRUE)
fitMeasures(srh.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 4, 3)
m[1, 1] <- "LogIncome_W1"
m[4, 1] <- "Education_W1"
m[2, 2] <- "totalCams_W2"
m[3, 2] <- "SRH_W2"
m[1, 3] <- "totalCams_W3"
m[4, 3] <- "SRH_W3"

# Drop controls from path diagram
srh.path_dia <- semptools::drop_nodes(
                object = semPlotModel(srh.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
srh.path_dia2 <- semPaths(srh.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram",
           layout = m, 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

# Create list to rearrange positions of node edges that were blocking each other
my_position_list <- c("totalCams_W3 ~ Education_W1" = .20, 
                      "SRH_W3 ~ LogIncome_W1" = .75)
# Change label position
srh.path_dia3 <- set_edge_label_position(srh.path_dia2, my_position_list)

srh.path_dia4 <- change_node_label(srh.path_dia3,
                           c(LogIncome_W1 = "Income W1",
                             Education_W1 = "Education W1",
                             totalCams_W2 = "Total CAMs W2",
                             SRH_W2 = "SRH W2", 
                             totalCams_W3 = "Total CAMs W3",
                             SRH_W3 = "SRH W3"),
                           label.cex = 1.1)

srh.path_dia5 <- mark_sig(srh.path_dia4, srh.model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))


plot(srh.path_dia5)


```
# Separate Models for Each Education Level 

## Less than High School 

```{r, echo=FALSE}
df.lshs <- subset(path.waves123.wide,Education_W1==1)

ls.model <- '
totalCams_W3 ~ totalCams_W2 + SumChron_W2 + Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
SumChron_W3 ~ totalCams_W2 + SumChron_W2 + Age_W3 + Sex_W1 + Marital_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
# Covariances
totalCams_W3 ~~ SumChron_W3
totalCams_W2 ~~ SumChron_W2
'

# Run model and get summary and fit statistics
ls.model.summ <- sem(ls.model, data = df.lshs)
summary(ls.model.summ, standardized = TRUE)
fitMeasures(ls.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 4, 3)
m[1, 1] <- "LogIncome_W1"
m[4, 1] <- "Education_W1"
m[1, 2] <- "totalCams_W2"
m[4, 2] <- "SumChron_W2"
m[1, 3] <- "totalCams_W3"
m[4, 3] <- "SumChron_W3"

# Drop controls from path diagram
ls.path_dia <- semptools::drop_nodes(
                object = semPlotModel(ls.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
ls.path <- semPaths(ls.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram", 
           exoCov = FALSE,
           layout = m, 
           nCharNodes = 0, nCharEdges = 0)

# Create list to rearrange positions of node edges that were blocking each other
my_position_list <- c("totalCams_W3 ~ SumChron_W2" = .20, 
                      "SumChron_W3 ~ totalCams_W2" = .75)
# Change label position
ls.path2 <- set_edge_label_position(ls.path, my_position_list)


ls.path3 <- mark_sig(ls.path2, model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))


plot(ls.path3)



```

## High School 

```{r, echo=FALSE}
df.hs <- subset(path.waves123.wide,Education_W1==2)

hs.model <- '
totalCams_W3 ~ totalCams_W2 + SumChron_W2 + Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
SumChron_W3 ~ totalCams_W2 + SumChron_W2 + Age_W3 + Sex_W1 + Marital_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
# Covariances
totalCams_W3 ~~ SumChron_W3
totalCams_W2 ~~ SumChron_W2
'

# Run model and get summary and fit statistics
hs.model.summ <- sem(hs.model, data = df.hs)
summary(hs.model.summ, standardized = TRUE)
fitMeasures(hs.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 4, 3)
m[1, 1] <- "LogIncome_W1"
m[4, 1] <- "Education_W1"
m[1, 2] <- "totalCams_W2"
m[4, 2] <- "SumChron_W2"
m[1, 3] <- "totalCams_W3"
m[4, 3] <- "SumChron_W3"

# Drop controls from path diagram
hs.path_dia <- semptools::drop_nodes(
                object = semPlotModel(hs.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
hs.path <- semPaths(hs.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram", 
           layout = m, 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

# Create list to rearrange positions of node edges that were blocking each other
my_position_list <- c("totalCams_W3 ~ SumChron_W2" = .20, 
                      "SumChron_W3 ~ totalCams_W2" = .75)
# Change label position
hs.path2 <- set_edge_label_position(hs.path, my_position_list)


hs.path3 <- mark_sig(hs.path2, model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))


plot(hs.path3)



```

## Some College 

```{r, echo=FALSE}
df.sc <- subset(path.waves123.wide,Education_W1==3)

sc.model <- '
totalCams_W3 ~ totalCams_W2 + SumChron_W2 + Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
SumChron_W3 ~ totalCams_W2 + SumChron_W2 + Age_W3 + Sex_W1 + Marital_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
# Covariances
totalCams_W3 ~~ SumChron_W3
totalCams_W2 ~~ SumChron_W2
'

# Run model and get summary and fit statistics
sc.model.summ <- sem(sc.model, data = df.sc)
summary(sc.model.summ, standardized = TRUE)
fitMeasures(sc.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 4, 3)
m[1, 1] <- "LogIncome_W1"
m[4, 1] <- "Education_W1"
m[1, 2] <- "totalCams_W2"
m[4, 2] <- "SumChron_W2"
m[1, 3] <- "totalCams_W3"
m[4, 3] <- "SumChron_W3"

# Drop controls from path diagram
sc.path_dia <- semptools::drop_nodes(
                object = semPlotModel(sc.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
sc.path <- semPaths(sc.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram", 
           layout = m,
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

# Create list to rearrange positions of node edges that were blocking each other
my_position_list <- c("totalCams_W3 ~ SumChron_W2" = .20, 
                      "SumChron_W3 ~ totalCams_W2" = .75)
# Change label position
sc.path2 <- set_edge_label_position(sc.path, my_position_list)


sc.path3 <- mark_sig(sc.path2, model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))


plot(sc.path3)



```

## College Grad 

```{r, echo=FALSE}
df.cg <- subset(path.waves123.wide,Education_W1==4)

cg.model <- '
totalCams_W3 ~ totalCams_W2 + SumChron_W2 + Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
SumChron_W3 ~ totalCams_W2 + SumChron_W2 + Age_W3 + Sex_W1 + Marital_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
# Covariances
totalCams_W3 ~~ SumChron_W3
totalCams_W2 ~~ SumChron_W2
'

# Run model and get summary and fit statistics
cg.model.summ <- sem(cg.model, data = df.cg)
summary(cg.model.summ, standardized = TRUE)
fitMeasures(cg.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 4, 3)
m[1, 1] <- "LogIncome_W1"
m[4, 1] <- "Education_W1"
m[1, 2] <- "totalCams_W2"
m[4, 2] <- "SumChron_W2"
m[1, 3] <- "totalCams_W3"
m[4, 3] <- "SumChron_W3"

# Drop controls from path diagram
cg.path_dia <- semptools::drop_nodes(
                object = semPlotModel(cg.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
cg.path <- semPaths(cg.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram",
           layout = m, 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

# Create list to rearrange positions of node edges that were blocking each other
my_position_list <- c("totalCams_W3 ~ SumChron_W2" = .20, 
                      "SumChron_W3 ~ totalCams_W2" = .75)
# Change label position
cg.path2 <- set_edge_label_position(cg.path, my_position_list)


cg.path3 <- mark_sig(cg.path2, model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))


plot(cg.path3)



```

## Grad/Professional

```{r, echo=FALSE}
df.gp <- subset(path.waves123.wide,Education_W1==5)

gp.model <- '
totalCams_W3 ~ totalCams_W2 + SumChron_W2 + Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
SumChron_W3 ~ totalCams_W2 + SumChron_W2 + Age_W3 + Sex_W1 + Marital_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
# Covariances
totalCams_W3 ~~ SumChron_W3
totalCams_W2 ~~ SumChron_W2
'

# Run model and get summary and fit statistics
gp.model.summ <- sem(gp.model, data = df.gp)
summary(gp.model.summ, standardized = TRUE)
fitMeasures(gp.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 4, 3)
m[1, 1] <- "LogIncome_W1"
m[4, 1] <- "Education_W1"
m[1, 2] <- "totalCams_W2"
m[4, 2] <- "SumChron_W2"
m[1, 3] <- "totalCams_W3"
m[4, 3] <- "SumChron_W3"

# Drop controls from path diagram
gp.path_dia <- semptools::drop_nodes(
                object = semPlotModel(gp.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
gp.path <- semPaths(gp.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram", 
           layout = m, 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

# Create list to rearrange positions of node edges that were blocking each other
my_position_list <- c("totalCams_W3 ~ SumChron_W2" = .20, 
                      "SumChron_W3 ~ totalCams_W2" = .75)
# Change label position
gp.path2 <- set_edge_label_position(gp.path, my_position_list)


gp.path3 <- mark_sig(gp.path2, model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))


plot(gp.path3)



```

## Multiple Group Analysis

```{r, echo=FALSE}
group.model <- '
totalCams_W3 ~ totalCams_W2 + SumChron_W2 + Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
SumChron_W3 ~ totalCams_W2 + SumChron_W2 + Age_W3 + Sex_W1 + Marital_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
# Covariances
totalCams_W3 ~~ SumChron_W3
totalCams_W2 ~~ SumChron_W2
'


# Run model and get summary and fit statistics
group.model.summ <- sem(group.model, data = path.waves123.wide, group = "Education_W1")
summary(group.model.summ, standardized = TRUE)
fitMeasures(group.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 4, 3)
m[1, 1] <- "LogIncome_W1"
m[4, 1] <- "Education_W1"
m[1, 2] <- "totalCams_W2"
m[4, 2] <- "SumChron_W2"
m[1, 3] <- "totalCams_W3"
m[4, 3] <- "SumChron_W3"

# Drop controls from path diagram
group.path_dia <- semptools::drop_nodes(
                object = semPlotModel(group.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
group.path <- semPaths(group.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram", 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

# Create list to rearrange positions of node edges that were blocking each other
my_position_list <- c("totalCams_W3 ~ SumChron_W2" = .20, 
                      "SumChron_W3 ~ totalCams_W2" = .75)
# Change label position
group.path2 <- set_edge_label_position(group.path, my_position_list)


group.path3 <- mark_sig(group.path2, group.model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))


plot(group.path3)



```

# Income Moderator

## Create Interaction terms and regress sumchron*income on total cams

```{r, echo = FALSE}
path.waves123.wide$LogIncomeCenter_W1 <- scale(path.waves123.wide$LogIncome_W1, scale = FALSE)
path.waves123.wide$SumChronCenter_W2 <- scale(path.waves123.wide$SumChron_W2, scale = FALSE)
path.waves123.wide$totalCamsCenter_W2 <- scale(path.waves123.wide$totalCams_W2, scale = FALSE)
path.waves123.wide$SumChronxLogIncome <- path.waves123.wide$SumChronCenter_W2*path.waves123.wide$LogIncomeCenter_W1
path.waves123.wide$totalCamsxLogIncome <- path.waves123.wide$totalCamsCenter_W2*path.waves123.wide$LogIncomeCenter_W1

# Add interaction term, just one outcome not controlling for totalCams at previous wave 
model = '
totalCams_W3 ~ SumChron_W2 + SumChronxLogIncome + Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
'
model.summ <- sem(model, data = path.waves123.wide)
summary(model.summ, standardized = TRUE, fit.measures=TRUE)

```

## Regress totalcams*logincome on sumchron

```{r,echo = FALSE}
model = '
SumChron_W3 ~ totalCams_W2 + totalCamsxLogIncome + Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
'
model.summ <- sem(model, data = path.waves123.wide)
summary(model.summ, standardized = TRUE, fit.measures=TRUE)
```

## Difference variables: Change in chronic conditions predict change in cams, income moderator wave 1 and wave 2

```{r, echo = FALSE}
# Subtract wave 1 from wave 2 chronic conditions 
path.waves123.wide$DiffChron_W1W2 <- path.waves123.wide$SumChron_W2 - path.waves123.wide$SumChron_W1
# Center 
path.waves123.wide$DiffChronCenter_W1W2 <- scale(path.waves123.wide$DiffChron_W1W2, scale = FALSE)
# Create interaction term 
path.waves123.wide$DiffChronxLogIncome <- path.waves123.wide$DiffChronCenter_W1W2*path.waves123.wide$LogIncomeCenter_W1
#Subtract wave 1 total cams from wave 2
path.waves123.wide$DiffCams_W1W2 <- path.waves123.wide$totalCams_W2 - path.waves123.wide$totalCams_W1

model = '
DiffCams_W1W2 ~ DiffChron_W1W2 + DiffChronxLogIncome
'
model.summ <- sem(model, data = path.waves123.wide)
summary(model.summ, standardized = TRUE, fit.measures=TRUE)

```


# Separate mediation models - chronic conditions

## Total Cams

### Income chron cam 

```{r, echo = FALSE}
set.seed(1234)
IncChroCam.model <- '
# Regressions
# Direct effect - log income 
totalCams_W3 ~ a*LogIncome_W1 
# Mediator effects - Income 
SumChron_W2 ~ e*LogIncome_W1 
totalCams_W3 ~ f*SumChron_W2
# Indirect effect 
ef := e*f
# total effect
total1 := a + (e*f)
# Controls
SumChron_W2 ~ Sex_W1 + Age_W2 + Marital_W2 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
totalCams_W3 ~ Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
'

# Run model and get summary and fit statistics
IncChroCam.model.summ <- sem(IncChroCam.model, data = path.waves123.wide, se = "bootstrap", bootstrap = 1000)
summary(IncChroCam.model.summ, standardized = TRUE)
#parameterestimates(IncChroCam.model.summ, boot.ci.type = "bca.simple", standardized = TRUE) %>% 
#  kable()
fitMeasures(IncChroCam.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 2, 3)
m[2, 1] <- "LogIncome_W1"
m[1, 2] <- "SumChron_W2"
m[2, 3] <- "totalCams_W3"

# Drop controls from path diagram
IncChroCam.path_dia <- semptools::drop_nodes(
                object = semPlotModel(IncChroCam.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
IncChroCam.path_dia2 <- semPaths(IncChroCam.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram",
           layout = m, 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

IncChroCam.path_dia3 <- change_node_label(IncChroCam.path_dia2,
                           c(LogIncome_W1 = "Income W1",
                             SumChron_W2 = "Sum Chron W2", 
                             totalCams_W3 = "Total CAMs W3"),
                           label.cex = 1.1)

IncChroCam.path_dia4 <- mark_sig(IncChroCam.path_dia3, IncChroCam.model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))

plot(IncChroCam.path_dia4)


```
Every 1% increase in logged income at wave 1 is associated with a .27 decrease in chronic conditions at wave 2. Adjusting for logged income, every additional chronic condition at wave 2 is associated with using 0.103 additional CAMs at wave 3. Logged income does not have a direct effect on total CAMs at wave 3 independent of the presence of chronic conditions.

### Income cam chron 

```{r, echo = FALSE}
set.seed(1234)
IncCamChro.model <- '
# Regressions
# Direct effect - log income 
SumChron_W3 ~ b*LogIncome_W1
# Mediator effects - Income 
totalCams_W2 ~ g*LogIncome_W1
SumChron_W3 ~ h*totalCams_W2
# Indirect effect 
gh := g*h
# total effect
total2 := b + (g*h)
# Controls
totalCams_W2 ~ Sex_W1 + Age_W2 + Marital_W2 + HlthInsurance_W2 + HLSelf_W2 + Spiritual_W2 + Religion_W2 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
SumChron_W3 ~ Age_W3 + Sex_W1 + Marital_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
'


# Run model and get summary and fit statistics
IncCamChro.model.summ <- sem(IncCamChro.model, data = path.waves123.wide, se = "bootstrap", bootstrap = 1000)
summary(IncCamChro.model.summ, standardized = TRUE)
fitMeasures(IncCamChro.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 2, 3)
m[2, 1] <- "LogIncome_W1"
m[1, 2] <- "totalCams_W2"
m[2, 3] <- "SumChron_W3"

# Drop controls from path diagram
IncCamChro.path_dia <- semptools::drop_nodes(
                object = semPlotModel(IncCamChro.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
IncCamChro.path_dia2 <- semPaths(IncCamChro.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram",
           layout = m, 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

IncCamChro.path_dia3 <- change_node_label(IncCamChro.path_dia2,
                           c(LogIncome_W1 = "Income W1",
                             SumChron_W3 = "Sum Chron W3", 
                             totalCams_W2 = "Total CAMs W2"),
                           label.cex = 1.1)

IncCamChro.path_dia4 <- mark_sig(IncCamChro.path_dia3, IncCamChro.model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))

plot(IncCamChro.path_dia4)



```
Every 1% increase in logged income is associated with a 0.401 decrease in chronic conditions at wave 3. Every 1% increase in logged income at wave 1 is associated with using an additional 0.168 CAMs at wave 2. Each additional CAM used at wave 2 is associated with a 0.132 increase in chronic conditions at wave 3. For every 0.168 increase in CAMs at wave 2 there is a 0.022 increase in chronic conditions at wave 3. 

### Education chron cam 

```{r, echo = FALSE}
set.seed(1234)
EducChroCam.model <- '
# Regressions
# Direct Effect - Education
totalCams_W3 ~ c*Education_W1
# Mediator effects - Income 
totalCams_W3 ~ f*SumChron_W2
# Mediator effects - Education
SumChron_W2 ~ j*Education_W1 
# Indirect effect 
jf := j*f
# total effect
total3 := c + (j*f)
# Controls
SumChron_W2 ~ Sex_W1 + Age_W2 + Marital_W2 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
totalCams_W3 ~ Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
# Covariances
'


# Run model and get summary and fit statistics
EducChroCam.model.summ <- sem(EducChroCam.model, data = path.waves123.wide, se = "bootstrap", bootstrap = 1000)
summary(EducChroCam.model.summ, standardized = TRUE)
fitMeasures(EducChroCam.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 2, 3)
m[2, 1] <- "Education_W1"
m[1, 2] <- "SumChron_W2"
m[2, 3] <- "totalCams_W3"

# Drop controls from path diagram
EducChroCam.path_dia <- semptools::drop_nodes(
                object = semPlotModel(EducChroCam.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
EducChroCam.path_dia2 <- semPaths(EducChroCam.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram",
           layout = m, 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

EducChroCam.path_dia3 <- change_node_label(EducChroCam.path_dia2,
                           c(Education_W1 = "Education W1",
                             SumChron_W2 = "Sum Chron W2", 
                             totalCams_W3 = "Total CAMs W3"),
                           label.cex = 1.1)

EducChroCam.path_dia4 <- mark_sig(EducChroCam.path_dia3, EducChroCam.model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))

plot(EducChroCam.path_dia4)



```

### Education cam chron 

```{r, echo = FALSE}
set.seed(1234)
EducCamChro.model <- '
# Regressions
# Direct Effect - Education
SumChron_W3 ~ d*Education_W1
# Mediator effects - Income 
SumChron_W3 ~ h*totalCams_W2
# Mediator effects - Education
totalCams_W2 ~ i*Education_W1
# Indirect effect 
ih := i*h
# total effect
total4 := d + (i*h)
# Controls
totalCams_W2 ~ Sex_W1 + Age_W2 + Marital_W2 + HlthInsurance_W2 + HLSelf_W2 + Spiritual_W2 + Religion_W2 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
SumChron_W3 ~ Age_W3 + Sex_W1 + Marital_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
'


# Run model and get summary and fit statistics
EducCamChro.model.summ <- sem(EducCamChro.model, data = path.waves123.wide, se = "bootstrap", bootstrap = 1000)
summary(EducCamChro.model.summ, standardized = TRUE)
fitMeasures(EducCamChro.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 2, 3)
m[2, 1] <- "Education_W1"
m[1, 2] <- "totalCams_W2"
m[2, 3] <- "SumChron_W3"

# Drop controls from path diagram
EducCamChro.path_dia <- semptools::drop_nodes(
                object = semPlotModel(EducCamChro.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
EducCamChro.path_dia2 <- semPaths(EducCamChro.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram",
           layout = m, 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

EducCamChro.path_dia3 <- change_node_label(EducCamChro.path_dia2,
                           c(Education_W1 = "Education W1",
                             SumChron_W3 = "Sum Chron W3", 
                             totalCams_W2 = "Total CAMs W2"),
                           label.cex = 1.1)

EducCamChro.path_dia4 <- mark_sig(EducCamChro.path_dia3, EducCamChro.model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))

plot(EducCamChro.path_dia4)



```

## Mind Body 

### Income chron mb 

```{r, echo = FALSE}
set.seed(1234)
IncChroCam.model <- '
# Regressions
# Direct effect - log income 
totalCams_W3 ~ a*LogIncome_W1 
# Mediator effects - Income 
SumChron_W2 ~ e*LogIncome_W1 
totalCams_W3 ~ f*SumChron_W2
# Indirect effect 
ef := e*f
# total effect
total1 := a + (e*f)
# Controls
SumChron_W2 ~ Sex_W1 + Age_W2 + Marital_W2 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
totalCams_W3 ~ Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
'

# Run model and get summary and fit statistics
IncChroCam.model.summ <- sem(IncChroCam.model, data = path.waves123.wide, se = "bootstrap", bootstrap = 1000)
summary(IncChroCam.model.summ, standardized = TRUE)
#parameterestimates(IncChroCam.model.summ, boot.ci.type = "bca.simple", standardized = TRUE) %>% 
#  kable()
fitMeasures(IncChroCam.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 2, 3)
m[2, 1] <- "LogIncome_W1"
m[1, 2] <- "SumChron_W2"
m[2, 3] <- "totalCams_W3"

# Drop controls from path diagram
IncChroCam.path_dia <- semptools::drop_nodes(
                object = semPlotModel(IncChroCam.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
IncChroCam.path_dia2 <- semPaths(IncChroCam.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram",
           layout = m, 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

IncChroCam.path_dia3 <- change_node_label(IncChroCam.path_dia2,
                           c(LogIncome_W1 = "Income W1",
                             SumChron_W2 = "Sum Chron W2", 
                             totalCams_W3 = "Total CAMs W3"),
                           label.cex = 1.1)

IncChroCam.path_dia4 <- mark_sig(IncChroCam.path_dia3, IncChroCam.model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))

plot(IncChroCam.path_dia4)


```

### Income mb chron 

### Education chron mb 

### Education mb chron 


# Separate mediation models - SRH 

## Income SRH cam 

```{r, echo = FALSE}
set.seed(1234)
IncSrhCam.model <- '
# Regressions
# Direct effect - log income 
totalCams_W3 ~ a*LogIncome_W1 
# Mediator effects - Income 
SRH_W2 ~ e*LogIncome_W1 
totalCams_W3 ~ f*SRH_W2
# Indirect effect 
ef := e*f
# total effect
total1 := a + (e*f)
# Controls
SRH_W2 ~ Sex_W1 + Age_W2 + Marital_W2 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
totalCams_W3 ~ Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
'

# Run model and get summary and fit statistics
IncSrhCam.model.summ <- sem(IncSrhCam.model, data = path.waves123.wide, se = "bootstrap", bootstrap = 1000)
summary(IncSrhCam.model.summ, standardized = TRUE)
fitMeasures(IncSrhCam.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 2, 3)
m[2, 1] <- "LogIncome_W1"
m[1, 2] <- "SRH_W2"
m[2, 3] <- "totalCams_W3"

# Drop controls from path diagram
IncSrhCam.path_dia <- semptools::drop_nodes(
                object = semPlotModel(IncSrhCam.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
IncSrhCam.path_dia2 <- semPaths(IncSrhCam.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram",
           layout = m, 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

IncSrhCam.path_dia3 <- change_node_label(IncSrhCam.path_dia2,
                           c(LogIncome_W1 = "Income W1",
                             SRH_W2 = "SRH W2", 
                             totalCams_W3 = "Total CAMs W3"),
                           label.cex = 1.1)

IncSrhCam.path_dia4 <- mark_sig(IncSrhCam.path_dia3, IncSrhCam.model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))

plot(IncSrhCam.path_dia4)


```

## Income cam SRH 

```{r, echo = FALSE}
set.seed(1234)
IncCamSrh.model <- '
# Regressions
# Direct effect - log income 
SRH_W3 ~ b*LogIncome_W1
# Mediator effects - Income 
totalCams_W2 ~ g*LogIncome_W1
SRH_W3 ~ h*totalCams_W2
# Indirect effect 
gh := g*h
# total effect
total2 := b + (g*h)
# Controls
totalCams_W2 ~ Sex_W1 + Age_W2 + Marital_W2 + HlthInsurance_W2 + HLSelf_W2 + Spiritual_W2 + Religion_W2 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
SRH_W3 ~ Age_W3 + Sex_W1 + Marital_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
'


# Run model and get summary and fit statistics
IncCamSrh.model.summ <- sem(IncCamSrh.model, data = path.waves123.wide, se = "bootstrap", bootstrap = 1000)
summary(IncCamSrh.model.summ, standardized = TRUE)
fitMeasures(IncCamSrh.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 2, 3)
m[2, 1] <- "LogIncome_W1"
m[1, 2] <- "totalCams_W2"
m[2, 3] <- "SRH_W3"

# Drop controls from path diagram
IncCamSrh.path_dia <- semptools::drop_nodes(
                object = semPlotModel(IncCamSrh.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
IncCamSrh.path_dia2 <- semPaths(IncCamSrh.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram",
           layout = m, 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

IncCamSrh.path_dia3 <- change_node_label(IncCamSrh.path_dia2,
                           c(LogIncome_W1 = "Income W1",
                             SRH_W3 = "SRH W3", 
                             totalCams_W2 = "Total CAMs W2"),
                           label.cex = 1.1)

IncCamSrh.path_dia4 <- mark_sig(IncCamSrh.path_dia3, IncCamSrh.model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))

plot(IncCamSrh.path_dia4)



```

## Education SRH cam 

```{r, echo = FALSE}
set.seed(1234)
EducSrhCam.model <- '
# Regressions
# Direct Effect - Education
totalCams_W3 ~ c*Education_W1
# Mediator effects - Income 
totalCams_W3 ~ f*SRH_W2
# Mediator effects - Education
SRH_W2 ~ j*Education_W1 
# Indirect effect 
jf := j*f
# total effect
total3 := c + (j*f)
# Controls
SRH_W2 ~ Sex_W1 + Age_W2 + Marital_W2 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
totalCams_W3 ~ Age_W3 + Sex_W1 + Marital_W3 + HlthInsurance_W3 + HLSelf_W3 + Spiritual_W3 + Religion_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
# Covariances
'


# Run model and get summary and fit statistics
EducSrhCam.model.summ <- sem(EducSrhCam.model, data = path.waves123.wide, se = "bootstrap", bootstrap = 1000)
summary(EducSrhCam.model.summ, standardized = TRUE)
fitMeasures(EducSrhCam.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 2, 3)
m[2, 1] <- "Education_W1"
m[1, 2] <- "SRH_W2"
m[2, 3] <- "totalCams_W3"

# Drop controls from path diagram
EducSrhCam.path_dia <- semptools::drop_nodes(
                object = semPlotModel(EducSrhCam.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
EducSrhCam.path_dia2 <- semPaths(EducSrhCam.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram",
           layout = m, 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

EducSrhCam.path_dia3 <- change_node_label(EducSrhCam.path_dia2,
                           c(Education_W1 = "Education W1",
                             SRH_W2 = "SRH W2", 
                             totalCams_W3 = "Total CAMs W3"),
                           label.cex = 1.1)

EducSrhCam.path_dia4 <- mark_sig(EducSrhCam.path_dia3, EducSrhCam.model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))

plot(EducSrhCam.path_dia4)



```

## Education cam SRH 

```{r, echo = FALSE}
set.seed(1234)
EducCamSrh.model <- '
# Regressions
# Direct Effect - Education
SRH_W3 ~ d*Education_W1
# Mediator effects - Income 
SRH_W3 ~ h*totalCams_W2
# Mediator effects - Education
totalCams_W2 ~ i*Education_W1
# Indirect effect 
ih := i*h
# total effect
total4 := d + (i*h)
# Controls
totalCams_W2 ~ Sex_W1 + Age_W2 + Marital_W2 + HlthInsurance_W2 + HLSelf_W2 + Spiritual_W2 + Religion_W2 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
SRH_W3 ~ Age_W3 + Sex_W1 + Marital_W3 + MexicanHispanic_W2 + Black_W2 + OtherRace_W2
'


# Run model and get summary and fit statistics
EducCamSrh.model.summ <- sem(EducCamSrh.model, data = path.waves123.wide, se = "bootstrap", bootstrap = 1000)
summary(EducCamSrh.model.summ, standardized = TRUE)
fitMeasures(EducCamSrh.model.summ, c("cfi", "rmsea", "srmr"))
#modificationindices(model.summ)

m <- matrix(NA, 2, 3)
m[2, 1] <- "Education_W1"
m[1, 2] <- "totalCams_W2"
m[2, 3] <- "SRH_W3"

# Drop controls from path diagram
EducCamSrh.path_dia <- semptools::drop_nodes(
                object = semPlotModel(EducCamSrh.model.summ),
                nodes = c("Age_W3", "Sex_W1", "Marital_W3", "HlthInsurance_W3", "HLSelf_W3", "Spiritual_W3", "Religion_W3", "MexicanHispanic_W2", "Black_W2", "OtherRace_W2", "Age_W2", "Marital_W2", "HlthInsurance_W2", "HLSelf_W2", "Spiritual_W2", "Religion_W2"))

# Create path diagram
EducCamSrh.path_dia2 <- semPaths(EducCamSrh.path_dia, whatLabels = "std",
           sizeMan = 8,
           edge.label.cex = .8,
           edge.color = "black",
           style = "ram",
           layout = m, 
           exoCov = FALSE,
           nCharNodes = 0, nCharEdges = 0)

EducCamSrh.path_dia3 <- change_node_label(EducCamSrh.path_dia2,
                           c(Education_W1 = "Education W1",
                             SRH_W3 = "SRH W3", 
                             totalCams_W2 = "Total CAMs W2"),
                           label.cex = 1.1)

EducCamSrh.path_dia4 <- mark_sig(EducCamSrh.path_dia3, EducCamSrh.model.summ, alpha = c("(n.s.)" = 1.00, "*" = .05))

plot(EducCamSrh.path_dia4)



```

