---
title: "CFA Attempt"
author: "Hannah Andrews"
date: "6/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Old NCCAM Domains 
NCCAM domains: 
- Mind-Body Treatments: hypnosis, meditation, guided imagery, relaxation, deep breathing, pray for self, others pray, prayer groups, healing rituals, biofeedback 
- Alternative Medical Systems: Acupuncture, ayurveda, folk medicine, naturopathy, homeopathy
- Manipulative Treatments: Chiropractic, massage
- Biologically based practices: herbs, vitamins. 
```{r, echo = FALSE}
# Specify the model
#NCCAM.model <- 'mindBody =~ aHypnosis + aRelaxMeditate + aImageryTech + aPrayer + aSpiritHeal + aBiofeedback
#                alternativeSystems =~ aAcupuncture + aHomeopathy + aExerciseMove 
#                manipulativeTreatments =~ aChiropractic + aMassage
#                biologicalBased =~ aVitamins + aHerbal + a SpecialDiet'

#acams[sapply(acams, is.numeric)] <- lapply(acams[sapply(acams, is.numeric)], as.factor) # convert all columns to factors


# Fit the model 
#fit <- cfa(NCCAM.model, data = acams, ordered = TRUE)

# display summary output
#summary(fit, fit.measures = TRUE)
```
