---
title: "Committee Meeting 12-11"
author: "Hannah Andrews"
date: "12/8/2020"
output:
html_document: default
pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
dir <- "C:/Users/hanna/Documents/git/AHL/R"
setwd(dir)

# Load Libraries ####
library(expss)
library(Hmisc)
library(dplyr)
library(summarytools)
library(magrittr)
library(tidyverse)
library(mosaic)
library(lattice)
library(ggplot2)
library(scales)
library(ggthemes)
library(ztable)
library(naniar)
library(pastecs)

# Load all MIDUS data into the global environment ####
load(file = "MIDUS1_recode.rda")
load(file = "MIDUS2_recode.rda")
load(file = "MIDUS3_recode.rda")
MIDUS3_recode

options(scipen=100)
options(digits=2)
options(tibble.print_max = Inf)
```

A little guidance for the following...

Dataset: Midlife in the United States (MIDUS). There are three waves, denoted MIDUS1, MIDUS2, and MIDUS3.

Variables: 
DV: Use of completementary and alternative medicine (CAM). There are dummies for each of 15 CAM modalities indicating any use in the past year, and a summed index that represents the number of CAMs used in the past year. TBD: Latent classes (LCA) and factors. I spent about a week on LCA and have some output we can look over, but it became clear I would need more time to come up with a class structure I was confident in. 

IVs: Income, education, gender, and race.

This document is organized into three sections: single variable distributions (CAM dummies, CAM summed index, income, education, gender, and race), bivariate distributions (CAM summed index by each IV), and missing case patterns. 

# Variable Distributions

## Complementary and Alternative Medicines (CAMs)

## CAM Dummies: Have you used any of the following CAMs in the past year?

### MIDUS 1 Indiviudal CAMs

```{r, echo = FALSE}
camslist <- c("aCamEverAcupuncture", "aCamEverBiofeedback", "aCamEverChiropractor", 
                "aCamEverEnergy", "aCamEverMoveTherapy", "aCamEverHerbTherapy", 
                "aCamEverMegaVitamins", "aCamEverHomeopathy", "aCamEverHypnosis", 
                "aCamEverImageTech", "aCamEverMassage", "aCamEverPray", 
                "aCamEverMeditate", "aCamEverSpecialDiet", "aCamEverSpiritHeal")
describe(MIDUS1_recode[camslist])
```

### MIDUS 2 Individual CAMs

```{r, echo = FALSE}
camslist <- c("bCamEverAcupuncture", "bCamEverBiofeedback", "bCamEverChiropractor", 
                "bCamEverEnergy", "bCamEverMoveTherapy", "bCamEverHerbTherapy", 
                "bCamEverMegaVitamins", "bCamEverHomeopathy", "bCamEverHypnosis", 
                "bCamEverImageTech", "bCamEverMassage", "bCamEverPray", 
                "bCamEverMeditate", "bCamEverSpecialDiet", "bCamEverSpiritHeal")
describe(MIDUS2_recode[camslist])
```

### MIDUS 3 Individual CAMs

```{r, echo = FALSE}
camslist <- c("cCamEverAcupuncture", "cCamEverBiofeedback", "cCamEverChiropractor", 
                "cCamEverEnergy", "cCamEverMoveTherapy", "cCamEverHerbTherapy", 
                "cCamEverMegaVitamins", "cCamEverHomeopathy", "cCamEverHypnosis", 
                "cCamEverImageTech", "cCamEverMassage", "cCamEverPray", 
                "cCamEverMeditate", "cCamEverSpecialDiet", "cCamEverSpiritHeal")
describe(MIDUS3_recode[camslist])
```

### MIDUS 1 CAM Summed Index 

```{r, echo = FALSE}
  summary(MIDUS1_recode$aCamSum)
  stat.desc(MIDUS1_recode$aCamSum, basic = F)
  boxplot(MIDUS1_recode$aCamSum, horizontal = TRUE, main="Number of CAMs used in the past year")
  M1aCamSum <- MIDUS1_recode$aCamSum
  hist(M1aCamSum, main = "Number of CAMs used in the past year",breaks=seq(0,32,1))
  M1CamSumPlot <- ggplot(data = MIDUS1_recode, aes(x = aCamSum)) + 
                      geom_density() # Create density plot
  M1CamSumPlot
#
```

### MIDUS 2 CAM Summed Index

```{r, echo = FALSE}
  summary(MIDUS2_recode$bCamSum)
  stat.desc(MIDUS2_recode$bCamSum, basic = F)
  boxplot(MIDUS2_recode$bCamSum, horizontal = TRUE, main="Number of CAMs used in the past year")
  M2bCamSum <- MIDUS2_recode$bCamSum
  hist(M2bCamSum, main = "Number of CAMs used in the past year",breaks=seq(0,32,1))
  M2CamSumPlot <- ggplot(data = MIDUS2_recode, aes(x = bCamSum)) + 
                      geom_density() # Create density plot
  M2CamSumPlot
#
```

### MIDUS 3 CAM Summed Index

```{r, echo = FALSE}
  describe(MIDUS3_recode$cCamSum)
  stat.desc(MIDUS3_recode$cCamSum, basic = F)
  boxplot(MIDUS3_recode$cCamSum, horizontal = TRUE, main="Number of CAMs used in the past year")
  M3cCamSum <- MIDUS3_recode$cCamSum
  hist(M3cCamSum, main = "Number of CAMs used in the past year",breaks=seq(0,32,1))
  M3CamSumPlot <- ggplot(data = MIDUS3_recode, aes(x = cCamSum)) + 
                      geom_density() # Create density plot
  M3CamSumPlot
#
```

## Income 

### MIDUS 1 Household Income 

```{r, echo = FALSE}
  summary(MIDUS1_recode$aHHIncome)
  stat.desc(MIDUS1_recode$aHHIncome, basic = F)
  boxplot(MIDUS1_recode$aHHIncome, horizontal = TRUE, main="Household Income")
  M1hhincome <- MIDUS1_recode$aHHIncome
  hist(M1hhincome, main = "Household Income",breaks=seq(0, 300000, 1000))
  M1HHIncomePlot <- ggplot(data = MIDUS1_recode, aes(x = aHHIncome)) + 
                      geom_density() # Create density plot
  M1HHIncomePlot
#
```

### MIDUS 2 Household Income 

```{r, echo = FALSE}
  summary(MIDUS2_recode$bHHIncome)
  stat.desc(MIDUS2_recode$bHHIncome, basic = F)
  boxplot(MIDUS2_recode$bHHIncome, horizontal = TRUE, main="Household Income")
  M2hhincome <- MIDUS2_recode$bHHIncome
  hist(M2hhincome, main = "Household Income",breaks=seq(0, 300000, 1000))
  M2HHIncomePlot <- ggplot(data = MIDUS2_recode, aes(x = bHHIncome)) + 
                      geom_density() # Create density plot
  M2HHIncomePlot
#
```

### MIDUS 3 Household Income 

```{r, echo = FALSE}
  summary(MIDUS3_recode$cHHIncome)
  stat.desc(MIDUS3_recode$cHHIncome, basic = F)
  boxplot(MIDUS3_recode$cHHIncome, horizontal = TRUE, main="Household Income")
  M3hhincome <- MIDUS3_recode$cHHIncome
  hist(M3hhincome, main = "Household Income",breaks=seq(0, 300000, 1000))
  M3HHIncomePlot <- ggplot(data = MIDUS3_recode, aes(x = cHHIncome)) + 
                      geom_density() # Create density plot
  M3HHIncomePlot
#  
```

## Education 

### MIDUS 1 Education 

```{r, echo = FALSE}
  table(MIDUS1_recode$aEducation, useNA = "always")  
M1EducationProportions <- MIDUS1_recode %>% 
  drop_na(aEducation) %>% 
  ggplot(aes(x = aEducation)) +
  geom_bar(fill = "lightblue", aes(y = (..count..)/sum(..count..))) +
  xlab("Education") +
  scale_y_continuous(labels = scales::percent, name = "Proportion") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  M1EducationProportions
#
```

### MIDUS 2 Education 

```{r, echo = FALSE}
  table(MIDUS2_recode$bEducation, useNA = "always")  
M2EducationProportions <- MIDUS2_recode %>% 
  drop_na(bEducation) %>% 
  ggplot(aes(x = bEducation)) +
  geom_bar(fill = "lightblue", aes(y = (..count..)/sum(..count..))) +
  xlab("Education") +
  scale_y_continuous(labels = scales::percent, name = "Proportion") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  M2EducationProportions
#
```

### MIDUS 3 Education 

```{r, echo = FALSE}
  table(MIDUS3_recode$cEducation, useNA = "always")  
M3EducationProportions <- MIDUS3_recode %>% 
  drop_na(cEducation) %>% 
  ggplot(aes(x = cEducation)) +
  geom_bar(fill = "lightblue", aes(y = (..count..)/sum(..count..))) +
  xlab("Education") +
  scale_y_continuous(labels = scales::percent, name = "Proportion") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  M3EducationProportions
#
```

## Gender 

### MIDUS 1 Gender

```{r, echo = FALSE}
describe(MIDUS1_recode$aGender)
```

### MIDUS 2 Gender 

``` {r, echo = FALSE}
describe(MIDUS2_recode$bGender)
```

### MIDUS 3 Gender 

```{r, echo = FALSE}
describe(MIDUS3_recode$cGender)
```

## Race and Ethnicity

### MIDUS 1 Race and Ethnicity

```{r, echo = FALSE}
# summary(MIDUS1$A1SS7)
table(MIDUS1_recode$aRace, useNA = "always") 
M1RaceProportions <- MIDUS1_recode %>% 
  drop_na(aRace) %>% 
  ggplot(aes(x = aRace)) +
  geom_bar(fill = "lightblue", aes(y = (..count..)/sum(..count..))) +
  xlab("Race and Ethnicity") +
  scale_y_continuous(labels = scales::percent, name = "Proportion") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  M1RaceProportions
#
```

### MIDUS 2 Race and Ethnicity

```{r, echo = FALSE}
table(MIDUS2_recode$bRace, useNA = "always") 
M2RaceProportions <- MIDUS2_recode %>% 
  drop_na(bRace) %>% 
  ggplot(aes(x = bRace)) +
  geom_bar(fill = "lightblue", aes(y = (..count..)/sum(..count..))) +
  xlab("Race and Ethnicity") +
  scale_y_continuous(labels = scales::percent, name = "Proportion") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  M2RaceProportions
#
```

### MIDUS 3 Race and Ethnicity

```{r, echo = FALSE}
table(MIDUS3_recode$cRace, useNA = "always") 
M3RaceProportions <- MIDUS3_recode %>% 
  drop_na(cRace) %>% 
  ggplot(aes(x = cRace)) +
  geom_bar(fill = "lightblue", aes(y = (..count..)/sum(..count..))) +
  xlab("Race and Ethnicity") +
  scale_y_continuous(labels = scales::percent, name = "Proportion") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  M3RaceProportions
#
```
# Bivariate Distributions 

## Income and CAMs

### MIDUS 1
```{r, echo = FALSE}
ggplot(MIDUS1_recode, aes(x = aHHIncome, y = aCamSum)) +
  geom_point()
mosaic::cor.test(MIDUS1_recode$aHHIncome, MIDUS1_recode$aCamSum)
#
```

### MIDUS 2 

```{r, echo = FALSE}
ggplot(MIDUS2_recode, aes(x = bHHIncome, y = bCamSum)) +
  geom_point()
mosaic::cor.test(MIDUS2_recode$bHHIncome, MIDUS2_recode$bCamSum)
#
```

### MIDUS 3

```{r, echo = FALSE}
ggplot(MIDUS3_recode, aes(x = cHHIncome, y = cCamSum)) +
  geom_point()
mosaic::cor.test(MIDUS3_recode$cHHIncome, MIDUS3_recode$cCamSum)
#
```

## Education and CAMs

### MIDUS 1 

```{r, echo = FALSE}
ggplot(MIDUS1_recode, aes(aCamSum, color = aEducation)) +
  geom_freqpoly(binwidth = 1)
ggplot(MIDUS1_recode, aes(x = aCamSum)) +
  geom_histogram(binwidth = 1) +
  facet_grid(~aEducation)
histogram(~ MIDUS1_recode$aCamSum | MIDUS1_recode$aEducation, 
          data = MIDUS1_recode,
          xlab = "Education",
          main = "CAM Sum by Education", 
          breaks = seq(0, 32, 1))
#table(MIDUS1_recode$aCamSum, MIDUS1_recode$aEducation)
```

### MIDUS 2 

```{r, echo = FALSE}
ggplot(MIDUS2_recode, aes(bCamSum, color = bEducation)) +
  geom_freqpoly(binwidth = 1)
ggplot(MIDUS2_recode, aes(x = bCamSum)) +
  geom_histogram(binwidth = 1) +
  facet_grid(~bEducation)
histogram(~ MIDUS2_recode$bCamSum | MIDUS2_recode$bEducation, 
          data = MIDUS2_recode,
          xlab = "Education",
          main = "CAM Sum by Education", 
          breaks = seq(0, 32, 1))
#table(MIDUS2_recode$bCamSum, MIDUS2_recode$bEducation)
```

### MIDUS 3

```{r, echo = FALSE}
ggplot(MIDUS3_recode, aes(cCamSum, color = cEducation)) +
  geom_freqpoly(binwidth = 1)
ggplot(MIDUS3_recode, aes(x = cCamSum)) +
  geom_histogram(binwidth = 1) +
  facet_grid(~cEducation)
histogram(~ MIDUS3_recode$cCamSum | MIDUS3_recode$cEducation, 
          data = MIDUS3_recode,
          xlab = "Education",
          main = "CAM Sum by Education", 
          breaks = seq(0, 32, 1))
#table(MIDUS3_recode$cCamSum, MIDUS3_recode$cEducation)
```

## Gender and CAM Summed Index

### MIDUS 1 

```{r, echo = FALSE}
table(MIDUS1_recode$aCamSum,MIDUS1_recode$aGender)
ggplot(MIDUS1_recode, aes(aCamSum, color = aGender)) +
  geom_freqpoly(binwidth = 1)
ggplot(MIDUS1_recode, aes(x = aCamSum)) +
  geom_histogram(binwidth = 1) +
  facet_grid(~aGender) +
  ggtitle("Summed CAM Index by Gender") +
  xlab("Number of CAMs used in the last year")
#
```

### MIDUS 2

```{r, echo = FALSE}
table(MIDUS2_recode$bCamSum,MIDUS2_recode$bGender)
ggplot(MIDUS2_recode, aes(bCamSum, color = bGender)) +
  geom_freqpoly(binwidth = 1)
ggplot(MIDUS2_recode, aes(x = bCamSum)) +
  geom_histogram(binwidth = 1) +
  facet_grid(~bGender) +
  ggtitle("Summed CAM Index by Gender") +
  xlab("Number of CAMs used in the last year")
#
```

### MIDUS 3

```{r, echo = FALSE}
table(MIDUS3_recode$cCamSum,MIDUS3_recode$cGender)
ggplot(MIDUS3_recode, aes(cCamSum, color = cGender)) +
  geom_freqpoly(binwidth = 1)
ggplot(MIDUS3_recode, aes(x = cCamSum)) +
  geom_histogram(binwidth = 1) +
  facet_grid(~cGender) +
  ggtitle("Summed CAM Index by Gender") +
  xlab("Number of CAMs used in the last year")
#
```

## Race and CAM Summed Index

### MIDUS 1

```{r, echo = FALSE}
table(MIDUS1_recode$aCamSum,MIDUS1_recode$aRace)
ggplot(MIDUS1_recode, aes(aCamSum, color = aRace)) +
  geom_freqpoly(binwidth = 1)
ggplot(MIDUS1_recode, aes(x = aCamSum)) +
  geom_histogram(binwidth = 1) +
  facet_grid(~aRace) +
  ggtitle("Summed CAM Index by Race") +
  xlab("Number of CAMs used in the last year")
#
```

### MIDUS 2

```{r, echo = FALSE}
table(MIDUS2_recode$bCamSum,MIDUS2_recode$bRace)
ggplot(MIDUS2_recode, aes(bCamSum, color = bRace)) +
  geom_freqpoly(binwidth = 1)
ggplot(MIDUS2_recode, aes(x = bCamSum)) +
  geom_histogram(binwidth = 1) +
  facet_grid(~bRace) +
  ggtitle("Summed CAM Index by Race") +
  xlab("Number of CAMs used in the last year")
#
```

### MIDUS 3

```{r, echo = FALSE}
table(MIDUS3_recode$cCamSum,MIDUS3_recode$cRace)
ggplot(MIDUS3_recode, aes(cCamSum, color = cRace)) +
  geom_freqpoly(binwidth = 1)
ggplot(MIDUS3_recode, aes(x = cCamSum)) +
  geom_histogram(binwidth = 1) +
  facet_grid(~cRace) +
  ggtitle("Summed CAM Index by Race") +
  xlab("Number of CAMs used in the last year")
#
```

# Missing Cases

## MIDUS 1

### Table: Total and percent missing on each variable

```{r, echo = FALSE}
varlist <- c("aEducation", "aHHIncome", "aGender", "aRace", "aCamSum", "aCamEverAcupuncture", "aCamEverBiofeedback", "aCamEverChiropractor", "aCamEverEnergy", "aCamEverMoveTherapy", "aCamEverHerbTherapy", "aCamEverMegaVitamins", "aCamEverHomeopathy", "aCamEverHypnosis", "aCamEverImageTech", "aCamEverMassage", "aCamEverPray", "aCamEverMeditate", "aCamEverSpecialDiet", "aCamEverSpiritHeal", "aCamEverOtherHeal")
M1_miss <- MIDUS1_recode[varlist]
miss_var_summary(M1_miss)
#
```

### Missing summary for each observation

```{r, echo = FALSE}
vis_miss(M1_miss)
#
```

## Missing pattern intersections all variables 

```{r, echo = FALSE}
gg_miss_upset(M1_miss,
              nsets = n_var_miss(M1_miss))
#
```

### Missing pattern intersections for all CAMs

```{r, echo = FALSE}
varlist <- c("aCamSum", "aCamEverAcupuncture", "aCamEverBiofeedback", "aCamEverChiropractor", "aCamEverEnergy", "aCamEverMoveTherapy", "aCamEverHerbTherapy", "aCamEverMegaVitamins", "aCamEverHomeopathy", "aCamEverHypnosis", "aCamEverImageTech", "aCamEverMassage", "aCamEverPray", "aCamEverMeditate", "aCamEverSpecialDiet", "aCamEverSpiritHeal", "aCamEverOtherHeal")
M1_cammiss <- MIDUS1_recode[varlist]
gg_miss_upset(M1_cammiss,
              nsets = n_var_miss(M1_cammiss))
#
```

### Missing patterns by summed index and all social variables 

```{r, echo = FALSE}
varlist <- c("aCamSum", "aEducation", "aHHIncome", "aGender", "aRace")
M1_camIvsmiss <- MIDUS1_recode[varlist]
gg_miss_upset(M1_camIvsmiss, 
              nsets = n_var_miss(M1_camIvsmiss))
#
```

## MIDUS 2

### Table: Total and percent missing on each variable

```{r, echo = FALSE}
varlist <- c("bEducation", "bHHIncome", "bGender", "bRace", "bCamSum", "bCamEverAcupuncture", "bCamEverBiofeedback", "bCamEverChiropractor", "bCamEverEnergy", "bCamEverMoveTherapy", "bCamEverHerbTherapy", "bCamEverMegaVitamins", "bCamEverHomeopathy", "bCamEverHypnosis", "bCamEverImageTech", "bCamEverMassage", "bCamEverPray", "bCamEverMeditate", "bCamEverSpecialDiet", "bCamEverSpiritHeal", "bCamEverOtherHeal")
M2_miss <- MIDUS2_recode[varlist]
miss_var_summary(M2_miss)
#
```

### Missing summary for each observation

```{r, echo = FALSE}
vis_miss(M2_miss)
#
```

## Missing pattern intersections all variables 

```{r, echo = FALSE}
gg_miss_upset(M2_miss,
              nsets = n_var_miss(M2_miss))
#
```

### Missing pattern intersections for all CAMs

```{r, echo = FALSE}
varlist <- c("bCamSum", "bCamEverAcupuncture", "bCamEverBiofeedback", "bCamEverChiropractor", "bCamEverEnergy", "bCamEverMoveTherapy", "bCamEverHerbTherapy", "bCamEverMegaVitamins", "bCamEverHomeopathy", "bCamEverHypnosis", "bCamEverImageTech", "bCamEverMassage", "bCamEverPray", "bCamEverMeditate", "bCamEverSpecialDiet", "bCamEverSpiritHeal", "bCamEverOtherHeal")
M2_cammiss <- MIDUS2_recode[varlist]
gg_miss_upset(M2_cammiss,
              nsets = n_var_miss(M2_cammiss))
#
```

### Missing patterns by summed index and all social variables 

```{r, echo = FALSE}
varlist <- c("bCamSum", "bEducation", "bHHIncome", "bGender", "bRace")
M2_camIvsmiss <- MIDUS2_recode[varlist]
gg_miss_upset(M2_camIvsmiss, 
              nsets = n_var_miss(M2_camIvsmiss))
#
```

## MIDUS 3

### Table: Total and percent missing on each variable

```{r, echo = FALSE}
varlist <- c("cEducation", "cHHIncome", "cGender", "cRace", "cCamSum", "cCamEverAcupuncture", "cCamEverBiofeedback", "cCamEverChiropractor", "cCamEverEnergy", "cCamEverMoveTherapy", "cCamEverHerbTherapy", "cCamEverMegaVitamins", "cCamEverHomeopathy", "cCamEverHypnosis", "cCamEverImageTech", "cCamEverMassage", "cCamEverPray", "cCamEverMeditate", "cCamEverSpecialDiet", "cCamEverSpiritHeal", "cCamEverOtherHeal")
M3_miss <- MIDUS3_recode[varlist]
miss_var_summary(M3_miss)
#
```

### Missing summary for each observation

```{r, echo = FALSE}
vis_miss(M3_miss)
#
```

## Missing pattern intersections all variables 

```{r, echo = FALSE}
gg_miss_upset(M3_miss,
              nsets = n_var_miss(M3_miss))
#
```

### Missing pattern intersections for all CAMs

```{r, echo = FALSE}
varlist <- c("cCamSum", "cCamEverAcupuncture", "cCamEverBiofeedback", "cCamEverChiropractor", "cCamEverEnergy", "cCamEverMoveTherapy", "cCamEverHerbTherapy", "cCamEverMegaVitamins", "cCamEverHomeopathy", "cCamEverHypnosis", "cCamEverImageTech", "cCamEverMassage", "cCamEverPray", "cCamEverMeditate", "cCamEverSpecialDiet", "cCamEverSpiritHeal", "cCamEverOtherHeal")
M3_cammiss <- MIDUS3_recode[varlist]
gg_miss_upset(M3_cammiss,
              nsets = n_var_miss(M3_cammiss))
#
```

### Missing patterns by summed index and all social variables 

```{r, echo = FALSE}
varlist <- c("cCamSum", "cEducation", "cHHIncome", "cGender", "cRace")
M3_camIvsmiss <- MIDUS3_recode[varlist]
gg_miss_upset(M3_camIvsmiss, 
              nsets = n_var_miss(M3_camIvsmiss))
#
```