---
title: "LCA"
author: "Hannah Andrews"
date: "11/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
dir <- "C:/Users/hanna/Documents/git/AHL/R"
setwd(dir)

# Load Libraries ####
library(expss)
library(Hmisc)
library(dplyr)
library(summarytools)
library(magrittr)
library(tidyverse)
library(mosaic)
library(lattice)
library(ggplot2)
library(scales)
library(ggthemes)
library(LCA)
library(poLCA)

# Load all MIDUS data into the global environment ####
load(file = "MIDUS1.rda")
load(file = "MIDUS2.rda")
load(file = "MIDUS3.rda")

```
# MIDUS 1
max_II <- -100000
min_bic <- 100000
for (i in 2:10) {
  lc <- poLCA(f, M1LCAData, nclass = i, maxiter = 3000,
              graph = TRUE, 
              nrep = 10, verbose = TRUE, calc.se = TRUE)
  if(lc$bic < min_bic) {
    min_bic <- lc$bic
    LCA_best_model <- lc
  }
}
LCA_best_model
## 1 Class
```{r echo = FALSE}
# Source: https://statistics.ohlsen-web.de/latent-class-analysis-polca/
M1LCAData <- MIDUS1_recode %>% 
  dplyr::select(aCamEverBiofeedback, aCamEverChiropractor, aCamEverEnergy, aCamEverMoveTherapy,
            aCamEverHerbTherapy, aCamEverMegaVitamins, aCamEverHomeopathy, aCamEverHypnosis,
            aCamEverImageTech, aCamEverMassage, aCamEverPray, aCamEverMeditate,
            aCamEverSpecialDiet, aCamEverSpiritHeal)
f <- with(M1LCAData, cbind(aCamEverBiofeedback, aCamEverChiropractor, aCamEverEnergy, aCamEverMoveTherapy,
            aCamEverHerbTherapy, aCamEverMegaVitamins, aCamEverHomeopathy, aCamEverHypnosis,
            aCamEverImageTech, aCamEverMassage, aCamEverPray, aCamEverMeditate,
            aCamEverSpecialDiet, aCamEverSpiritHeal) ~ 1)

# Create a matrix to hold the best LCA model for each number of classes. 
lcamat <- matrix(NA, nrow = 10, ncol = 3)
colnames(lcamat) <- c("Maximum Log-Likelihood", "AIC", "BIC")

lc2 <- poLCA(f, M1LCAData, nclass = 2, graphs = TRUE,
             na.rm = TRUE, nrep = 500, maxiter = 3000)


```

```{r, echo = FALSE}
MIDUS1.lc2 <- poLCA(cbind(aCamEverBiofeedback, aCamEverChiropractor, aCamEverEnergy, aCamEverMoveTherapy,
            aCamEverHerbTherapy, aCamEverMegaVitamins, aCamEverHomeopathy, aCamEverHypnosis,
            aCamEverImageTech, aCamEverMassage, aCamEverPray, aCamEverMeditate,
            aCamEverSpecialDiet, aCamEverSpiritHeal) ~ 1, 
           data = as.data.frame(MIDUS1_recode), 
           nclass = 2,
           nrep = 10)
MIDUS1.lc3 <- poLCA(cbind(aCamEverBiofeedback, aCamEverChiropractor, aCamEverEnergy, aCamEverMoveTherapy,
            aCamEverHerbTherapy, aCamEverMegaVitamins, aCamEverHomeopathy, aCamEverHypnosis,
            aCamEverImageTech, aCamEverMassage, aCamEverPray, aCamEverMeditate,
            aCamEverSpecialDiet, aCamEverSpiritHeal) ~ 1, 
           data = as.data.frame(MIDUS1_recode), 
           nclass = 3,
           nrep = 10)
```
